### 2018-05-25 tex/manuscript.tex
some intro

\documentclass[twocolumn, linenumbers, superscriptaddress]{revtex4-1}

\usepackage{graphicx}
\usepackage{blindtext}

\begin{document}
	%abstract - approx. 150 words
	\begin{abstract}
		\blindtext
	\end{abstract}

	%title and author block
	{\title{Microstructure-derived flexibility in the acorn weevil exoskeleton}

	\date{\today}

	\author{M. Andrew Jansen}
		\email[corresponding author, email:~]{majanse1@asu.edu}
		\affiliation{School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA}
	\author{Nikhilesh Chawla}
		\affiliation{School for Engineering of Matter, Energy, and Transport, Arizona State University, Tempe, AZ 85287, USA}
	\author{Nico M. Franz}
		\affiliation{School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA}

	\maketitle
	}

%introduction - nature materials requires that introduction is given NO SECTION HEADER
%Brief explanation of relevant behavior and biology, prior work on ecology, head shape, cuticle mechanics, and rostral microstructure.

		We report novel modifications to the composite microstructure of the exoskeleton in the snout of acorn weevils (Coleoptera: Curculionidae) belonging to the genus \textit{Curculio} Linnaeus, 1756.

		As a weevil (snout beetle), members of the genus \textit{Curculio} are typified by the presence of a highly elongate structure on the head, called the rostrum (snout). 
		This structure is a hollow, cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts. 
		The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
		By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be thought of as a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \ref{sec:microstructure}).

	These modifications enable the snout to be flexed until straight, remaining within the elastic limits of the material, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure.
	Thus, the flexibility of the rostrum appears to be derived exclusively through modification of the composite architecture and fiber arrangement in the exoskeleton.
	Support for this hypothesis has come from three lines of evidence: first, examination of the cuticle microstructure across the length of the snout has revealed modifications to the composite structure of the rostrum; second, tensile testing of the structure has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure among species; and, third, fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure.
	We additionally report on the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
	To our knowledge, this is the first time such modifications have been reported for enhancing structural elasticity in the insect exoskeleton.

%4 sections of results ideally, nobody really uses subsections, 6 display items max.
	\section*{Microstructure}\label{sec:microstructure} 
		Display 1: Pictures of heads, macrofiber arrangement, and exo-endo ratio at base and apex, across species
		Explain how everything is laid out across the length of the rostrum, emphasizing that this is key to predicting and understanding the mechanical behavior of the snout during bending.

	\section*{Tensile testing and fracture mechanics}
		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.

	\section*{Fatigue testing of \textit{Curculio caryae}}

%conclusion	- with intro and "results" should be no more than 3000 words
	\section*{Conclusion}

%actual methods after references
	\section*{Methods}
		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.

%50 references (as a guideline)
	\begin{thebibliography}{50}
		\section*{References}	
		\bibitem{lamport94}
			Leslie Lamport,
			\textit{\LaTeX: a document preparation system},
			Addison Wesley, Massachusetts,
			2nd edition,
			1994.

	\end{thebibliography}

%brief and 'without effusive statements'
%thank everyone who helped with data collection, especially jason (tensile testing), sridhar (if I use micro-CT scan data), dave (SEM), charlie (specimens), paige (if I use confocal data), Sal (if I use histology/staining data or confocal data, he helped with specimen clearing).
%Financial support/ funding sources/ grants.
	\begin{acknowledgements}

	\end{acknowledgements}

%author contributions
%MAJ: did tensile tests, SEM, confocal (if applicable), data analysis.
%MAJ, NC: designed research, micro-CT (if applicable)
%MAJ, NMF: field collection of test specimens, procurement of specimen loans, specimen prep.
%MAJ, NC, NMF: All authors contributed to the writing, review, and revision of the manuscript.
	\section*{Author contributions}

	\section*{Additional information}
		Supplementary information is available in the online version of the paper.
		Reprints and permissions information is available online at www.nature.com/reprints.
		Correspondence and requests for materials should be addressed to M.A.J.

	\section*{Competing financial interests}
		The authors declare no competing financial interests.

	\newpage

%provide detailed methods here, subsection for each method, statistical analysis, statements of code and data avilability.
	\section*{Methods}
		\subsection*{Method 1}

		\subsection*{Method 2}

		\subsection*{Method 3}

		\subsection*{Statistical analysis}

		\subsection*{Code availability}

		\subsection*{Data availability}

%references for methods only
	\begin{thebibliography}{50}
		\section*{References}	
		\bibitem{lamport94}
			Leslie Lamport,
			\textit{\LaTeX: a document preparation system},
			Addison Wesley, Massachusetts,
			2nd edition,
			1994.

	\end{thebibliography}
\end{document}

### 2018-06-17 tex/manuscript.tex
rough draft intro

@@
 		The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
-		By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be thought of as a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \ref{sec:microstructure}).
+		By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \ref{sec:microstructure}).
+		We have found that the composite profile of the rostrum is strongly differentiated from the head capsule and other body parts, with modification of both the relative layer thicknesses and fiber orientation angles of cuticle regions (viz.
+exocuticle and endocuticle), which we describe in detail below.
 
-	These modifications enable the snout to be flexed until straight, remaining within the elastic limits of the material, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure.
-	Thus, the flexibility of the rostrum appears to be derived exclusively through modification of the composite architecture and fiber arrangement in the exoskeleton.
-	Support for this hypothesis has come from three lines of evidence: first, examination of the cuticle microstructure across the length of the snout has revealed modifications to the composite structure of the rostrum; second, tensile testing of the structure has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure among species; and, third, fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure.
+	These modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	Thus, the flexibility of the rostrum appears to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
+	Support for this hypothesis has come from three lines of evidence: first, examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species; second, tensile testing of the structure has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species; and, third, fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
 	We additionally report on the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.

### 2018-06-18 tex/manuscript.tex
intro edits

@@
 \usepackage{blindtext}
+\usepackage{gensymb}
+\usepackage{textgreek}
+\usepackage{xcolor}
+\usepackage[colorlinks]{hyperref}
+	\hypersetup{
+	colorlinks,
+	linkcolor={red!50!black},
+	citecolor={blue!60!black},
+	urlcolor={blue!40!black}
+	}
+\usepackage{cleveref}
 
@@
 		The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
-		By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \ref{sec:microstructure}).
+		By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \cref{sec:microstructure}).
+		Acorn weevils use this structure to excavate sites for egg-laying (oviposition) and feeding on a variety of fruits, including acorns, Japanese camellia, hazelnuts, pecans, chestnuts, and chinquapins.
+
+		During oviposition, a female engages in a unique "drilling" behavior that causes significant, apparently elastic, deformation of the rostrum.
+		The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the bore-hole.
+		Once the apex of the snout is inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
+		The female will maintain tension on the rostrum in this position, continuing to ingest the substrate and rotate around the bore-hole, while slowly inserting the rostrum further into the excavated channel.
+		Once the rostrum is fully inserted, usually up to the eyes, she will pull her snout from the bore-hole and deposit several eggs into the site.
+		By maintaining constant tension on the rostrum and rotating around the bore-hole, the female is able to flex the snout into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
+		While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending (over 90$\degree$) incurred during the process of oviposition.
+
 		We have found that the composite profile of the rostrum is strongly differentiated from the head capsule and other body parts, with modification of both the relative layer thicknesses and fiber orientation angles of cuticle regions (viz.
 exocuticle and endocuticle), which we describe in detail below.
+	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	Thus, the flexibility of the rostrum appears to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
 
-	These modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
-	Thus, the flexibility of the rostrum appears to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
-	Support for this hypothesis has come from three lines of evidence: first, examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species; second, tensile testing of the structure has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species; and, third, fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
-	We additionally report on the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
-	To our knowledge, this is the first time such modifications have been reported for enhancing structural elasticity in the insect exoskeleton.
+	Support for this hypothesis has come from three lines of evidence: 
+	\begin{enumerate}
+	\item Examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species.
+	\item Tensile testing of the rostrum has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species.
+	\item Fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
+	\end{enumerate}
+
+	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
+	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.
 
 %4 sections of results ideally, nobody really uses subsections, 6 display items max.
-	\section*{Microstructure}\label{sec:microstructure} 
-		Display 1: Pictures of heads, macrofiber arrangement, and exo-endo ratio at base and apex, across species
+	\section{Microstructure}\label{sec:microstructure} 
+		Display 1: Pictures of heads, macrofiber arrangement, and exo-endo ratio at base and apex, across species.
 		Explain how everything is laid out across the length of the rostrum, emphasizing that this is key to predicting and understanding the mechanical behavior of the snout during bending.
+		No change in total thickness, only relative layer thickness,
+		No
 
-	\section*{Tensile testing and fracture mechanics}
+	\section{Tensile testing and fracture mechanics}
 		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
 
-	\section*{Fatigue testing of \textit{Curculio caryae}}
+	\section{Fatigue testing of \textit{Curculio caryae}}
 
 %conclusion	- with intro and "results" should be no more than 3000 words
-	\section*{Conclusion}
+	\section{Conclusion}
 
 %actual methods after references
-	\section*{Methods}
+	\section{Methods}
 		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.

### 2018-06-18 tex/manuscript.tex
intro cleanup
began sec 1

@@
 \usepackage{gensymb}
-\usepackage{textgreek}
 \usepackage{xcolor}
@@
 	As a weevil (snout beetle), members of the genus \textit{Curculio} are typified by the presence of a highly elongate structure on the head, called the rostrum (snout). 
-		This structure is a hollow, cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts. 
+	This structure is a hollow, strongly curved (over 90$\degree$), cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
 	The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
@@
 	During oviposition, a female engages in a unique "drilling" behavior that causes significant, apparently elastic, deformation of the rostrum.
-		The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the bore-hole.
-		Once the apex of the snout is inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
+	The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the perimeter of the bore-hole.
+	Once the apex of the snout is fully inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
 	The female will maintain tension on the rostrum in this position, continuing to ingest the substrate and rotate around the bore-hole, while slowly inserting the rostrum further into the excavated channel.
@@
 	By maintaining constant tension on the rostrum and rotating around the bore-hole, the female is able to flex the snout into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
-		While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending (over 90$\degree$) incurred during the process of oviposition.
+	While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending incurred during the process of oviposition.
 
@@
 %4 sections of results ideally, nobody really uses subsections, 6 display items max.
-	\section{Microstructure}\label{sec:microstructure} 
+	\section{Microstructure of the rostrum}\label{sec:microstructure} 
 		Display 1: Pictures of heads, macrofiber arrangement, and exo-endo ratio at base and apex, across species.
 		Explain how everything is laid out across the length of the rostrum, emphasizing that this is key to predicting and understanding the mechanical behavior of the snout during bending.
-		No change in total thickness, only relative layer thickness,
-		No
+		No change in total thickness, only relative layer thickness, no resilin
+		outline general arrangement of normal beetle/weevil cuticle (copy/edit from previous)
+		detail specifics of head capsule cuticle 
+		explain modification of rostral cuticle
 

### 2018-06-26 tex/manuscript.tex
added gitignore folder

--- 2018-06-18 tex/manuscript.tex
@@
 \usepackage[colorlinks]{hyperref}
-	\hypersetup{
-	colorlinks,
+	\hypersetup{colorlinks,
 	linkcolor={red!50!black},
@@
 		No change in total thickness, only relative layer thickness,
-		No
+		No resilin either.
 

### 2018-11-09 tex/manuscript.tex
edits
edit
some text

--- 2018-06-18 tex/manuscript.tex
@@
 \usepackage{cleveref}
+\graphicspath{{figures/}}
 
@@
 	%title and author block
-	{\title{Microstructure-derived flexibility in the acorn weevil exoskeleton}
+	{\title{Microstructure-derived strength in the acorn weevil exoskeleton}
 
@@
 exocuticle and endocuticle), which we describe in detail below.
-	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
-	Thus, the flexibility of the rostrum appears to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
+	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
 
@@
 		Explain how everything is laid out across the length of the rostrum, emphasizing that this is key to predicting and understanding the mechanical behavior of the snout during bending.
+		No change in total thickness, only relative layer thickness,
+		No resilin either.
 		No change in total thickness, only relative layer thickness, no resilin
@@
 	\section*{Author contributions}
+		\begin{description}
+		\item[Andrew Jansen] Conducted sectioning and staining, microscopy and imaging, tensile and fatigue testing, statistical analysis, and participated in manuscript preparation.
+		\item[Nikhilesh Chawla] Facilitated microscopy, tensile and fatigue testing, and participated in manuscript preparation.
+		\item[Nico Franz] Facilitated specimen acquisition and imaging, participated in manuscript preparation.
+		\end{description}
 
@@
 	\section*{Methods}
-		\subsection*{Method 1}
+		\subsection*{Histological sectioning}
 
-		\subsection*{Method 2}
+		\subsection*{Tensile and fatigue testing}
 
-		\subsection*{Method 3}
+		\subsection*{Specimen imaging and microscopy}
 
 		\subsection*{Statistical analysis}
+			\paragraph*{General approach}
+				To explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models to the data using maximum likelihood estimation.
+				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
+				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion model of trait evolution along the phylogeny.
+				Response variables and covariates were natural-log transformed, as needed, to ensure model residuals were normally distributed and homoscedastic.
+				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test and $R^{2}_{\sigma}$-difference test between the fully-specified model and a model lacking the phylogenetic effect.
+			\paragraph*{Hypothesis testing}
+				The following three hypotheses were tested using PGLMMs fitted using ML estimation:
+				\begin{enumerate}
+				\item The maximum sustained tensile force is proportional to the cross-sectional area of the endocuticle, and \emph{not} that of the exocuticle.
+				\item The ultimate tensile strength of the samples is inversely proportional to the ratio of exocuticle to endocuticle at the location of fracture.
+				\item Young's modulus of the samples is proportional to the length of the snout.
+				\end{enumerate}
+
+				We fitted a fully-specified model with the cross-sectional area of endocuticle and exocuticle at the site of fracture as fixed effects, including an interaction term, and with maximum tensile force sustained prior to fracture as a response variable.
+				This model was then compared to models with only cross-sectional area of either endocuticle or exocuticle as the sole fixed effect in the model.
+				We then tested the first hypothesis by using likelihood-ratio tests and $R^{2}_{\beta*}$-difference tests between each of the three models.
+
+				The hypothesis that
+
+			\paragraph*{Model selection and fitting}
+			\paragraph*{Estimating phylogenetic signal}
 

### 2019-01-10 tex/manuscript.tex
added microstructure section

@@
 		\affiliation{School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA}
+	\author{Jason Williams}
+		\affiliation{School for Engineering of Matter, Energy, and Transport, Arizona State University, Tempe, AZ 85287, USA}
 	\author{Nikhilesh Chawla}
@@
 %Brief explanation of relevant behavior and biology, prior work on ecology, head shape, cuticle mechanics, and rostral microstructure.
+
+%650 words here
 
@@
 	Support for this hypothesis has come from three lines of evidence:
+
 	\begin{enumerate}
@@
 
-%4 sections of results ideally, nobody really uses subsections, 6 display items max.
-	\section{Microstructure of the rostrum}\label{sec:microstructure} 
-		Display 1: Pictures of heads, macrofiber arrangement, and exo-endo ratio at base and apex, across species.
-		Explain how everything is laid out across the length of the rostrum, emphasizing that this is key to predicting and understanding the mechanical behavior of the snout during bending.
-		No change in total thickness, only relative layer thickness,
-		No resilin either.
-		No change in total thickness, only relative layer thickness, no resilin
-		outline general arrangement of normal beetle/weevil cuticle (copy/edit from previous)
-		detail specifics of head capsule cuticle 
-		explain modification of rostral cuticle
+	MIGHT BE GOOD TO SIGNPOST HERE...
+we discuss each of these lines of evidence in the sections below beginning with a consideration of the impact of microstructure, then we talk about mechanical testing, blah blah.
+Emphasize that this is key to predicting and understanding the mechanical behavior of the snout during bending.
+%stick to 600 words for each results section and 500 words for conclusions
+	\section{Microstructure of the rostrum}\label{sec:microstructure} %750 words here (1400 total)
+		The cuticle of arthropods, including insects, is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
+		The cuticle is strongly influenced by its chemical properties, especially the degree of tanning and overall water content, but the arrangement of the embedded chitin fibers is primarily responsible for the its unique mechanical behavior \cite{Klocke2011,Vincent2004}.
+		Accordingly, the cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.
 
-	\section{Tensile testing and fracture mechanics}
+		The insect cuticle can generally be divided into three regions (Figs.~INSERT-REFS-HERE), including (1) endocuticle, which is the most compliant and innermost region; and (3) exocuticle, which is the outermost region, usually the stiffest and hardest, and coated by a waxy secretion (epicuticle) that is not of great mechanical significance in most groups.
+		Between these regions lies (2) the mesocuticle, which is similar in microstructure to exocuticle, but exhibits a lesser degree of tanning (sclerotization) and usually serves as a transition zone to the endocuticle \cite{Klocke2011,Vincent1982,Vincent2004}.
+
+		In general, these regions are confluent, and not necessarily sharply defined; for instance, the mesocuticle is not always evident, and is sometimes considered part of the exocuticle.
+		In beetles, both the exocuticle and mesocuticle (when present) are laminate, and have numerous laminae of unidirectional chitin fibers, one layer thick (2-4nm each), embedded in a proteinaceous matrix.
+		These layers are stacked at a more or less constant angle to each other in a helicoidal arrangement, referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
+		This microstructural arrangement allows the cuticle to exhibit transverse isotropy, or complete isotropy in some cases, despite the strong anisotropy of $\alpha$-chitin.
+
+		Conversely, the endocuticle of Coleoptera appears to take advantage of this anisotropy to improve fracture resistance of the exoskeleton.
+		Beetles in particular have a highly modified endocuticle made up of large (5-20$\mu$m OD) unidirectional bundles of chitin, called ``balken'' (German for beam, strut) or macrofibers; these macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
+		Typically, the endocuticle contains several layers of macrofibers, in which adjacent layers form pairs. 
+		The layers within each pair are pseudo-orthogonal (i.e., angled approx.
+$90\degree$ to each other, see Figs.~INSERT-REFS-HERE), while the stacking angle between pairs is typically much shallower, although some other configurations have been observed \cite{Hepburn1973,Kamp2010}.
+		It is thought that this arrangement of the macrofibers contributes to the high toughness of beetle exoskeletons by inhibiting crack formation and propagation between successive layers \cite{Kamp2010,Kamp2015,Hepburn1973}.
+
+		We have found that in \textit{Curculio}, the head capsule (as in the rest of the body) exhibits this general arrangement of the endocuticle, with an angle of approximately $30\degree$ between successive pseudo-orthogonal plies.
+		Additionally, the head capsule has roughly equal thicknesses of exocuticle and endocuticle along its surface in cross section; qualitatively, manipulating this part of the head with forceps quickly reveals that this part of the head is quite rigid.
+
+		By contrast, the region beyond the scrobe (antennal channel) is quite flexible, even in fully desiccated specimens.
+		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has substantially different composite properties in two key traits (see Fig~INSERT-REFS-HERE):
+
+		\begin{enumerate}
+			\item The exocuticle is reduced to a thin shell, with the endocuticle thickened to offset this reduction.
+			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approx. $45\degree$ to the longitudinal axis of the snout.
+		\end{enumerate}
+
+		In previous work we identified these modifications to the composite arrangement of the cuticle, but only within the species \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
+		These same modifications to the cuticle of the rostral apex have now been uncovered in a further six phylogenetically disparate species, listed below, thus indicating that this is likely a genus-wide phenomenon.
+
+		Notably, we could find no evidence of resilin, as indicated by florescent microscopy, anywhere in the cuticle of the head (including the rostrum).
+		We also observed no difference in the total thickness of the cuticle between the head capsule and rostral apex, only differences in relative layer thicknesses.
+		In the next section, we demonstrate that that the endocuticle does not vary in tensile strength across the head, making it unlikely that there are significant differences in sclerotization or chitin composition within the exocuticle and endocuticle that might result in a more compliant cuticle.
+		We therefor hypothesize that these differences in the microstructure of the cuticle in different regions of the head are solely responsible for the observed flexibility of the snout.
+
+	\section{Tensile testing and fracture mechanics} %shoot for 550 words
 		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
 
-	\section{Fatigue testing of \textit{Curculio caryae}}
+	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 550 words
 
-%conclusion	- with intro and "results" should be no more than 3000 words
-	\section{Conclusion}
+	\section{Conclusion} %shoot for 500 words
 
@@
 
-%50 references (as a guideline)
+%50 references (as a guideline, in order of appearance in text)
 	\begin{thebibliography}{50}
 		\section*{References}	
-		\bibitem{lamport94}
-			Leslie Lamport,
-			\textit{\LaTeX: a document preparation system},
-			Addison Wesley, Massachusetts,
-			2nd edition,
-			1994.
+			\bibitem{Klocke2011}
+				Klocke, D. \& Schmitz, H.
+				Water as a major modulator of the mechanical properties of insect cuticle.
+				\textit{Acta Biomater.}
+				\textbf{7,}
+				2935--2942
+				(2011).
+
+			\bibitem{Vincent1982}
+				Vincent, J. F. V.
+				\textit{Structural biomaterials}
+				(Halsted Press,
+				New York,
+				1982).
+
+			\bibitem{Vincent2004}
+				Vincent, J. F. V. \& Wegst, U. G. K.
+				Design and mechanical properties of insect cuticle.				
+				\textit{Arthropod Struct. Dev.}
+				\textbf{33:3,}
+				187--199,
+				(2004).
+
+			\bibitem{Nikolov2011}
+				Nikolov, S. et al.
+				Robustness and optimal use of design principles of arthropod exoskeletons studied by ab initio-based multiscale simulations.
+				\textit{J. Mech. Behav. Biomed. Mater.}
+				\textbf{4:2,}
+				129--145,
+				(2011).
+
+			\bibitem{Nikolov2010}
+				Nikolov, S. et al.
+				Revealing the design principles of high-performance biological composites using Ab initio and multiscale simulations: The example of lobster cuticle.
+				\textit{Adv. Mater.}
+				\textbf{22:4,}
+				519--526,
+				(2010).
+
+			\bibitem{Blackwell1980}
+				Blackwell, J. \& Weih, M.
+				Structure of chitin-protein complexes: ovipositor of the ichneumon fly \textit{Megarhyssa}.
+				\textit{J. Mol. Biol.}
+				\textbf{137:1,}
+				49--60,
+				(1980).
+
+			\bibitem{Bouligand1972}
+				Bouligand, Y.
+				Twisted fibrous arrangements in biological materials and cholesteric mesophases.
+				\textit{Tissue Cell}
+				\textbf{4:2,}
+				189--217,
+				(1972).
+
+			\bibitem{Neville1976}
+				Neville, A. C., Parry, D. A. \& Woodhead-Galloway, J.
+				The chitin crystallite in arthropod cuticle.
+				\textit{J. Cell Sci.}
+				\textbf{21:1,}
+				73--82,
+				(1976).
+
+			\bibitem{Cheng2009}
+				Cheng, L., Wang, L., \& Karlsson, A. M.
+				Mechanics-based analysis of selected features of the exoskeletal microstructure of \textit{Popillia japonica}.
+				\textit{Mater. Res.}
+				\textbf{24:11,}
+				3253--3267,
+				(2009).
+
+			\bibitem{Hepburn1973}
+				Hepburn, H. R. \& Ball, A.
+				On the structure and mechanical properties of beetle shells.
+				\textit{J. Mater. Sci.}
+				\textbf{8:5,}
+				618--623,
+				(1973).
+
+			\bibitem{Kamp2010}
+				van de Kamp, T. \& Greven, H.
+				On the architecture of beetle elytra.
+				\textit{Entomol. Heute}
+				\textbf{22,}
+				191--204,
+				(2010).
+
+			\bibitem{Kamp2015}
+				van de Kamp, T., Riedel, A. \& Greven, H.
+				Micromorphology of the elytral cuticle of beetles, with an emphasis on weevils (Coleoptera : Curculionoidea)
+				\textit{Arthropod Struct. Dev.}
+				\textbf{45:1,}
+				14--22,
+				(2015).
 
@@
 %brief and 'without effusive statements'
-%thank everyone who helped with data collection, especially jason (tensile testing), sridhar (if I use micro-CT scan data), dave (SEM), charlie (specimens), paige (if I use confocal data), Sal (if I use histology/staining data or confocal data, he helped with specimen clearing).
+%thank everyone who helped with data collection, dave (SEM), charlie (specimens), paige (if I use confocal data), Sal (if I use histology/staining data or confocal data, he helped with specimen clearing).
 %Financial support/ funding sources/ grants.
@@
 		\item[Andrew Jansen] Conducted sectioning and staining, microscopy and imaging, tensile and fatigue testing, statistical analysis, and participated in manuscript preparation.
+		\item[Jason Williams] Conducted tensile and fatigue testing, participated in manuscript preparation.
 		\item[Nikhilesh Chawla] Facilitated microscopy, tensile and fatigue testing, and participated in manuscript preparation.

### 2019-01-10 tex/manuscript.tex
finished rough microstructure and rough figures

@@
 %stick to 600 words for each results section and 500 words for conclusions
-	\section{Microstructure of the rostrum}\label{sec:microstructure} %750 words here (1400 total)
+	\section{Microstructure of the rostrum}\label{sec:microstructure}
+	%800 words here (1500 total, shorten later)
 		The cuticle of arthropods, including insects, is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
@@
 		By contrast, the region beyond the scrobe (antennal channel) is quite flexible, even in fully desiccated specimens.
-		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has substantially different composite properties in two key traits (see Fig~INSERT-REFS-HERE):
+		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has substantially different composite properties from the head capsule in two key traits (see Fig~INSERT-REFS-HERE):
 
@@
 
+		The cuticle of the rostral base and region near the scrobe exhibits a gradual transition between these cuticle arrangements, such that there is a smoothly discontinuous profile in the composite properties of the cuticle along an anterior-posterior gradient.
 		In previous work we identified these modifications to the composite arrangement of the cuticle, but only within the species \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
@@
 		In the next section, we demonstrate that that the endocuticle does not vary in tensile strength across the head, making it unlikely that there are significant differences in sclerotization or chitin composition within the exocuticle and endocuticle that might result in a more compliant cuticle.
-		We therefor hypothesize that these differences in the microstructure of the cuticle in different regions of the head are solely responsible for the observed flexibility of the snout.
+		We therefore hypothesize that these differences in the microstructure of the cuticle in different regions of the head are solely responsible for the observed flexibility of the snout.
 
-	\section{Tensile testing and fracture mechanics} %shoot for 550 words
+	\section{Tensile testing and fracture mechanics} %shoot for 600 words
 		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
 
-	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 550 words
+	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
 

### 2019-01-12 tex/manuscript.tex
revised microstructure

--- 2019-01-10 tex/manuscript.tex
@@
 
-%650 words here
+%~650 words here
 
@@
 Emphasize that this is key to predicting and understanding the mechanical behavior of the snout during bending.
+
 %stick to 600 words for each results section and 500 words for conclusions
-	\section{Microstructure of the rostrum}\label{sec:microstructure} %750 words here (1400 total)
-		The cuticle of arthropods, including insects, is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
-		The cuticle is strongly influenced by its chemical properties, especially the degree of tanning and overall water content, but the arrangement of the embedded chitin fibers is primarily responsible for the its unique mechanical behavior \cite{Klocke2011,Vincent2004}.
-		Accordingly, the cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.
+	\section{Microstructure of the rostrum}\label{sec:microstructure}
+	%~700 words here (1500 total, shorten later)
+		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
+		The arrangement of the embedded chitin fibers is primarily responsible for the mechanical behavior of the cuticle, although this is modulated by the degree of tanning and water retention.
+\cite{Klocke2011,Vincent2004}.
+		Accordingly, cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.
 
-		The insect cuticle can generally be divided into three regions (Figs.~INSERT-REFS-HERE), including (1) endocuticle, which is the most compliant and innermost region; and (3) exocuticle, which is the outermost region, usually the stiffest and hardest, and coated by a waxy secretion (epicuticle) that is not of great mechanical significance in most groups.
-		Between these regions lies (2) the mesocuticle, which is similar in microstructure to exocuticle, but exhibits a lesser degree of tanning (sclerotization) and usually serves as a transition zone to the endocuticle \cite{Klocke2011,Vincent1982,Vincent2004}.
+		Insect cuticle is divisible into three regions (Figs.~INSERT-REFS-HERE), including (1) endocuticle, which is the most compliant and innermost region; and (3) exocuticle, which is the stiffest, hardest, and outermost region.
+		Between these lies (2) mesocuticle, which is similar in microstructure to exocuticle, but less sclerotized (tanned), usually acting as a thin transition zone \cite{Klocke2011,Vincent1982,Vincent2004}.
 
-		In general, these regions are confluent, and not necessarily sharply defined; for instance, the mesocuticle is not always evident, and is sometimes considered part of the exocuticle.
-		In beetles, both the exocuticle and mesocuticle (when present) are laminate, and have numerous laminae of unidirectional chitin fibers, one layer thick (2-4nm each), embedded in a proteinaceous matrix.
+		In general, these regions are confluent, and not necessarily sharply defined; for instance, mesocuticle is not always evident, and is sometimes considered part of the exocuticle.
+		In beetles, both exocuticle and mesocuticle (when present) are laminate, and have numerous laminae of unidirectional chitin fibers, each layer a single fiber thick (2-4nm), embedded in a proteinaceous matrix.
 		These layers are stacked at a more or less constant angle to each other in a helicoidal arrangement, referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
-		This microstructural arrangement allows the cuticle to exhibit transverse isotropy, or complete isotropy in some cases, despite the strong anisotropy of $\alpha$-chitin.
+		This microstructure allows exocuticle and mesocuticle to exhibit transverse isotropy, or complete isotropy in some cases, despite the strong anisotropy of $\alpha$-chitin.
 
 		Conversely, the endocuticle of Coleoptera appears to take advantage of this anisotropy to improve fracture resistance of the exoskeleton.
-		Beetles in particular have a highly modified endocuticle made up of large (5-20$\mu$m OD) unidirectional bundles of chitin, called ``balken'' (German for beam, strut) or macrofibers; these macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
-		Typically, the endocuticle contains several layers of macrofibers, in which adjacent layers form pairs. 
-		The layers within each pair are pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other, see Figs.~INSERT-REFS-HERE), while the stacking angle between pairs is typically much shallower, although some other configurations have been observed \cite{Hepburn1973,Kamp2010}.
-		It is thought that this arrangement of the macrofibers contributes to the high toughness of beetle exoskeletons by inhibiting crack formation and propagation between successive layers \cite{Kamp2010,Kamp2015,Hepburn1973}.
+		Beetles in particular have a highly modified endocuticle comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
+		Macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
+		Typically, endocuticle contains several layers of macrofibers, with the adjacent layers forming pairs. 
+		The layers \emph{within} each pair are pseudo-orthogonal (i.e., angled approx.
+$90\degree$ to each other, see Figs.~INSERT-REFS-HERE), while the stacking angle \emph{between} pairs is typically acute, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
+		It is thought that this arrangement of the macrofibers contributes to the toughness of beetle exoskeletons by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
 
-		We have found that in \textit{Curculio}, the head capsule (as in the rest of the body) exhibits this general arrangement of the endocuticle, with an angle of approximately $30\degree$ between successive pseudo-orthogonal plies.
-		Additionally, the head capsule has roughly equal thicknesses of exocuticle and endocuticle along its surface in cross section; qualitatively, manipulating this part of the head with forceps quickly reveals that this part of the head is quite rigid.
+		We have found that in \textit{Curculio}, the head capsule (which is similar to the rest of the body) fits the general profile of endocuticle, with an angle of approximately $30\degree$ between successive pseudo-orthogonal plies.
+		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle; compared to the snout, this part of the head is fairly rigid.
 
 		By contrast, the region beyond the scrobe (antennal channel) is quite flexible, even in fully desiccated specimens.
-		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has substantially different composite properties in two key traits (see Fig~INSERT-REFS-HERE):
+		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has a different composite structure (see Fig~INSERT-REFS-HERE) than the head capsule, namely:
 
@@
 			\item The exocuticle is reduced to a thin shell, with the endocuticle thickened to offset this reduction.
-			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approx. $45\degree$ to the longitudinal axis of the snout.
+			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approx. $\pm45\degree$ to the longitudinal axis of the snout.
 		\end{enumerate}
 
-		In previous work we identified these modifications to the composite arrangement of the cuticle, but only within the species \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
-		These same modifications to the cuticle of the rostral apex have now been uncovered in a further six phylogenetically disparate species, listed below, thus indicating that this is likely a genus-wide phenomenon.
+		Correspondingly, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.		
+		In previous work we identified these modifications to the composite structure of the cuticle, but only within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
+		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (listed below), indicating that this is likely a genus-wide trait.
 
 		Notably, we could find no evidence of resilin, as indicated by florescent microscopy, anywhere in the cuticle of the head (including the rostrum).
-		We also observed no difference in the total thickness of the cuticle between the head capsule and rostral apex, only differences in relative layer thicknesses.
-		In the next section, we demonstrate that that the endocuticle does not vary in tensile strength across the head, making it unlikely that there are significant differences in sclerotization or chitin composition within the exocuticle and endocuticle that might result in a more compliant cuticle.
-		We therefor hypothesize that these differences in the microstructure of the cuticle in different regions of the head are solely responsible for the observed flexibility of the snout.
+		We also observed no difference in total cuticle thickness between the head capsule and rostral apex, only differences in relative layer thicknesses.
+		Below we demonstrate that that the endocuticle does not vary in tensile strength across the rostrum, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
+		The available evidence therefore suggests that the relative flexibility of the snout is solely derived from the composite profile of its cuticle.
 
-	\section{Tensile testing and fracture mechanics} %shoot for 550 words
+	\section{Tensile testing and fracture mechanics} %shoot for 600 words
+		Based on our understanding of the microstructure of the cuticle, it is possible to make several predictions about the mechanical behavior of the snout.
+		First, we can predict that the endocuticle will contribute more to the tensile strength of the rostrum than the exocuticle, because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
+		Second, the endocuticle will be less brittle than the exocuticle, therefore we should expect that fractures will initiate in the exocuticle.
 		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
+		In addition to making the cuticle flexible, this profile makes it less likely to break, suggesting a possible means by which the system evolved (negative selection of breakage, maybe say this in conclusions).
 
-	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 550 words
+		Outline:
+		Intro
+			To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the heads of 6 species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
+		Tensile testing
+			We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
+			Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
+			 All of these associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
+
+			This strongly suggests that there is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology.
+
+	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
 

### 2019-01-15 tex/manuscript.tex
skeleton of tensile, fracture, and fatigue

@@
 
-	\section{Tensile testing and fracture mechanics} %shoot for 600 words
-		Based on our understanding of the microstructure of the cuticle, it is possible to make several predictions about the mechanical behavior of the snout.
-		First, we can predict that the endocuticle will contribute more to the tensile strength of the rostrum than the exocuticle, because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
-		Second, the endocuticle will be less brittle than the exocuticle, therefore we should expect that fractures will initiate in the exocuticle.
-		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
-		In addition to making the cuticle flexible, this profile makes it less likely to break, suggesting a possible means by which the system evolved (negative selection of breakage, maybe say this in conclusions).
+	\section{Tensile testing of the rostrum} %shoot for 600 words
+		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
+		Although the heads were rehydrated by immersion in de-ionized water for 24 hours, we observed comparatively brittle fracture (but see below), in contrast to GIVE EXAMPLES HERE.
+		The tensile behavior for the cuticle of these weevils is a result of its microstructure, which lacks pores, etc (give reasons).
+		THIS IS ALSO A GOOD SPOT TO SIGNPOST WHAT I FOUND IN GENERAL.
 
-		Outline:
-		Intro
-			To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the heads of 6 species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
-		Tensile testing
+		\subsection{Tensile behavior}
 			We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
 			Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
-			 All of these associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
+			In other words, UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
+			These associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
 
-			This strongly suggests that there is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology.
+			These data have three important implications:
+			\begin{enumerate}
+				\item There is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology. 
+				\item The endocuticle contributes more to the tensile strength of the rostrum than the exocuticle, possibly because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
+				\item Thus, in addition to making the cuticle more flexible, the altered apical composite profile makes it less likely to break, suggesting a possible means by which the system evolved (via negative selection of breakage, maybe say this in conclusions).
+			\end{enumerate}
+
+			Finally, we observed that Young's modulus of elasticity is higher in specimens with a longer snout.
+			This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) would need to be more flexible to avoid fracture during oviposition.
+			Based on preliminary confocal microscopy data, we speculate that this phenomenon may be the result of a longer scrobe/transition/gradient from the basal profile to the apical profile in longer rostra.
+			A decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
+			[OR SAY MORE CONCISELY THAT: based on prelim CLSM data we believe that the snout is reinforced against bucking with thicker exocuticle in the base of these species; as a result, the higher volume fraction of exocuticle in the snout increases young's modulus for such species]
+
+		\subsection{Fractography}
+			Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).
+
+			In cross-section, the exocuticle consistently presented a smooth, nearly continuous fracture surface, indicative of relatively brittle fracture, likely due to the microarchitecture and resultant transverse isotropy of the Bouligand structure (explain this).
+			Endocuticle, on the other hand exhibited severe delamination, ply-splitting, and fiber pulling, consistent with viscoelastic/plastic behavior shown in previous studies and congruent with theoretical consideration of the microstructure of this material (name something here).
+			These patterns indicate that the endocuticle is probably less brittle than the exocuticle, most likely due to the alignment of the $\alpha$-chitin fibers in each macrofiber.
+
+			In addition, the exocuticle typically appears to fracture before the endocuticle, with shear-cusp formation evident over unbroken endocuticle.
+			We note, however, that the exocuticle of weevils/beetles is anchored to the endocuticle by cross-linking fibers in a transition zone described by Kamp et al. (see refs.).
+			The presence of shear-cusps therefore indicates that the fibers of the endocuticle are shearing past each other within each ply along the radial/normal plane (type II shearing???).
+			Furthermore, given the delamination observed between plies, the layers of endocuticle are liekly shearing past each other (type III shearing) along the transverse plane.
+			We therefore infer that, under tension, the endocuticle tends to deform visco-elastically and plastically along the longitudinal axes of the macrofibers, while the overlying exocuticle exhibits brittle fracture due to shearing between the stretching endocuticle fibers to which it is anchored.
+
+			Additionally, the fracture surfaces show a characteristic failure mode, based on the pattern of fiber dislocation in the plies of the endocuticle.
+			(copy description from DATA PRESENTATION, for figure 5B).
+
+			From this pattern we hypothesize that the exocuticle-rich gular sutures are the most likely site for the initiation of void nucleation and failure of the integrated rostral cuticle in cross section.
+			Structure failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
+			We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrallyl
+			Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.
 
 	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
+		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
+		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.
 
-	\section{Conclusion} %shoot for 500 words
+		~400K cycles, complete elongation, rehydrated, coated in grease to prevent loss of moisture and stiffening of the specimen.
+		We observed visco-elastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
+		Fmax decreased logarithmically with cycle number, etc., and the specimen appeared to have deformed plastically during the test.
+		We initially believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that the specimen returned to its original shape.
+		The specimen did not show any evidence of fractures or shear cusps anywhere in the surface of the exocuticle, and, furthermore, the tensile strength of the specimen was consistent with other members of its species.
+		Given this surprising result, it appears that the specimen was undamaged by the testing.
+		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/visco-elastic.
+
+		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is resposible for what we observed.
+		The endocuticle is made of aligned $\alpha$-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
+		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
+		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform visco-elastically/visco-plastically.
+		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain without any damage to the specimen.
+
+	\section{Conclusions} %shoot for 500 words
 

### 2019-01-17 tex/manuscript.tex
critique
needs new narrative direction

@@
 	%title and author block
-	{\title{Microstructure-derived strength in the acorn weevil exoskeleton}
+	{\title{Microstructure and mechanical behavior of the acorn weevil exoskeleton}
 
@@
 		Beetles in particular have a highly modified endocuticle comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
+		The macrofibers are strongly anisotropic, with an elastic modulus of ~8-9GPa along their longitudinal axis, but only 1-2 GPa along the transverse plane.
 		Macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
@@
 		\end{enumerate}
+
+		We believe that each of these modifications contributes to the observed flexibility of the snout, but in different ways.
+		Because the exocuticle is generally more rigid (and brittle) than the endocuticle due to its transverse isotropy and microstructure, a reduction in the relative thickness of this region would likely have the effect of increasing the flexibility of the resulting composite.
+		Additionally, aligning the layers of endocuticle as a 45 deg cross ply would likely reduce the bending moment of the structure.
+		The combined effect of these modifications is that the apex of the rostrum is able to bend until completely straight without fracture.
 
@@
 			\begin{enumerate}
-				\item There is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology. 
+				\item There is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology. (singh 2016, jansen 2016)
 				\item The endocuticle contributes more to the tensile strength of the rostrum than the exocuticle, possibly because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
@@
 
-			Finally, we observed that Young's modulus of elasticity is higher in specimens with a longer snout.
+			Finally, we observed that the elastic modulus is higher in specimens with a longer snout.
 			This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) would need to be more flexible to avoid fracture during oviposition.
@@
 			From this pattern we hypothesize that the exocuticle-rich gular sutures are the most likely site for the initiation of void nucleation and failure of the integrated rostral cuticle in cross section.
-			Structure failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
-			We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrallyl
+			Structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
+			We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally
 			Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.
@@
 	\section{Conclusions} %shoot for 500 words
+			Overall conclusion - the only difference between the flexible part and the stiff part is the microstructure; there is no difference in the mechanical behavior of the endocuticle at different locations on the snout and across species.
 

### 2019-01-20 tex_v2/cover_letter.tex
rewrite and restructure narrative

\documentclass[10pt,letterpaper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[letterpaper,left=1in, right=1in, top=1in, bottom=1in]{geometry}

\makeatletter
\let\@texttop\relax
\makeatother

\address{Andrew Jansen\\
Life Sciences A-Wing\\
Arizona State University\\
427 E. Tyler Mall\\
PO Box 874501\\
Tempe, AZ 85287-4501} 
\signature{Andrew Jansen\\
Corresponding Author – majanse1@asu.edu\\
Graduate Student\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA} 
\begin{document} 
\begin{letter}{NAME\\
Editor\\
email@address\\
address 1\\
address 2\\
address 3}
\opening{Dear. !!!EDITOR-NAME-HERE!!!,}

I am writing to submit our manuscript entitled, ``Microstructure-derived strength in the acorn weevil exoskeleton'', for consideration for publication in Nature Materials.
This paper DOES THIS NEW IMPACTFUL THING.

We WHAT WE DO HEREIN.
This work represents WHAT THIS WORK SIGNIFIES.

Our manuscript therefore serves two distinct, but equally important ends: (1) PURPOSE 1, and (2) PURPOSE 2.
All of the authors have agreed to submit the manuscript to Nature Materials; we feel that this journal will permit us to reach the ideal readership to disseminate our findings and to promote these methods.

We suggest the following individuals be contacted as potential reviewers: Dr. ONE (EMAIL), for WHAT REASON, as well as Dr. TWO (EMAIL) and Dr. THREE (email), both of whom are REASON.

This paper is not being considered for publication at any other journal.
We have carefully followed your manuscript preparation guidelines, and look forward to your comments.
If this manuscript is not suitable to your needs, please notify us as soon as possible.
We thank you for your time and consideration.

\closing{Sincerely,}

\newpage

\cc{Jason Williams\\
Co-author\\
Research Associate\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nikhilesh Chawla\\
Co-author\\
Professor\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nico Franz\\
Co-author\\
Professor\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA} 
%\ps{adding a postscript} 
%\encl{list of enclosed material} 
\end{letter} 
\end{document}

### 2019-01-20 tex_v2/manuscript.tex
now the narrative makes sense!!

--- 2019-01-12 tex/manuscript.tex
@@
 	%title and author block
-	{\title{Microstructure-derived strength in the acorn weevil exoskeleton}
+	{\title{Exoskeletal microstructure and tensile behavior of the acorn weevil rostrum}
 
@@
 
-%introduction - nature materials requires that introduction is given NO SECTION HEADER
-%Brief explanation of relevant behavior and biology, prior work on ecology, head shape, cuticle mechanics, and rostral microstructure.
+%manuscript outline:
 
-%~650 words here
+%standard intro, talk about behavior, rostrum bending, and cuticle
+%weevil cuticle is known to be stronger in compression than in tension
+%in Curculio this means that the greatest risk of breakage would come from tension during bending of the snout
+%curculio appear to have mitigated the risk of breakage by making the rostral apex flexible
+%prior work suggests that this flexibility comes from a modified composite profile in the cuticle of the rostral apex
+%different parts of the head thus appear to have differently specialized cuticle to deal with extreme bending
+%never any comparison across species, fractographic analysis never attempted to explain how snout is most likely to break
 
-	We report novel modifications to the composite microstructure of the exoskeleton in the snout of acorn weevils (Coleoptera: Curculionidae) belonging to the genus \textit{Curculio} Linnaeus, 1756.
+%we therefore set out to characterize the modifications to the cuticle of the rostral apex that presumably allow bending.
+%because material rigidity and toughness are often inversely proportional, we also investigate how the composite structure of the cuticle contributes to the tensile strength of the snout, whether the flexible regions exhibit higher tensile strenth, and conduct fractographic analysis on tensile testing specimens to determine the failure modes of the rostrum under tension.
+%We additionally present the results of a load cycling test for a representative specimen of a species with a highly elongate snout to explore the possibility that bending of the rostrum results in elastic deformation of the cuticle.
 
-	As a weevil (snout beetle), members of the genus \textit{Curculio} are typified by the presence of a highly elongate structure on the head, called the rostrum (snout). 
-	This structure is a hollow, strongly curved (over 90$\degree$), cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
-	The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
-	By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \cref{sec:microstructure}).
-	Acorn weevils use this structure to excavate sites for egg-laying (oviposition) and feeding on a variety of fruits, including acorns, Japanese camellia, hazelnuts, pecans, chestnuts, and chinquapins.
+%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich gular sutures, which would be under tension during the drilling behavior in the live insect.
 
-	During oviposition, a female engages in a unique "drilling" behavior that causes significant, apparently elastic, deformation of the rostrum.
-	The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the perimeter of the bore-hole.
-	Once the apex of the snout is fully inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
-	The female will maintain tension on the rostrum in this position, continuing to ingest the substrate and rotate around the bore-hole, while slowly inserting the rostrum further into the excavated channel.
-	Once the rostrum is fully inserted, usually up to the eyes, she will pull her snout from the bore-hole and deposit several eggs into the site.
-	By maintaining constant tension on the rostrum and rotating around the bore-hole, the female is able to flex the snout into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
-	While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending incurred during the process of oviposition.
+%this has important implications for understanding the evolution of this highly exaggerated structure, ie suggesting negative selection of breakage instead of positive selection of flexibility.
+possibly a single ancestral shift for each modification, opened a large region of morphospace for exploration, resulting in present day diversity of the genus.
 
-	We have found that the composite profile of the rostrum is strongly differentiated from the head capsule and other body parts, with modification of both the relative layer thicknesses and fiber orientation angles of cuticle regions (viz.
-exocuticle and endocuticle), which we describe in detail below.
-	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
-	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
+%-------------------------------------------------------------------------------------------------
+%introduction - nature materials requires that introduction is given NO SECTION HEADER, 500 words
 
-	Support for this hypothesis has come from three lines of evidence:
+	\section{Microstructure of the \textit{Curculio} rostrum}\label{sec:microstructure}
 
-	\begin{enumerate}
-	\item Examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species.
-	\item Tensile testing of the rostrum has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species.
-	\item Fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
-	\end{enumerate}
+	\section{Force-controlled loading to fracture}
 
-	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
-	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.
+	\section{Load cycling of \textit{Curculio caryae}}
 
-	MIGHT BE GOOD TO SIGNPOST HERE...
-we discuss each of these lines of evidence in the sections below beginning with a consideration of the impact of microstructure, then we talk about mechanical testing, blah blah.
-Emphasize that this is key to predicting and understanding the mechanical behavior of the snout during bending.
+	\section{Fractography of test specimens}
 
-%stick to 600 words for each results section and 500 words for conclusions
-	\section{Microstructure of the rostrum}\label{sec:microstructure}
-	%~700 words here (1500 total, shorten later)
-		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
-		The arrangement of the embedded chitin fibers is primarily responsible for the mechanical behavior of the cuticle, although this is modulated by the degree of tanning and water retention.
-\cite{Klocke2011,Vincent2004}.
-		Accordingly, cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.
-
-		Insect cuticle is divisible into three regions (Figs.~INSERT-REFS-HERE), including (1) endocuticle, which is the most compliant and innermost region; and (3) exocuticle, which is the stiffest, hardest, and outermost region.
-		Between these lies (2) mesocuticle, which is similar in microstructure to exocuticle, but less sclerotized (tanned), usually acting as a thin transition zone \cite{Klocke2011,Vincent1982,Vincent2004}.
-
-		In general, these regions are confluent, and not necessarily sharply defined; for instance, mesocuticle is not always evident, and is sometimes considered part of the exocuticle.
-		In beetles, both exocuticle and mesocuticle (when present) are laminate, and have numerous laminae of unidirectional chitin fibers, each layer a single fiber thick (2-4nm), embedded in a proteinaceous matrix.
-		These layers are stacked at a more or less constant angle to each other in a helicoidal arrangement, referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
-		This microstructure allows exocuticle and mesocuticle to exhibit transverse isotropy, or complete isotropy in some cases, despite the strong anisotropy of $\alpha$-chitin.
-
-		Conversely, the endocuticle of Coleoptera appears to take advantage of this anisotropy to improve fracture resistance of the exoskeleton.
-		Beetles in particular have a highly modified endocuticle comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
-		Macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
-		Typically, endocuticle contains several layers of macrofibers, with the adjacent layers forming pairs. 
-		The layers \emph{within} each pair are pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other, see Figs.~INSERT-REFS-HERE), while the stacking angle \emph{between} pairs is typically acute, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
-		It is thought that this arrangement of the macrofibers contributes to the toughness of beetle exoskeletons by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
-
-		We have found that in \textit{Curculio}, the head capsule (which is similar to the rest of the body) fits the general profile of endocuticle, with an angle of approximately $30\degree$ between successive pseudo-orthogonal plies.
-		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle; compared to the snout, this part of the head is fairly rigid.
-
-		By contrast, the region beyond the scrobe (antennal channel) is quite flexible, even in fully desiccated specimens.
-		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has a different composite structure (see Fig~INSERT-REFS-HERE) than the head capsule, namely:
-
-		\begin{enumerate}
-			\item The exocuticle is reduced to a thin shell, with the endocuticle thickened to offset this reduction.
-			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approx. $\pm45\degree$ to the longitudinal axis of the snout.
-		\end{enumerate}
-
-		Correspondingly, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.		
-		In previous work we identified these modifications to the composite structure of the cuticle, but only within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
-		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (listed below), indicating that this is likely a genus-wide trait.
-
-		Notably, we could find no evidence of resilin, as indicated by florescent microscopy, anywhere in the cuticle of the head (including the rostrum).
-		We also observed no difference in total cuticle thickness between the head capsule and rostral apex, only differences in relative layer thicknesses.
-		Below we demonstrate that that the endocuticle does not vary in tensile strength across the rostrum, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
-		The available evidence therefore suggests that the relative flexibility of the snout is solely derived from the composite profile of its cuticle.
-
-	\section{Tensile testing and fracture mechanics} %shoot for 600 words
-		Based on our understanding of the microstructure of the cuticle, it is possible to make several predictions about the mechanical behavior of the snout.
-		First, we can predict that the endocuticle will contribute more to the tensile strength of the rostrum than the exocuticle, because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
-		Second, the endocuticle will be less brittle than the exocuticle, therefore we should expect that fractures will initiate in the exocuticle.
-		The predicted behavior of the snout is borne out by the data, where UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
-		In addition to making the cuticle flexible, this profile makes it less likely to break, suggesting a possible means by which the system evolved (negative selection of breakage, maybe say this in conclusions).
-
-		Outline:
-		Intro
-			To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the heads of 6 species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
-		Tensile testing
-			We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
-			Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
-			 All of these associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
-
-			This strongly suggests that there is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology.
-
-	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
-
-	\section{Conclusion} %shoot for 500 words
+	\section{Conclusions} %shoot for 500 words
 
@@
 		\section*{References}	
-			\bibitem{Klocke2011}
-				Klocke, D. \& Schmitz, H.
-				Water as a major modulator of the mechanical properties of insect cuticle.
-				\textit{Acta Biomater.}
-				\textbf{7,}
-				2935--2942
-				(2011).
-
-			\bibitem{Vincent1982}
-				Vincent, J. F. V.
-				\textit{Structural biomaterials}
-				(Halsted Press,
-				New York,
-				1982).
-
-			\bibitem{Vincent2004}
-				Vincent, J. F. V. \& Wegst, U. G. K.
-				Design and mechanical properties of insect cuticle.				
-				\textit{Arthropod Struct. Dev.}
-				\textbf{33:3,}
-				187--199,
-				(2004).
-
-			\bibitem{Nikolov2011}
-				Nikolov, S. et al.
-				Robustness and optimal use of design principles of arthropod exoskeletons studied by ab initio-based multiscale simulations.
-				\textit{J. Mech. Behav. Biomed. Mater.}
-				\textbf{4:2,}
-				129--145,
-				(2011).
-
-			\bibitem{Nikolov2010}
-				Nikolov, S. et al.
-				Revealing the design principles of high-performance biological composites using Ab initio and multiscale simulations: The example of lobster cuticle.
-				\textit{Adv. Mater.}
-				\textbf{22:4,}
-				519--526,
-				(2010).
-
-			\bibitem{Blackwell1980}
-				Blackwell, J. \& Weih, M.
-				Structure of chitin-protein complexes: ovipositor of the ichneumon fly \textit{Megarhyssa}.
-				\textit{J. Mol. Biol.}
-				\textbf{137:1,}
-				49--60,
-				(1980).
-
-			\bibitem{Bouligand1972}
-				Bouligand, Y.
-				Twisted fibrous arrangements in biological materials and cholesteric mesophases.
-				\textit{Tissue Cell}
-				\textbf{4:2,}
-				189--217,
-				(1972).
-
-			\bibitem{Neville1976}
-				Neville, A. C., Parry, D. A. \& Woodhead-Galloway, J.
-				The chitin crystallite in arthropod cuticle.
-				\textit{J. Cell Sci.}
-				\textbf{21:1,}
-				73--82,
-				(1976).
-
-			\bibitem{Cheng2009}
-				Cheng, L., Wang, L., \& Karlsson, A. M.
-				Mechanics-based analysis of selected features of the exoskeletal microstructure of \textit{Popillia japonica}.
-				\textit{Mater. Res.}
-				\textbf{24:11,}
-				3253--3267,
-				(2009).
-
-			\bibitem{Hepburn1973}
-				Hepburn, H. R. \& Ball, A.
-				On the structure and mechanical properties of beetle shells.
-				\textit{J. Mater. Sci.}
-				\textbf{8:5,}
-				618--623,
-				(1973).
-
-			\bibitem{Kamp2010}
-				van de Kamp, T. \& Greven, H.
-				On the architecture of beetle elytra.
-				\textit{Entomol. Heute}
-				\textbf{22,}
-				191--204,
-				(2010).
-
-			\bibitem{Kamp2015}
-				van de Kamp, T., Riedel, A. \& Greven, H.
-				Micromorphology of the elytral cuticle of beetles, with an emphasis on weevils (Coleoptera : Curculionoidea)
-				\textit{Arthropod Struct. Dev.}
-				\textbf{45:1,}
-				14--22,
-				(2015).
-
+			\bibitem{lamport94}
+				Leslie Lamport,
+				\textit{\LaTeX: a document preparation system},
+				Addison Wesley, Massachusetts,
+				2nd edition,
+				1994.
 	\end{thebibliography}

### 2019-01-22 tex_v2/manuscript.tex
intro 1st pass

--- 2019-01-17 tex/manuscript.tex
@@
 
+%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----_-----------------------------------
+\usepackage{xargs}
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+\newcommandx{\info}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{#2}}
+\newcommandx{\improvement}[2][1=]{\todo[linecolor=purple,backgroundcolor=purple!25,bordercolor=purple,#1]{#2}}
+\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
+%-------------------------------------------------------------------------------
+
 \begin{document}
@@
 	%title and author block
-	{\title{Microstructure and mechanical behavior of the acorn weevil exoskeleton}
-
+	{\title{Exoskeletal microstructure and tensile behavior of the acorn weevil rostrum}
+	%CONSIDER CHANGING THE TITLE AFTER WRITING MANUSCRIPT
 	\date{\today}
@@
 
-%introduction - nature materials requires that introduction is given NO SECTION HEADER
-%Brief explanation of relevant behavior and biology, prior work on ecology, head shape, cuticle mechanics, and rostral microstructure.
+{%manuscript outline:
+%3000 word limit, 150 word abstract, 500 word intro, 250 word fig caps
 
-%~650 words here
+%standard intro, talk about behavior, rostrum bending, and cuticle
+%weevil cuticle is known to be stronger in compression than in tension
+%in Curculio this means that the greatest risk of breakage would come from tension during bending of the snout
+%curculio appear to have mitigated the risk of breakage by making the rostral apex flexible
+%prior work suggests that this flexibility comes from a modified composite profile in the cuticle of the rostral apex
+%different parts of the head thus appear to have differently specialized cuticle to deal with extreme bending
+%only observed in 1 species, never any comparison across species/in genus, fractographic analysis never attempted to explain how snout is most likely to break/ how the cuticle fails (ie what failure modes are the weevils fighting against/trying to mitigate)
 
-	We report novel modifications to the composite microstructure of the exoskeleton in the snout of acorn weevils (Coleoptera: Curculionidae) belonging to the genus \textit{Curculio} Linnaeus, 1756.
+%we therefore set out to characterize the modifications to the cuticle of the rostral apex that presumably allow bending, using six species representing a mix of phylogenetically related and disparate species.
+%because material rigidity and toughness are often inversely proportional, we also investigate how the composite structure of the cuticle contributes to the tensile strength of the snout, whether the flexible regions exhibit higher tensile strenth, and conduct fractographic analysis on tensile testing specimens to determine the failure modes of the rostrum/cuticle under tension.
+%We additionally present the results of a load cycling test for a representative specimen of a species with a highly elongate and curved snout to explore the possibility that bending of the rostrum results in elastic deformation of the cuticle.
 
-	As a weevil (snout beetle), members of the genus \textit{Curculio} are typified by the presence of a highly elongate structure on the head, called the rostrum (snout). 
-	This structure is a hollow, strongly curved (over 90$\degree$), cylindrical extension of the exoskeleton of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
-	The space inside of the rostrum contains the esophagus, various muscles and tendons used for feeding, and hemolymph that serves as a rough equivalent to blood in insects.
-	By contrast, the solid shell of the rostrum is comprised entirely of cuticle, which can be considered a laminate composite consisting of various arrangements of chitin fibers embedded in a protein matrix (see \cref{sec:microstructure}).
-	Acorn weevils use this structure to excavate sites for egg-laying (oviposition) and feeding on a variety of fruits, including acorns, Japanese camellia, hazelnuts, pecans, chestnuts, and chinquapins.
+%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, present in all examined species, and likely across the genus (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich gular sutures, which would be under tension during the drilling behavior in the live insect.
 
-	During oviposition, a female engages in a unique "drilling" behavior that causes significant, apparently elastic, deformation of the rostrum.
-	The female will insert the snout into an incision made with the mandibles, eating the material as she proceeds, while rotating her head and body around the perimeter of the bore-hole.
-	Once the apex of the snout is fully inserted, she will push up and forward with her front legs, forcing the rostrum to bend until it is nearly straight.
-	The female will maintain tension on the rostrum in this position, continuing to ingest the substrate and rotate around the bore-hole, while slowly inserting the rostrum further into the excavated channel.
-	Once the rostrum is fully inserted, usually up to the eyes, she will pull her snout from the bore-hole and deposit several eggs into the site.
-	By maintaining constant tension on the rostrum and rotating around the bore-hole, the female is able to flex the snout into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
-	While this behavior has been observed in many species of \textit{Curculio}, we have lacked a fundamental understanding of how female \textit{Curculio} rostra can withstand the repeated, often extreme bending incurred during the process of oviposition.
+%in conclusion, the cuticle of the rostral apex is modified to increase flexibility, allowing for nearly elastic bending, while simultaneously increasing tensile strength, thus resisting the predominant fracture modes observed in unaltered cuticle.
+%to our knowledge, this is the first time that modifications to the composite profile of the cuticle have been observed to optimized mechanical behavior in insects, or something like that.
 
-	We have found that the composite profile of the rostrum is strongly differentiated from the head capsule and other body parts, with modification of both the relative layer thicknesses and fiber orientation angles of cuticle regions (viz.
-exocuticle and endocuticle), which we describe in detail below.
-	We posit that these modifications enable the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
-	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
+%this has important implications for understanding the evolution of this highly exaggerated structure, ie suggesting negative selection of breakage instead of positive selection of flexibility.
+possibly a single ancestral shift for each modification, opened a large region of morphospace for exploration, resulting in present day diversity of the genus.
 
-	Support for this hypothesis has come from three lines of evidence:
+%-------------------------------------------------------------------------------
+%introduction - nature materials requires that introduction is given NO SECTION HEADER, 500 words
+}	
+\section*{Introduction}
+%Intro outline - 500 word maximum
+%	Paragraph 1 - Cuticle of beetles
+%	Paragraph 2 - Weevils and behavior
+%	Paragraph 3 - modified cuticle in curculio
+%	Paragraph 4 - study objectives
+%	Paragraph 5 - short segue/ signposting
+	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite typified by variously arranged $\alpha$-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
+	Although $\alpha$-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
+	Beetle cuticle is divided into two structurally and mechanically distinct regions, the (outer) exocuticle and (inner) endocuticle.
+	The exocuticle is characterized by a transversely-isotropic helicoidal structure that is comparatively rigid (though brittle) and has been the subject of many prior studies\unsure{numbers and refs}.
+	By contrast, the endocuticle of beetles is comprised of large, aligned bundles (macrofibers) of chitin that are strongly anisotropic\unsure{numbers again}.
+	Arranged in unidirectional laminae, the macrofibers contribute to the toughness of the exoskeleton by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
 
-	\begin{enumerate}
-	\item Examination of the cuticle microstructure across the length of the snout has revealed consistent modification to the composite structure of the rostrum among \textit{Curculio} species.
-	\item Tensile testing of the rostrum has demonstrated that the mechanical strength of the cuticle components are consistent along the length of the structure and between species.
-	\item Fatigue testing has shown that a highly curved rostrum is capable of flexing hundreds of thousands of times without damage to the structure, and is apparently elastic.
-	\end{enumerate}
+	In general, impact-prone areas and exaggerated structures, such as horns and legs exhibit cuticle organization that resists deformation and fracture.
+	Acorn weevils in the genus \textit{Curculio} are typified by an elongate structure of the head, called the rostrum (snout), which instead exhibits unusual distal flexibility.
+	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
+	Despite being composed of the same material as other rigid body parts, the snout can be repeatedly bent without evident damage.
+	This structure is used by the female to feed and excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
+	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
+
+	While this behavior has been observed in many species of \textit{Curculio}, it was unclear how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
+	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
+	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts.
+	Because recent studies have shown that the yield strength of the weevil rostrum exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
+	We relate an observed increase in the volume fraction of endocuticle in the rostrum to higher tensile strength at the rostral apex in all tested species, and find that a strongly curved rostrum can be flexed repeatedly without harm to the structure.
 
 	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
-	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.
-
-	MIGHT BE GOOD TO SIGNPOST HERE...
-we discuss each of these lines of evidence in the sections below beginning with a consideration of the impact of microstructure, then we talk about mechanical testing, blah blah.
-Emphasize that this is key to predicting and understanding the mechanical behavior of the snout during bending.
-
-%stick to 600 words for each results section and 500 words for conclusions
-	\section{Microstructure of the rostrum}\label{sec:microstructure}
-	%~700 words here (1500 total, shorten later)
+	Based on our findings we posit that the composite profile of the rostral apex enables the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
+	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.\info{630 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
+}	
+	\section{Microstructure of the \textit{Curculio} rostrum}\label{sec:microstructure}
 		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
@@
 
-	\section{Tensile testing of the rostrum} %shoot for 600 words
+	\section{Force-controlled loading to fracture}
 		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
@@
 
-		\subsection{Tensile behavior}
 		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
@@
 
-		\subsection{Fractography}
+	\section{Load cycling of \textit{Curculio caryae}}
+		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
+		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.
+
+		~400K cycles, complete elongation, rehydrated, coated in grease to prevent loss of moisture and stiffening of the specimen.
+		We observed visco-elastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
+		Fmax decreased logarithmically with cycle number, etc., and the specimen appeared to have deformed plastically during the test.
+		We initially believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that the specimen returned to its original shape.
+		The specimen did not show any evidence of fractures or shear cusps anywhere in the surface of the exocuticle, and, furthermore, the tensile strength of the specimen was consistent with other members of its species.
+		Given this surprising result, it appears that the specimen was undamaged by the testing.
+		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/visco-elastic.
+
+		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is resposible for what we observed.
+		The endocuticle is made of aligned $\alpha$-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
+		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
+		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform visco-elastically/visco-plastically.
+		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain without any damage to the specimen.
+
+	\section{Fractography of test specimens}
 		Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).
@@
 		Structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
-			We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally
+		We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally.
 		Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.
 
-	\section{Fatigue testing of \textit{Curculio caryae}} %shot for 600 words
-		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
-		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.
+	\section{Conclusions}
+	%1-2 'short paragraphs'
 
-		~400K cycles, complete elongation, rehydrated, coated in grease to prevent loss of moisture and stiffening of the specimen.
-		We observed visco-elastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
-		Fmax decreased logarithmically with cycle number, etc., and the specimen appeared to have deformed plastically during the test.
-		We initially believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that the specimen returned to its original shape.
-		The specimen did not show any evidence of fractures or shear cusps anywhere in the surface of the exocuticle, and, furthermore, the tensile strength of the specimen was consistent with other members of its species.
-		Given this surprising result, it appears that the specimen was undamaged by the testing.
-		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/visco-elastic.
-
-		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is resposible for what we observed.
-		The endocuticle is made of aligned $\alpha$-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
-		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
-		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform visco-elastically/visco-plastically.
-		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain without any damage to the specimen.
-
-	\section{Conclusions} %shoot for 500 words
-			Overall conclusion - the only difference between the flexible part and the stiff part is the microstructure; there is no difference in the mechanical behavior of the endocuticle at different locations on the snout and across species.
-
+	\section{Methods}
 	%actual methods after references
-	\section{Methods}
 		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.
 
+	\begin{thebibliography}{50}
 	%50 references (as a guideline, in order of appearance in text)
-	\begin{thebibliography}{50}
 		\section*{References}	

### 2019-01-28 tex_v2/manuscript.tex
more modeling

@@
 
+\usepackage{amsmath}
 \usepackage{graphicx}
+%TODO remove package: blindtext
 \usepackage{blindtext}
@@
 
-%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----_-----------------------------------
+%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----------------------------------------
 \usepackage{xargs}
@@
 \newcommandx{\improvement}[2][1=]{\todo[linecolor=purple,backgroundcolor=purple!25,bordercolor=purple,#1]{#2}}
-\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
+\newcommandx{\noshow}[2][1=]{\todo[disable,#1]{#2}}
 %-------------------------------------------------------------------------------
@@
 	By contrast, the endocuticle of beetles is comprised of large, aligned bundles (macrofibers) of chitin that are strongly anisotropic\unsure{numbers again}.
-	Arranged in unidirectional laminae, the macrofibers contribute to the toughness of the exoskeleton by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
+	Arranged in unidirectional laminae, the macrofibers improve cuticle toughness by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
 
@@
 	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
-	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts.
+	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we predict the effect of these differences using Classical Laminate Theory (CLT).
 	Because recent studies have shown that the yield strength of the weevil rostrum exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
@@
 }	
-	\section{Microstructure of the \textit{Curculio} rostrum}\label{sec:microstructure}
-		Insect cuticle is made up of chitin and numerous uncharacterized proteins, which effectively act as a matrix into which the chitin is embedded \cite{Nikolov2011,Nikolov2010,Vincent2004}.
-		The arrangement of the embedded chitin fibers is primarily responsible for the mechanical behavior of the cuticle, although this is modulated by the degree of tanning and water retention.
-\cite{Klocke2011,Vincent2004}.
-		Accordingly, cuticle is often considered as a fiber-reinforced composite material; however, in beetles, the composite arrangement of the chitin fibers varies by cuticle region.
+	\section{Microstructure of the \textit{Curculio} rostrum}
+		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4nm) embedded in a proteinaceous matrix.
+		These layers are stacked at a more or less constant angle to each other, thus forming a quasi-isotropic laminate referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}.
+		This layout effectively produces a transversely isotropic composite, mitigating the strong anisotropy of $\alpha$-chitin to yield a versatile building material for the exoskeleton.
 
-		Insect cuticle is divisible into three regions (Figs.~INSERT-REFS-HERE), including (1) endocuticle, which is the most compliant and innermost region; and (3) exocuticle, which is the stiffest, hardest, and outermost region.
-		Between these lies (2) mesocuticle, which is similar in microstructure to exocuticle, but less sclerotized (tanned), usually acting as a thin transition zone \cite{Klocke2011,Vincent1982,Vincent2004}.
+		Conversely, the endocuticle of Coleoptera retains the anisotropy of $\alpha$-chitin to improve fracture resistance of the cuticle.
+		Beetle endocuticle is unique among arthropods and is comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
+		Chitin macrofibers are orthotropic \improvement{insert E1, E2, E3 here} and arranged in unidirectional plies, as in Figs. \info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
+		Typically, adjacent macrofiber plies are paired and pseudo-orthogonal (i.e., angled approx.
+$90\degree$ to each other, see Figs.\info{~INSERT-REFS-HERE}), with a constant stacking angle \emph{between} pairs \info{what kind of laminate is this?}, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
+		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure.
+		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010,Kamp2015,Hepburn1973}.
 
-		In general, these regions are confluent, and not necessarily sharply defined; for instance, mesocuticle is not always evident, and is sometimes considered part of the exocuticle.
-		In beetles, both exocuticle and mesocuticle (when present) are laminate, and have numerous laminae of unidirectional chitin fibers, each layer a single fiber thick (2-4nm), embedded in a proteinaceous matrix.
-		These layers are stacked at a more or less constant angle to each other in a helicoidal arrangement, referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
-		This microstructure allows exocuticle and mesocuticle to exhibit transverse isotropy, or complete isotropy in some cases, despite the strong anisotropy of $\alpha$-chitin.
-
-		Conversely, the endocuticle of Coleoptera appears to take advantage of this anisotropy to improve fracture resistance of the exoskeleton.
-		Beetles in particular have a highly modified endocuticle comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
-		The macrofibers are strongly anisotropic, with an elastic modulus of ~8-9GPa along their longitudinal axis, but only 1-2 GPa along the transverse plane.
-		Macrofibers are aligned in layers, as in Figs~INSERT-FIGS-HERE \cite{Kamp2010,Kamp2015}.
-		Typically, endocuticle contains several layers of macrofibers, with the adjacent layers forming pairs. 
-		The layers \emph{within} each pair are pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other, see Figs.~INSERT-REFS-HERE), while the stacking angle \emph{between} pairs is typically acute, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
-		It is thought that this arrangement of the macrofibers contributes to the toughness of beetle exoskeletons by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
-
-		We have found that in \textit{Curculio}, the head capsule (which is similar to the rest of the body) fits the general profile of endocuticle, with an angle of approximately $30\degree$ between successive pseudo-orthogonal plies.
-		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle; compared to the snout, this part of the head is fairly rigid.
-
-		By contrast, the region beyond the scrobe (antennal channel) is quite flexible, even in fully desiccated specimens.
-		Serial thin sectioning of the snout has demonstrated that the cuticle in this region has a different composite structure (see Fig~INSERT-REFS-HERE) than the head capsule, namely:
+		Serial thin sectioning and SEM of fractured \textit{Curculio} specimens has revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
+		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle\info{thickness measurements here}.
+		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}E) in two key characteristics:
 
 		\begin{enumerate}
-			\item The exocuticle is reduced to a thin shell, with the endocuticle thickened to offset this reduction.
-			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approx. $\pm45\degree$ to the longitudinal axis of the snout.
+			\item The exocuticle is reduced to a thin shell\info{measurement here}, with the endocuticle thickened \info{measurement here} to offset this reduction to maintain a constant cuticle thickness in the head.
+			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
 		\end{enumerate}
 
-		We believe that each of these modifications contributes to the observed flexibility of the snout, but in different ways.
-		Because the exocuticle is generally more rigid (and brittle) than the endocuticle due to its transverse isotropy and microstructure, a reduction in the relative thickness of this region would likely have the effect of increasing the flexibility of the resulting composite.
+		In previous work we identified these modifications to the composite structure of the cuticle within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
+		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (listed below), indicating that this is likely a genus-wide trait.
+		We note that in all species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.
+
+		To investigate the effect of these cuticle modifications, we estimated the engineering and flexural elastic moduli of the cuticle in both the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
+		We previously derived the effective elastic constants of the cuticle regions of \textit{C. longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
+		The cuticle of the head capsule has moduli\info{insert numbers here}.
+		The cuticle of the rostral apex has moduli\info{insert numbers here}.
+
+		Interestingly, each of these modifications contributes to the observed flexibility of the snout.
+
+		\info{some explanation in source code}
+		\noshow{Because the exocuticle is generally more rigid (and brittle) than the endocuticle due to its transverse isotropy and microstructure, a reduction in the relative thickness of this region would likely have the effect of increasing the flexibility of the resulting composite.
 		Additionally, aligning the layers of endocuticle as a 45 deg cross ply would likely reduce the bending moment of the structure.
-		The combined effect of these modifications is that the apex of the rostrum is able to bend until completely straight without fracture.
+		The combined effect of these modifications is that the apex of the rostrum is able to bend until completely straight without fracture.}
 
-		Correspondingly, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.		
-		In previous work we identified these modifications to the composite structure of the cuticle, but only within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
-		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (listed below), indicating that this is likely a genus-wide trait.
-
-		Notably, we could find no evidence of resilin, as indicated by florescent microscopy, anywhere in the cuticle of the head (including the rostrum).
-		We also observed no difference in total cuticle thickness between the head capsule and rostral apex, only differences in relative layer thicknesses.
-		Below we demonstrate that that the endocuticle does not vary in tensile strength across the rostrum, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
-		The available evidence therefore suggests that the relative flexibility of the snout is solely derived from the composite profile of its cuticle.
+		Below we demonstrate that the endocuticle does not vary in tensile strength across the rostrum or between species, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
+		The available evidence therefore suggests that the relative flexibility of the snout is primarily derived from the composite profile of its cuticle.
 
@@
 		We note, however, that the exocuticle of weevils/beetles is anchored to the endocuticle by cross-linking fibers in a transition zone described by Kamp et al. (see refs.).
-		The presence of shear-cusps therefore indicates that the fibers of the endocuticle are shearing past each other within each ply along the radial/normal plane (type II shearing???).
+		The presence of shear-cusps therefore indicates that the fibers of the endocuticle are shearing past each other within each ply along the radial/normal plane (type II shearing???, extension shear-coupling).
 		Furthermore, given the delamination observed between plies, the layers of endocuticle are liekly shearing past each other (type III shearing) along the transverse plane.
@@
 
+		\subsection*{Constitutive modeling of the cuticle}
+			\paragraph*{General Approach}
+			\paragraph*{Cuticle profile of model}
+				The cuticle of \textit{C. longinasus} is 50 microns thick; we use \textit{C. longinasus} because we have constitutive models for the cuticle regions of this species.
+				We assumed equal layer thicknesses in the endocuticle of the basal cuticle, and equal thicknesses of exocuticle and endocuticle.
+				In the apical cuticle we assumed the exocuticle and upper 8 layers of the endocuticle were each 5 microns thick, with 4 thinner layers of endocuticle, each 1.5 microns thick.
+				The angles of the endocuticle layers for both types of cuticle are as described above.
+			\paragraph*{Permutations}
+				We additionally calculated models for two hypothetical hybrid cuticles: one model has the layer thicknesses of the apex, but fiber orientations of the base, while the second has the fiber orientations of the apex, but the layer thicknesses of the base.
+			\paragraph*{Classical Laminate Plate Theory}
+				We begin by calculating the 2D reduced stiffness matrix for each part of the cuticle.
+				For orthotropic materials with the principal axes parallel to the ply edges, the reduced stiffness matrix is defined as follows:
+
+				\begin{equation}
+				[Q] =
+					\begin{bmatrix}
+						Q_{11} & Q_{12} & 0 \\
+						Q_{21} & Q_{22} & 0 \\
+						0 & 0 & Q_{66}
+					\end{bmatrix}\,,
+				\end{equation}
+
+				and where:
+
+				\begin{equation}
+				\begin{aligned}
+					Q_{11} & = \frac{E_{1}}{1 - \nu_{12}\nu_{21}}\,, \\
+					Q_{12} & = \frac{E_{1}\nu_{21}}{1 - \nu_{12}\nu_{21}} = Q_{21}\,, \\
+					Q_{21} & = \frac{E_{2}\nu_{12}}{1 - \nu_{12}\nu_{21}} = Q_{12}\,, \\
+					Q_{22} & = \frac{E_{2}}{1 - \nu_{12}\nu_{21}}\,, \\
+					Q_{66} & = G_{12}\,.
+				\end{aligned}
+				\end{equation}
+
+				For each layer $k$, the reduced stiffness matrix is transformed to account for the layer orientation angle $\theta$ within the laminate coordinate system, yielding a reduced transformed stiffness matrix according to:
+
+				\begin{equation}
+					[\bar{Q}] = [T]^{-1}[Q][T]^{-T}\,,
+				\end{equation}
+
+				where the transformation matrix $[T]$ is defined as:
+
+				\begin{equation}
+				[T] =
+				 \begin{bmatrix}
+					 \cos^2\theta & \sin^2\theta & 2\cos\theta\sin\theta \\
+					 \sin^2\theta & \cos^2\theta & -2\cos\theta\sin\theta \\
+					 -\cos\theta\sin\theta & \cos\theta\sin\theta & \cos^2\theta - \sin^2\theta
+				 \end{bmatrix}\,.
+				\end{equation}
+
+				Using the lay-ups specified for the cuticle permutations, we calculate the extensional stiffness matrix $[A]$, bending stiffness matrix $[D]$, and bending-extension coupling matrix $B$ for each laminate consisting of $n$ layers at a distance $z$ from the laminate mid-plane.
+				The elements of these matrices can be found according to:
+
+				\begin{equation}
+					\begin{aligned}
+						A_{ij} & = \sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z_{k} - z_{k-1})\,, \\
+						B_{ij} & = \frac{1}{2}\sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z^2_{k} - z^2_{k-1})\,, \\
+						D_{ij} & = \frac{1}{3}\sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z^3_{k} - z^3_{k-1})\,.
+					\end{aligned}
+				\end{equation}
+
+				These stiffness matrices relate vectors of resultant forces $\{N\}$ and bending moments $\{M\}$ to mid-surface strains and curvatures $\{\epsilon^{\degree}\}$ and $\{\kappa\}$, respectively, in the laminate according to the following relationship:
+
+				\begin{equation}
+					\begin{Bmatrix}
+						\{N\} \\
+						\{M\}
+					\end{Bmatrix}
+					=
+					\begin{bmatrix}
+						[A] & [B] \\
+						[B] & [D]
+					\end{bmatrix}
+					\begin{Bmatrix}
+					\{\epsilon^{\degree}\} \\
+					\{\kappa\}
+					\end{Bmatrix}\,.
+				\end{equation}
+
+				For symmetric laminates, $[B] = 0$, and therefore:
+
+				\begin{equation}
+				\label{eqn::abbd}
+					\begin{aligned}
+						\{N\} & = [A]\{\epsilon^{\degree}\}\,, \\
+						\{M\} & = [D]\{\kappa\}\,,						
+					\end{aligned}\,,
+				\end{equation}
+
+				or, in expanded form:
+
+				\begin{equation}
+					\begin{aligned}
+						\begin{Bmatrix}
+							N_{xx} \\
+							N_{yy} \\
+							N_{xy}
+						\end{Bmatrix}
+						& =
+						\begin{bmatrix}
+							A_{11} & A_{12} & A_{16} \\
+							A_{21} & A_{22} & A_{26} \\
+							A_{61} & A_{62} & A_{66}
+						\end{bmatrix}
+						\begin{Bmatrix}
+							\epsilon^{\degree}_{xx} \\
+							\epsilon^{\degree}_{yy} \\
+							\gamma^{\degree}_{xy}
+						\end{Bmatrix}\,,
+						\\
+						\begin{Bmatrix}
+							M_{xx} \\
+							M_{yy} \\
+							M_{xy}
+						\end{Bmatrix}
+						& =
+						\begin{bmatrix}
+							D_{11} & D_{12} & D_{16} \\
+							D_{21} & D_{22} & D_{26} \\
+							D_{61} & D_{62} & D_{66}
+						\end{bmatrix}
+						\begin{Bmatrix}
+							\kappa_{xx} \\
+							\kappa_{yy} \\
+							\kappa_{xy}
+						\end{Bmatrix}\,.
+					\end{aligned}
+				\end{equation}
+
+				If we make the simplifying assumptions (see ref.\info{reddy, 2007 and jones, 1998}) that \textbf{(1)} the laminate experiences pure axial loading and transverse bending (i.e., $N_{yy} = N_{xy} = 0$ and $M_{yy} = M_{xy} = 0$, respectively) and \textbf{(2)} the laminate is a beam of sufficiently high aspect ratio to minimize the Poisson effect and anisotropic shear coupling (i.e., below we effectively let $A^*_{12} = A^*_{16} = 0$ and $D^*_{12} = D^*_{16} = 0$), then we can calculate the in-plane effective flexural and axial Young's moduli of the laminate along the x-axis.
+
+				For axial Young's modulus of the laminate, we first define the average membrane stresses in the laminate as:
+
+				\begin{equation}
+					\{\bar{\sigma}^m\} = \frac{\{N\}}{z_1 - z_n}\,.
+				\end{equation}
+
+				By substitution in Eq.~\ref{eqn::abbd}, we find:
+
+				\begin{equation}
+					\begin{aligned}
+						\begin{Bmatrix}
+							\bar{\sigma}^m_{xx} \\
+							\bar{\sigma}^m_{yy} \\
+							\bar{\tau}^m_{xy}
+						\end{Bmatrix}
+						& = \frac{1}{(z_1 - z_n)}
+						\begin{bmatrix}
+							A_{11} & A_{12} & A_{16} \\
+							A_{21} & A_{22} & A_{26} \\
+							A_{61} & A_{62} & A_{66}
+						\end{bmatrix}
+						\begin{Bmatrix}
+							\epsilon^m_{xx} \\
+							\epsilon^m_{yy} \\
+							\gamma^m_{xy}
+						\end{Bmatrix}\,,
+					\end{aligned}
+				\end{equation}
+
+				and, by inverting this equation (let $A^* = A^{-1}$) and substituting based on the assumptions above, we infer:
+
+				\begin{equation}
+					\epsilon^m_{xx} = (z_1 - z_n)A^{*}_{11}\bar{\sigma}^m_{xx}\,.
+				\end{equation}
+
+				We therefore define Young's modulus for effective axial elasticity as:
+
+				\begin{equation}
+					E^m_{xx} = \frac{\bar{\sigma}^m_{xx}}{\epsilon^m_{xx}} = \frac{1}{A^*_{11}(z_1 - z_k)}
+				\end{equation}
+
+				To find the transverse flexural Young's modulus of the laminate, we first specify the moment-curvature relation of an Euler-Bernoulli beam:
+
+				\begin{equation}
+				\label{eqn::bee}
+				M = EI\kappa\,.
+				\end{equation}
+
+				Along the x-axis, the second moment of area for a rectangular cross-section is:
+
+				\begin{equation}
+					I_{yy} = \frac{b(z_1 - z_n)^3}{12}\,,
+				\end{equation}
+
+				Given the assumption that $M_{yy} = M_{xy} = 0$, the moment along the x-axis is related to the moment of the beam by:
+
+				\begin{equation}
+					M = M_{xx}b\,.
+				\end{equation}
+
+				Thus, given the assumption that $D_{12} = D_{16} = 0$, Young's modulus for the effective transverse flexural elasticity of the laminate can be found by making Eq.~\ref{eqn::bee} specific to transverse flexure of the x-axis and rearranging the terms:
+
+				\begin{equation}
+					E^m_{xx} = \frac{12M_{xx}}{(z_1 - z_n)^3\kappa_{xx}}
+				\end{equation}
+
+				From inversion of Eq.~\ref{eqn::abbd} (let $D^* = D^{-1}$) this reduces to:
+
+				\begin{equation}
+					E^m_{xx} = \frac{12}{(z_1 - z_n)^3D^*_{11}}
+				\end{equation}
+
 		\subsection*{Statistical analysis}
@@
 			\paragraph*{Model selection and fitting}
+
 			\paragraph*{Estimating phylogenetic signal}

### 2019-01-28 matlab/effective_moduli.m

@@ -1,0 +1,45 @@
+%Author: Andrew Jansen
+%Date: January 2019
+%Title: Effective moduli of Curculio cuticle
+
+%--------------------------------------------------------------------------
+%Elastic constants for exocuticle, denoted by 'x'
+%E denotes Young's modulus along material axis denoted by number
+Ex1 = 6.4;
+Ex2 = 6.4;
+%nu denotes Poisson ratio along material axes denoted by numbers
+nux12 = 0.36;
+nux21 = 0.36;
+%G denotes shear modulus along material axes denoted by numbers
+Gx12 = 2.5;
+%p denotes the denominator used to find elements of stiffness matrix
+px = 1 - nux12 * nux21;
+
+%--------------------------------------------------------------------------
+%For macrofiber (a.k.a. 'balken') layer in endocuticle, denoted by 'b'
+Eb1 = 8.5;
+Eb2 = 0.52;
+nub12 = 0.49;
+nub21 = 0.03;
+Gb12 = 0.17;
+
+function Q = lamina(S)
+%Elements of 2D reduced stiffness matrix, denoted Q
+pb = 1 - nub12 * nub21;
+%E*nu is averaged to correct for rounding error in the published values
+%because Eb1 * nub21 = Eb2 * nub12 is actually true for this material.
+Enub_avg = (Eb1 * nub21 + Eb2 * nub12) / 2;
+
+Qb11 = Eb1 / pb;
+Qb12 = Enub_avg / pb;
+Qb16 = 0;
+Qb21 = Enub_avg / pb;
+Qb22 = Eb2 / pb;
+Qb26 = 0;
+Qb61 = 0;
+Qb62 = 0;
+Qb66 = Gb12;
+Q = [Qb11 Qb12 Qb16;
+      Qb21 Qb22 Qb26;
+      Qb61 Qb62 Qb66];
+end

### 2019-01-29 tex_v2/manuscript.tex
moduli

@@
 \usepackage{gensymb}
+\usepackage{textgreek}
 \usepackage{xcolor}
@@
 %	Paragraph 5 - short segue/ signposting
-	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite typified by variously arranged $\alpha$-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
-	Although $\alpha$-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
+	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite typified by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
+	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
 	Beetle cuticle is divided into two structurally and mechanically distinct regions, the (outer) exocuticle and (inner) endocuticle.
-	The exocuticle is characterized by a transversely-isotropic helicoidal structure that is comparatively rigid (though brittle) and has been the subject of many prior studies\unsure{numbers and refs}.
+	The exocuticle is characterized by a transversely-isotropic helicoidal structure that is comparatively rigid (though brittle) and has been the subject of numerous prior studies\unsure{numbers and refs}.
 	By contrast, the endocuticle of beetles is comprised of large, aligned bundles (macrofibers) of chitin that are strongly anisotropic\unsure{numbers again}.
@@
 	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
-	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we predict the effect of these differences using Classical Laminate Theory (CLT).
+	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we estimate the effect of these differences using Classical Laminate Plate Theory (CLPT).
+
 	Because recent studies have shown that the yield strength of the weevil rostrum exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
@@
 	\section{Microstructure of the \textit{Curculio} rostrum}
-		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4nm) embedded in a proteinaceous matrix.
+		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
 		These layers are stacked at a more or less constant angle to each other, thus forming a quasi-isotropic laminate referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}.
-		This layout effectively produces a transversely isotropic composite, mitigating the strong anisotropy of $\alpha$-chitin to yield a versatile building material for the exoskeleton.
+		This layout effectively produces a transversely isotropic composite, mitigating the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton.
 
-		Conversely, the endocuticle of Coleoptera retains the anisotropy of $\alpha$-chitin to improve fracture resistance of the cuticle.
-		Beetle endocuticle is unique among arthropods and is comprised of large (5-20$\mu$m OD) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic \improvement{insert E1, E2, E3 here} and arranged in unidirectional plies, as in Figs. \info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
+		Beetle endocuticle, however, is unique among arthropods and is comprised of large (5-20\,{\textmu}m OD) unidirectional bundles of chitin, called macrofibers.
+		Chitin macrofibers are orthotropic ($E_1 = 8.5$\,GPa, $E_2 = E_3 = 0.52$\,GPa \cite{Jansen2016}) and arranged in unidirectional plies, as in Figs. \info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
 		Typically, adjacent macrofiber plies are paired and pseudo-orthogonal (i.e., angled approx.
@@
 		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is resposible for what we observed.
-		The endocuticle is made of aligned $\alpha$-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
+		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
 		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
@@
 		Endocuticle, on the other hand exhibited severe delamination, ply-splitting, and fiber pulling, consistent with viscoelastic/plastic behavior shown in previous studies and congruent with theoretical consideration of the microstructure of this material (name something here).
-		These patterns indicate that the endocuticle is probably less brittle than the exocuticle, most likely due to the alignment of the $\alpha$-chitin fibers in each macrofiber.
+		These patterns indicate that the endocuticle is probably less brittle than the exocuticle, most likely due to the alignment of the \textalpha-chitin fibers in each macrofiber.
 
@@
 		\section*{References}
+			\bibitem{Jansen2016}
+				Jansen, M. A., Singh, S. S., Chawla, N., \& Franz, N. M.
+				A multilayer micromechanical model of the cuticle of \textit{Curculio longinasus} Chittenden, 1927 (Coleoptera: Curculionidae).
+				\textit{J. Struct. Biol.}
+				\textbf{195: 2,}
+				139--158,
+				(2016).
+
 			\bibitem{Klocke2011}
@@
 				We assumed equal layer thicknesses in the endocuticle of the basal cuticle, and equal thicknesses of exocuticle and endocuticle.
-				In the apical cuticle we assumed the exocuticle and upper 8 layers of the endocuticle were each 5 microns thick, with 4 thinner layers of endocuticle, each 1.5 microns thick.
+				In the apical cuticle we assumed the exocuticle and upper 8 layers of the endocuticle were each 5 microns thick, with 4 thinner layers of endocuticle, each 1.25\,{\textmu}m thick.
 				The angles of the endocuticle layers for both types of cuticle are as described above.
@@
 				\begin{equation}
-					E^m_{xx} = \frac{\bar{\sigma}^m_{xx}}{\epsilon^m_{xx}} = \frac{1}{A^*_{11}(z_1 - z_k)}
+					E^m_{xx} = \frac{\bar{\sigma}^m_{xx}}{\epsilon^m_{xx}} = \frac{1}{(z_1 - z_k)A^*_{11}}
 				\end{equation}
@@
 				\begin{equation}
-					E^m_{xx} = \frac{12M_{xx}}{(z_1 - z_n)^3\kappa_{xx}}
+					E^f_{xx} = \frac{12M_{xx}}{(z_1 - z_n)^3\kappa_{xx}}
 				\end{equation}
@@
 				\begin{equation}
-					E^m_{xx} = \frac{12}{(z_1 - z_n)^3D^*_{11}}
+					E^f_{xx} = \frac{12}{(z_1 - z_n)^3D^*_{11}}
 				\end{equation}

### 2019-01-29 matlab/effective_moduli.m

@@ -3,17 +3,22 @@
 %Title: Effective moduli of Curculio cuticle
 
 %--------------------------------------------------------------------------
+%Definition of elastic constants by cuticle region
+%--------------------------------------------------------------------------
 %Elastic constants for exocuticle, denoted by 'x'
 %E denotes Young's modulus along material axis denoted by number
 Ex1 = 6.4;
 Ex2 = 6.4;
+
 %nu denotes Poisson ratio along material axes denoted by numbers
 nux12 = 0.36;
 nux21 = 0.36;
+
 %G denotes shear modulus along material axes denoted by numbers
 Gx12 = 2.5;
-%p denotes the denominator used to find elements of stiffness matrix
-px = 1 - nux12 * nux21;
+
+%Matrix of constants for exocuticle
+exo = [Ex1 Ex2 nux12 nux21 Gx12];
 
 %--------------------------------------------------------------------------
 %For macrofiber (a.k.a. 'balken') layer in endocuticle, denoted by 'b'
@@ -23,23 +28,162 @@ nub12 = 0.49;
 nub21 = 0.03;
 Gb12 = 0.17;
 
-function Q = lamina(S)
-%Elements of 2D reduced stiffness matrix, denoted Q
-pb = 1 - nub12 * nub21;
-%E*nu is averaged to correct for rounding error in the published values
-%because Eb1 * nub21 = Eb2 * nub12 is actually true for this material.
-Enub_avg = (Eb1 * nub21 + Eb2 * nub12) / 2;
+endo = [Eb1 Eb2 nub12 nub21 Gb12];
 
-Qb11 = Eb1 / pb;
-Qb12 = Enub_avg / pb;
-Qb16 = 0;
-Qb21 = Enub_avg / pb;
-Qb22 = Eb2 / pb;
-Qb26 = 0;
-Qb61 = 0;
-Qb62 = 0;
-Qb66 = Gb12;
-Q = [Qb11 Qb12 Qb16;
-      Qb21 Qb22 Qb26;
-      Qb61 Qb62 Qb66];
+cuticle = [exo; endo];
+%--------------------------------------------------------------------------
+%Cuticle lay-up: layer thicknesses, orientation angles & stacking sequence  
+%--------------------------------------------------------------------------
+%sequence for basal portion of head
+base_z = (1 / 480) * [-24 -12:-1 1:12 24]; %48 layers, 50 microns => mm
[…]

### 2019-01-30 tex_v2/manuscript.tex
edits

@@
 }	
-\section*{Introduction}
+\section*{Introduction} \info{650 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
 %Intro outline - 500 word maximum
@@
 	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
-	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.\info{630 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
+	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.
 }	
 	\section{Microstructure of the \textit{Curculio} rostrum}
+	\info{850 words, reduce in revisions}
 		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
@@
 
-		Beetle endocuticle, however, is unique among arthropods and is comprised of large (5-20\,{\textmu}m OD) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic ($E_1 = 8.5$\,GPa, $E_2 = E_3 = 0.52$\,GPa \cite{Jansen2016}) and arranged in unidirectional plies, as in Figs. \info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
+		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter) unidirectional bundles of chitin, called macrofibers.
+		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, as in Figs.
+\info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
 		Typically, adjacent macrofiber plies are paired and pseudo-orthogonal (i.e., angled approx.
@@
 
-		Serial thin sectioning and SEM of fractured \textit{Curculio} specimens has revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
+		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens has revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
 		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle\info{thickness measurements here}.
@@
 
-		To investigate the effect of these cuticle modifications, we estimated the engineering and flexural elastic moduli of the cuticle in both the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
+		To investigate the effect of these cuticle modifications, we estimated uni-axial membrane and transverse flexural Young's moduli of the cuticle in both the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
 		We previously derived the effective elastic constants of the cuticle regions of \textit{C. longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
-		The cuticle of the head capsule has moduli\info{insert numbers here}.
-		The cuticle of the rostral apex has moduli\info{insert numbers here}.
+		The cuticle was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively, in the head capsule; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
 
-		Interestingly, each of these modifications contributes to the observed flexibility of the snout.
+		TO assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications, namely:
+		\begin{enumerate}
+			\item A hybrid cuticle with the angle stacking sequence of regular cuticle (the head capsule) but possessing the layer thicknesses of the rostral apex.
+$E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases, respectively.
+			\item A hybrid cuticle with the layer thicknesses of regular cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle).
+$E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases, respectively.
+		\end{enumerate}
 
-		\info{some explanation in source code}
-		\noshow{Because the exocuticle is generally more rigid (and brittle) than the endocuticle due to its transverse isotropy and microstructure, a reduction in the relative thickness of this region would likely have the effect of increasing the flexibility of the resulting composite.
-		Additionally, aligning the layers of endocuticle as a 45 deg cross ply would likely reduce the bending moment of the structure.
-		The combined effect of these modifications is that the apex of the rostrum is able to bend until completely straight without fracture.}
+		Interestingly, each of the cuticle modifications noted in the rostral apex  individually decreased the elastic moduli of the cuticle; however, they appear to have a synergistic combined effect on cuticle elasticity, rather than a simple additive effect.
+		This result suggests that both modifications are necessary in order for the snout to function properly in the living animal, where the combined effect allows the rostrum to bend until completely straight without fracture.
 
-		Below we demonstrate that the endocuticle does not vary in tensile strength across the rostrum or between species, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
+		\change{maybe edit, move to conclusions, or just delete this paragraph}Below we demonstrate that the endocuticle does not vary in tensile strength across the rostrum or between species, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
 		The available evidence therefore suggests that the relative flexibility of the snout is primarily derived from the composite profile of its cuticle.
@@
 	\section{Force-controlled loading to fracture}
+	%475 words, try to keep around this length
 		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
@@
 	\section{Load cycling of \textit{Curculio caryae}}
+	%375 words, add no more than 100, ONLY IF NEEDED
 		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
@@
 		~400K cycles, complete elongation, rehydrated, coated in grease to prevent loss of moisture and stiffening of the specimen.
-		We observed visco-elastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
+		We observed viscoelastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
 		Fmax decreased logarithmically with cycle number, etc., and the specimen appeared to have deformed plastically during the test.
@@
 		Given this surprising result, it appears that the specimen was undamaged by the testing.
-		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/visco-elastic.
+		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/viscoelastic.
 
-		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is resposible for what we observed.
-		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystaline structure is enforced by hydrogen bonds between individual chitin chains along their length.
+		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is responsible for what we observed.
+		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains along their length.
 		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
-		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform visco-elastically/visco-plastically.
+		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform viscoelastically/viscoplastically.
 		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain without any damage to the specimen.
@@
 	\section{Fractography of test specimens}
+	%currently ~400 words, keep to this length
 		Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).
@@
 		The presence of shear-cusps therefore indicates that the fibers of the endocuticle are shearing past each other within each ply along the radial/normal plane (type II shearing???, extension shear-coupling).
-		Furthermore, given the delamination observed between plies, the layers of endocuticle are liekly shearing past each other (type III shearing) along the transverse plane.
-		We therefore infer that, under tension, the endocuticle tends to deform visco-elastically and plastically along the longitudinal axes of the macrofibers, while the overlying exocuticle exhibits brittle fracture due to shearing between the stretching endocuticle fibers to which it is anchored.
+		Furthermore, given the delamination observed between plies, the layers of endocuticle are likely shearing past each other (type III shearing) along the transverse plane.
+		We therefore infer that, under tension, the endocuticle tends to deform viscoelastically and plastically along the longitudinal axes of the macrofibers, while the overlying exocuticle exhibits brittle fracture due to shearing between the stretching endocuticle fibers to which it is anchored.
 
@@
 	\section{Conclusions}
-	%1-2 'short paragraphs'
+	%1-2 'short paragraphs', 300 words maximum
 

### 2019-01-30 matlab/effective_moduli.m

@@ -36,9 +36,10 @@ cuticle = [exo; endo];
 %--------------------------------------------------------------------------
 %sequence for basal portion of head
 base_z = (1 / 480) * [-24 -12:-1 1:12 24]; %48 layers, 50 microns => mm
-base_theta = 30 * [0 3:8 -(-5:0); 0:5 -(-8:-3) 0]; %correct angle sequence
+base_theta = 30 * [ 3:8 -(-5:0); 0:5 -(-8:-3)]; %30 deg angle sequence
+%subtract 5 deg  after matrix to maximize flexural stiffness
 base_seq = [1, repelem(2, 24) 1]; %1 for exocuticle, 2 for endocuticle
-base_layup = [base_z; base_theta(1:26); base_seq];
+base_layup = [base_z; 0 base_theta(1:24) 0; base_seq];
 
 %--------------------------------------------------------------------------
 %sequence for apical portion of head
@@ -49,7 +50,7 @@ apex_seq = [1, repelem(2, 24) 1]; %same sequence as base
 apex_layup = [apex_z; apex_theta(1:26); apex_seq];
 
 %--------------------------------------------------------------------------
-%permutations: exo thickness, endo angle; exo angle, endo thickness
+%permutations: base thickness, apex angle; base angle, apex thickness
 mod_theta_layup = [base_layup(1, :); apex_layup(2, :); base_layup(3, :)];
 
 mod_z_layup = [apex_layup(1, :); base_layup(2, :); base_layup(3, :)];

### 2019-02-01 tex_v2/manuscript.tex
updates

@@
 %	Paragraph 5 - short segue/ signposting
-	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite typified by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
+	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
 	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
@@
 	This structure is used by the female to feed and excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
-	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit.
+	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit; a single female may prepare hundreds of such sites during the adult stage.
 
@@
 		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, as in Figs.
+		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.
 \info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
@@
 		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle\info{thickness measurements here}.
-		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}E) in two key characteristics:
+		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}) in two key characteristics:
 
@@
 		In previous work we identified these modifications to the composite structure of the cuticle within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
-		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (listed below), indicating that this is likely a genus-wide trait.
-		We note that in all species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.
+		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
+		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.
 
@@
 
-		TO assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications, namely:
+		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications, namely:
 		\begin{enumerate}
@@
 
-		Interestingly, each of the cuticle modifications noted in the rostral apex  individually decreased the elastic moduli of the cuticle; however, they appear to have a synergistic combined effect on cuticle elasticity, rather than a simple additive effect.
+		Each of the cuticle modifications noted in the rostral apex  individually decreased the elastic moduli of the cuticle; however, they appear to have a synergistic combined effect on cuticle elasticity, rather than a simple additive effect.
 		This result suggests that both modifications are necessary in order for the snout to function properly in the living animal, where the combined effect allows the rostrum to bend until completely straight without fracture.

### 2019-02-07 tex_v2/manuscript.tex
added analysis pipeline, data, and plots

@@
 		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
-		Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
+		Consequently, there was a negative correlation between utlimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
 		In other words, UTS is strongly correlated with the cross-sectional area of the endocuticle across species.

### 2019-02-07 analysis/phylo_regress.R

@@ -1,0 +1,739 @@
+# Andrew Jansen
+# created x.10.2018
+
+# Miscellaneous Functions ------------------------------------------------------
+# Import the data
+GetData = function(file.name) {
+  test.data = data.frame(read.csv(file.name, sep = ','))
+  return(test.data)
+}
+
+# Analysis 1 -------------------------------------------------------------------
+# Fits models with and without log transformation and phylogenetic covariance
+Analysis1 = function(phylogeny, raw.data) {
+  
+  # Fit models with variable transformation and phylogenetic covariance
+  simple.model = lme(phenotype ~ cofactor,
+                     random = ~ 1 | species,
+                     data = raw.data,
+                     method = 'ML')
+  
+  simple.brownian = lme(phenotype ~ cofactor,
+                        random = ~ 1 | species,
+                        data = raw.data,
+                        correlation = phylogeny,
+                        method = 'ML')
+  
+  y.transformed = lme(log(phenotype) ~ cofactor,
+                           random = ~ 1 | species,
+                           data = raw.data,
+                           method = 'ML')
+  
+  y.brownian = lme(log(phenotype) ~ cofactor,
+                             random = ~ 1 | species,
+                             data = raw.data,
+                             correlation = phylogeny,
+                             method = 'ML')
+  
+  x.transformed = lme(phenotype ~ log(cofactor),
+                      random = ~ 1 | species,
+                      data = raw.data,
+                      method = 'ML')
+  
+  x.brownian = lme(phenotype ~ log(cofactor),
+                   random = ~ 1 | species,
+                   data = raw.data,
+                   correlation = phylogeny,
+                   method = 'ML')
+  
+  xy.transformed = lme(log(phenotype) ~ log(cofactor),
+                       random = ~ 1 | species,
+                       data = raw.data,
+                       method = 'ML')
+  
+  xy.brownian = lme(log(phenotype) ~ log(cofactor),
+                    random = ~ 1 | species,
+                    data = raw.data,
+                    correlation = phylogeny,
+                    method = 'ML')
+  
[…]

### 2019-02-07 analysis/phylo_regress_bayesian.R

@@ -1,0 +1,332 @@
+# Andrew Jansen
+# created iv.15.2018
+
+# Miscellaneous Functions ------------------------------------------------------
+#Import the data
+GetData = function(file.name) {
+  test.data = data.frame(read.csv(file.name, sep = ','))
+  return(test.data)
+}
+
+#For ANOVA
+SqDev = function(x) {var(x) * (length(x) - 1)}
+
+# Analysis 1 -------------------------------------------------------------------
+# Compute co-variance matrix from inverse
+Analysis1 = function(phylo) {
+  inv.phylo = MCMCglmm::inverseA(phylo, nodes="TIPS")
+  A = solve(inv.phylo$Ainv)
+  rownames(A) = rownames(inv.phylo$Ainv)
+  print(A)
+  return(A)
+}
+
+# Analysis 2 -------------------------------------------------------------------
+# fit hierarchical phylogenetic model
+Analysis2 = function(raw.data, A, priors) {
+  phylo_model = brm(
+    phen ~ spec_mean_cf + (1|phylo) + (1|species),
+    data = raw.data,
+    family = gaussian(),
+    cov_ranef = list(phylo = A),
+    prior = priors,
+    sample_prior = TRUE#,
+    #chains = 16,
+    #cores = 8,
+    #iter = 2000,
+    #warmup = 1000
+  )
+  print(summary(phylo_model))
+  plot(phylo_model)
+  gg = plot(marginal_effects(phylo_model))[[1]]
+  plot(gg + geom_point(inherit.aes = FALSE,
+                  data = raw.data,
+                  mapping = aes(x = spec_mean_cf, y = phen, color = phylo)))
+  
+  return(phylo_model)
+}
+
+# Analysis 3 -------------------------------------------------------------------
+# Hypothesis testing and phylogenetic signal
+Analysis3 = function(phylo_model) {
+  hyp = paste("sd_phylo__Intercept^2 /",
+              "(sd_phylo__Intercept^2 + sd_species__Intercept^2 + sigma^2) = 0")
+  hyp = hypothesis(phylo_model, hyp, class = NULL)
+  
+  print(hyp)
+  plot(hyp)
+}
+
+# Analysis 4 -------------------------------------------------------------------
+# Update model and re-run phylogenetic signal
[…]

### 2019-02-07 analysis/phylo_regress_messy.R

@@ -1,0 +1,316 @@
+# Andrew Jansen
+# created iv.15.2018
+
+# Miscellaneous Functions ------------------------------------------------------
+#Import the data
+GetData = function(file.name) {
+  test.data = data.frame(read.csv(file.name, sep = ','))
+  return(test.data)
+}
+
+#For ANOVA
+SqDev = function(x) {var(x) * (length(x) - 1)}
+
+# Analysis 1 -------------------------------------------------------------------
+# ADD NAMES FOR LABELS ON GRAPHS TO FUNCTION INPUT
+Analysis1 = function(phylo, raw.data, axis.labels) {
+  
+  lm_mod = lme(phen ~ cf,
+               random = ~ 1|species,
+               data = raw.data,
+               correlation = corBlomberg(6, phylo, fixed = TRUE),
+               method = 'ML')
+  
+  lm_mod_alt = lme(phen ~ cf,
+                   random = ~ 1|species,
+                   data = raw.data,
+                   method = 'ML')
+  
+  print(summary(lm_mod))
+  print(summary(lm_mod_alt))
+  print(anova(lm_mod))
+  print(anova(lm_mod_alt))
+  
+  print(lrtest(lm_mod, lm_mod_alt))
+  
+  #create data.frame with new values for predictors
+  #more than one predictor is possible
+  cf = raw.data$cf
+  new.dat <- data.frame(cf)
+  
+  #predict response
+  new.dat$pred <- predict(lm_mod, newdata=new.dat, level=0)
+  
+  #create design matrix
+  Designmat <- model.matrix(eval(eval(lm_mod$call$fixed)[-2]),
+                            new.dat[-ncol(new.dat)])
+  
+  #compute standard error for predictions
+  predvar = diag(Designmat %*% lm_mod$varFix %*% t(Designmat))
+  new.dat$SE = sqrt(predvar) 
+  new.dat$SE2 = sqrt(predvar+lm_mod$sigma^2)
+  
+  gg = ggplot(new.dat,aes(x=cf,y=pred)) + 
+    geom_ribbon(aes(ymin=pred-2*SE2,ymax=pred+2*SE2),alpha=0.2,fill='black') +
+    geom_ribbon(aes(ymin=pred-2*SE,ymax=pred+2*SE),alpha=0.2,fill='black') +
[…]

### 2019-02-07 analysis/tensile.py

@@ -1,0 +1,450 @@
+# -*- coding: utf-8 -*-
+'''
+Created on Tue Aug  1 13:59:50 2017
+
+@author: Flashcock
+'''
+
+import os
+from scipy.stats import linregress as linreg
+from itertools import takewhile
+from numpy import array, poly1d, sqrt, absolute, arcsin, pi
+from scipy.optimize import curve_fit, minimize_scalar
+from scipy.integrate import quad
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from mpl_toolkits.mplot3d import Axes3D
+from numpy import log as ln
+
+class specimen:
+    '''
+    Args should be formatted as follows:
+        asuid (str or int), species (str), length (str), exo (str), endo (str),
+        total (str), N (list [float]), mm (list [float]), Sec (list [float]).
+    These are converted to their respective types below and used to calculate
+        true and engineering stress and strain, young's modulus, and UTS.
+    '''
+    def __init__(self, asuid, species, length, exo, endo, total, N, mm, Sec):
+        self.__asuid__ = asuid
+        self.__species__ = species.lower()
+
+        self.__length__ = float(length)
+        self.__exo__ = float(exo)
+        self.__endo__ = float(endo)
+        self.__total__ = float(total)
+
+        self.__N__ = array(N)
+        self.__mm__ = array(mm) - min(mm)
+        self.__Sec__ = array(Sec) - Sec[0]
+
+        self.__sigmaN__ = self.__N__ / self.__total__ #P/A0
+        self.__epsilonN__ = (self.__mm__ / self.__length__) #dL/L0
+        self.__sigmaT__ = self.__sigmaN__ * (1 + self.__epsilonN__) #s(1+e)
+        self.__epsilonT__ = ln(1 + self.__epsilonN__) #ln(1+e)
+
+        self.__fmax__ = max(N)
+        self.__uts__ = max(self.__sigmaN__) #evaluated from engineering stress
+
+        self.__dmax__ = max(self.__mm__)
+        self.__ufs__ = max(self.__epsilonN__) #evaluated from engineering strain
+
+
+        self.__Eindex__ = []
+        lower = 0.00 #self.__ufs__ * 0
+        upper = 0.015 #self.__ufs__ * 0.20
+        self.__Eindex__.append(len([x for x in takewhile(lambda x: x[1] <= lower, enumerate(self.__epsilonN__))]))
[…]

### 2019-02-08 tex_v2/manuscript.tex
file cleanup

--- 2019-02-01 tex_v2/manuscript.tex
@@
 
-		\change{maybe edit, move to conclusions, or just delete this paragraph}Below we demonstrate that the endocuticle does not vary in tensile strength across the rostrum or between species, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum.
-		The available evidence therefore suggests that the relative flexibility of the snout is primarily derived from the composite profile of its cuticle.
+	\section{Force-controlled loading to fracture}
+	%550 words, maybe reduce?? Still must add numbers...
 
-	\section{Force-controlled loading to fracture}
-	%475 words, try to keep around this length
 		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
 		Although the heads were rehydrated by immersion in de-ionized water for 24 hours, we observed comparatively brittle fracture (but see below), in contrast to GIVE EXAMPLES HERE.
-		The tensile behavior for the cuticle of these weevils is a result of its microstructure, which lacks pores, etc (give reasons).
-		THIS IS ALSO A GOOD SPOT TO SIGNPOST WHAT I FOUND IN GENERAL.
 
+		In general, the behavior of the rostrum is non-linear viscoelastic, and, as in other biological materials, exhibits a stress-strain curve characterized by a steady increase in stiffness, beginning with a toe-in region and terminating in apparently brittle fracture after a period of strain hardening.
 		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
-		Consequently, there was a negative correlation between utlimate tensile strenth of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture, indicating that as the proportion of exocuticle increased, tensile strength decreased.
-		In other words, UTS is strongly correlated with the cross-sectional area of the endocuticle across species.
+		Consequently, there was a negative correlation between utlimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
+		In other words, UTS is positively correlated with the relative proportion of the endocuticle across species.
+
+		Although the proportion of exocuticle is predicted to be positively correlated with the stiffness of the resulting composite, we found no such associations because of the anterior-posterior gradient in cuticle profile.
+		Because the cross-sectional areas of the cuticle regions vary across the length of the head, it is meaningless to attempt to correlate these areas to properties of the entire snout.
+		Instead, we found that the uniaxial elastic moduli of the snout under low strain (E-low) was inversely correlated with the ultimate strain and toughness of the snout (as was the secant modulus, E-sec).
 		These associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
 
-		These data have three important implications:
-		\begin{enumerate}
-			\item There is very little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology. (singh 2016, jansen 2016)
-			\item The endocuticle contributes more to the tensile strength of the rostrum than the exocuticle, possibly because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
-			\item Thus, in addition to making the cuticle more flexible, the altered apical composite profile makes it less likely to break, suggesting a possible means by which the system evolved (via negative selection of breakage, maybe say this in conclusions).
-		\end{enumerate}
+		From our results, we infer that there is little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum (singh 2016, jansen 2016).
+		Furthermore, the endocuticle apparently contributes more to the tensile strength of the rostrum than the exocuticle, likely because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
+		Thus, in addition to making the cuticle more flexible, the altered apical composite profile makes the rostrum less likely to break, suggesting a possible means by which this system evolved\info{(via negative selection of breakage, maybe say this in conclusions)}.
 
-		Finally, we observed that the elastic modulus is higher in specimens with a longer snout.
+		\info{DO I really want to discuss this? Occupies precious space...}Finally, we observed that the elastic modulus is higher in specimens with a longer snout.
 		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) would need to be more flexible to avoid fracture during oviposition.
-		Based on preliminary confocal microscopy data, we speculate that this phenomenon may be the result of a longer scrobe/transition/gradient from the basal profile to the apical profile in longer rostra.
+		Based on preliminary confocal microscopy data, we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
 		A decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
-		[OR SAY MORE CONCISELY THAT: based on prelim CLSM data we believe that the snout is reinforced against bucking with thicker exocuticle in the base of these species; as a result, the higher volume fraction of exocuticle in the snout increases young's modulus for such species]
 
@@
 	\section{Fractography of test specimens}
-	%currently ~400 words, keep to this length
+	%currently ~400 words, keep to this length or reduce to 2 short paragraphs
 		Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).

### 2019-02-08 analysis/phylo_regress.R

@@ -452,8 +452,8 @@ Analysis4 = function(phylogeny, raw.data, axis.labels) {
 main = function() {
   # Clear workspace
   rm(list = ls())
-  setwd(paste('C:/Users/Andrew/Documents/Current Projects/',
-              'Curculio Mechanical Testing/Tensile Testing/Analysis', sep=''))
+  setwd(paste('C:/Users/Andrew/Documents/GitHub/',
+              'curculio-tensile/analysis', sep=''))
   
   # Load gplots for plotCI
   library(ggplot2)
@@ -589,7 +589,7 @@ main = function() {
   writeLines('Comparison 4a: Etan low vs. length')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
-  raw.data$phenotype = raw.data$uts / raw.data$ufs
+  raw.data$phenotype = raw.data$Etanl
   raw.data$cofactor = raw.data$length
   Analysis1(phylo.vcv, raw.data)
 
@@ -597,10 +597,10 @@ main = function() {
   readline(prompt = 'Etan high vs. length: press enter to continue.')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
-  writeLines('Comparison 4a: Etan high vs. length')
+  writeLines('Comparison 4b: Etan high vs. length')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
-  raw.data$phenotype = raw.data$uts / raw.data$ufs
+  raw.data$phenotype = raw.data$Etanh
   raw.data$cofactor = raw.data$length
   Analysis1(phylo.vcv, raw.data)
 
@@ -608,7 +608,7 @@ main = function() {
   readline(prompt = 'Esec vs. length: press enter to continue.')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
-  writeLines('Comparison 4a: Esec low vs. length')
+  writeLines('Comparison 4c: Esec vs. length')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
   raw.data$phenotype = raw.data$uts / raw.data$ufs
@@ -727,7 +727,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$Esec)
   raw.data$cofactor = log(raw.data$length)
-  comparison = c('ln(Secant Modulus [MPa]', 'ln(Rostrum Length [mm2])')
[…]

### 2019-02-08 analysis/phylo_regress_bayesian.R

@@ -1,332 +1,0 @@
-# Andrew Jansen
-# created iv.15.2018
-
-# Miscellaneous Functions ------------------------------------------------------
-#Import the data
-GetData = function(file.name) {
-  test.data = data.frame(read.csv(file.name, sep = ','))
-  return(test.data)
-}
-
-#For ANOVA
-SqDev = function(x) {var(x) * (length(x) - 1)}
-
-# Analysis 1 -------------------------------------------------------------------
-# Compute co-variance matrix from inverse
-Analysis1 = function(phylo) {
-  inv.phylo = MCMCglmm::inverseA(phylo, nodes="TIPS")
-  A = solve(inv.phylo$Ainv)
-  rownames(A) = rownames(inv.phylo$Ainv)
-  print(A)
-  return(A)
-}
-
-# Analysis 2 -------------------------------------------------------------------
-# fit hierarchical phylogenetic model
-Analysis2 = function(raw.data, A, priors) {
-  phylo_model = brm(
-    phen ~ spec_mean_cf + (1|phylo) + (1|species),
-    data = raw.data,
-    family = gaussian(),
-    cov_ranef = list(phylo = A),
-    prior = priors,
-    sample_prior = TRUE#,
-    #chains = 16,
-    #cores = 8,
-    #iter = 2000,
-    #warmup = 1000
-  )
-  print(summary(phylo_model))
-  plot(phylo_model)
-  gg = plot(marginal_effects(phylo_model))[[1]]
-  plot(gg + geom_point(inherit.aes = FALSE,
-                  data = raw.data,
-                  mapping = aes(x = spec_mean_cf, y = phen, color = phylo)))
-  
-  return(phylo_model)
-}
-
-# Analysis 3 -------------------------------------------------------------------
-# Hypothesis testing and phylogenetic signal
-Analysis3 = function(phylo_model) {
-  hyp = paste("sd_phylo__Intercept^2 /",
-              "(sd_phylo__Intercept^2 + sd_species__Intercept^2 + sigma^2) = 0")
-  hyp = hypothesis(phylo_model, hyp, class = NULL)
-  
-  print(hyp)
-  plot(hyp)
-}
-
-# Analysis 4 -------------------------------------------------------------------
-# Update model and re-run phylogenetic signal
[…]

### 2019-02-08 analysis/phylo_regress_messy.R

@@ -1,316 +1,0 @@
-# Andrew Jansen
-# created iv.15.2018
-
-# Miscellaneous Functions ------------------------------------------------------
-#Import the data
-GetData = function(file.name) {
-  test.data = data.frame(read.csv(file.name, sep = ','))
-  return(test.data)
-}
-
-#For ANOVA
-SqDev = function(x) {var(x) * (length(x) - 1)}
-
-# Analysis 1 -------------------------------------------------------------------
-# ADD NAMES FOR LABELS ON GRAPHS TO FUNCTION INPUT
-Analysis1 = function(phylo, raw.data, axis.labels) {
-  
-  lm_mod = lme(phen ~ cf,
-               random = ~ 1|species,
-               data = raw.data,
-               correlation = corBlomberg(6, phylo, fixed = TRUE),
-               method = 'ML')
-  
-  lm_mod_alt = lme(phen ~ cf,
-                   random = ~ 1|species,
-                   data = raw.data,
-                   method = 'ML')
-  
-  print(summary(lm_mod))
-  print(summary(lm_mod_alt))
-  print(anova(lm_mod))
-  print(anova(lm_mod_alt))
-  
-  print(lrtest(lm_mod, lm_mod_alt))
-  
-  #create data.frame with new values for predictors
-  #more than one predictor is possible
-  cf = raw.data$cf
-  new.dat <- data.frame(cf)
-  
-  #predict response
-  new.dat$pred <- predict(lm_mod, newdata=new.dat, level=0)
-  
-  #create design matrix
-  Designmat <- model.matrix(eval(eval(lm_mod$call$fixed)[-2]),
-                            new.dat[-ncol(new.dat)])
-  
-  #compute standard error for predictions
-  predvar = diag(Designmat %*% lm_mod$varFix %*% t(Designmat))
-  new.dat$SE = sqrt(predvar) 
-  new.dat$SE2 = sqrt(predvar+lm_mod$sigma^2)
-  
-  gg = ggplot(new.dat,aes(x=cf,y=pred)) + 
-    geom_ribbon(aes(ymin=pred-2*SE2,ymax=pred+2*SE2),alpha=0.2,fill='black') +
-    geom_ribbon(aes(ymin=pred-2*SE,ymax=pred+2*SE),alpha=0.2,fill='black') +
[…]

### 2019-02-08 analysis/tensile.py

@@ -50,16 +50,29 @@ class specimen:
 
 
         self.__Eindex__ = []
-        lower = 0.00 #self.__ufs__ * 0
-        upper = 0.015 #self.__ufs__ * 0.20
+        lower = self.__ufs__ * 0
+        upper = self.__ufs__ * 0.33
         self.__Eindex__.append(len([x for x in takewhile(lambda x: x[1] <= lower, enumerate(self.__epsilonN__))]))
         self.__Eindex__.append(len([y for y in takewhile(lambda y: y[1] <= upper, enumerate(self.__epsilonN__))]))
         xvals = self.__epsilonN__[self.__Eindex__[0]:self.__Eindex__[1]]
         yvals = self.__sigmaN__[self.__Eindex__[0]:self.__Eindex__[1]]
         self.__E__ = linreg(xvals, yvals) #engineering stress and strain via Hooke's Law
 
-        self.__Etanl__ = self.__E__
+        self.__Etanl__ = self.__E__[0]
         self.__Etanh__ = 0
+
+        t_upper = 0
+        t_lower = 0
+
+        for i in range(len(self.__epsilonN__)-1):
+            ub = self.__epsilonN__[i+1]
+            lb = self.__epsilonN__[i]
+            uh = self.__sigmaN__[i+1]
+            lh = self.__sigmaN__[i]
+            t_upper += (ub - lb) * uh
+            t_lower += (ub - lb) * lh
+
+        self.__U__ = (t_upper + t_lower) / 2
 
     def asuid(self):
         return self.__asuid__
@@ -164,6 +177,9 @@ class specimen:
     def ufs(self):
         return self.__ufs__
 
+    def U(self):
+        return self.__U__
+
 def as_si(x, ndp):
     s = '{x:0.{ndp:d}e}'.format(x=x, ndp=ndp)
     m, e = s.split('e')
@@ -240,8 +256,15 @@ def stress_strain(spec_set, save=True, i=0, j=-1):
         ax3.set_xlabel('Strain (ln(L/L0) or dL/L0)')
 
         print(spec.asuid()+' '+spec.species())
-        new_l = spec.ufs() * 0.65
-        new_u = spec.ufs() * 0.85
+        new_l = spec.ufs() * 0.00
+        new_u = spec.ufs() * 0.33
+        spec.recalcE(lower = new_l, upper = new_u, graphical=False, normal=False)
+        ax3.plot(spec.epsilonN()[spec.Eindex(0):spec.Eindex(1)],
[…]

### 2019-02-11 tex_v2/manuscript.tex
all but conclusions and methods

@@
 
-%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, present in all examined species, and likely across the genus (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich gular sutures, which would be under tension during the drilling behavior in the live insect.
+%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, present in all examined species, and likely across the genus (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich occipital sulci, which would be under tension during the drilling behavior in the live insect.
 
@@
 }
-\section*{Introduction} \info{650 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
-%Intro outline - 500 word maximum
-%	Paragraph 1 - Cuticle of beetles
-%	Paragraph 2 - Weevils and behavior
-%	Paragraph 3 - modified cuticle in curculio
-%	Paragraph 4 - study objectives
-%	Paragraph 5 - short segue/ signposting
+
+\section*{Introduction} \info{670 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
+%Intro - 670 words, reduce to 500 or so
 	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
-	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its uniquely layered microstructure\unsure{give more numbers here}.
+	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure\unsure{give more numbers here}.
 	Beetle cuticle is divided into two structurally and mechanically distinct regions, the (outer) exocuticle and (inner) endocuticle.
@@
 	Despite being composed of the same material as other rigid body parts, the snout can be repeatedly bent without evident damage.
-	This structure is used by the female to feed and excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
+	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
 	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit; a single female may prepare hundreds of such sites during the adult stage.
@@
 	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
-	Based on our findings we posit that the composite profile of the rostral apex enables the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	Based on our findings, we posit that the composite profile of the rostral apex enables the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
 	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
@@
 	\section{Microstructure of the \textit{Curculio} rostrum}
-	\info{850 words, reduce in revisions}
+	\info{780 words, reduce in revisions}
 		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
@@
 	\section{Force-controlled loading to fracture}
-	%550 words, maybe reduce?? Still must add numbers...
+	%600 words... still must add numbers...
 
-		To better characterize the mechanical behavior of the snout as an integrated whole, we performed tensile testing on the snouts of six species in the genus \textit{Curculio}, representing a mixture of closely and distantly related taxa.
-		Although the heads were rehydrated by immersion in de-ionized water for 24 hours, we observed comparatively brittle fracture (but see below), in contrast to GIVE EXAMPLES HERE.
+		To better characterize the mechanical behavior of the rostrum at failure, we performed tensile testing on the snouts of six \textit{Curculio} species, representing a mixture of closely and distantly related taxa.
+		Heads of female specimens of each species were rehydrated by immersion in de-ionized water for 24 hours
+		Each specimen was subjected to force-controlled, uniaxial loading to fracture at a constant stress rate (detailed in Methods).
+		We then examined the correspondence between composite structure and mechanical behavior of the snout, using phylogenetic linear mixed effects models to account for phylogenetic non-independence in the residual variances of the models.
 
-		In general, the behavior of the rostrum is non-linear viscoelastic, and, as in other biological materials, exhibits a stress-strain curve characterized by a steady increase in stiffness, beginning with a toe-in region and terminating in apparently brittle fracture after a period of strain hardening.
-		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle.
-		Consequently, there was a negative correlation between utlimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
-		In other words, UTS is positively correlated with the relative proportion of the endocuticle across species.
+		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in apparently brittle fracture after a period of strain hardening.
+		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
+		Consequently, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
+		Inversely, the failure strength of the cuticle is positively associated with the relative proportion of the endocuticle.
 
-		Although the proportion of exocuticle is predicted to be positively correlated with the stiffness of the resulting composite, we found no such associations because of the anterior-posterior gradient in cuticle profile.
-		Because the cross-sectional areas of the cuticle regions vary across the length of the head, it is meaningless to attempt to correlate these areas to properties of the entire snout.
-		Instead, we found that the uniaxial elastic moduli of the snout under low strain (E-low) was inversely correlated with the ultimate strain and toughness of the snout (as was the secant modulus, E-sec).
-		These associations were found to be statistically significant and independent of species membership, rostrum length, and location on the snout.
+		Although CLPT predicts that the proportion of exocuticle is positively correlated with the stiffness of generalized cuticle, we found no associations between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
+		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.
+		Instead, we found that the uniaxial elastic modulus of the snout under low strain (E-low) was inversely correlated with the ultimate strain and toughness of the snout (as was the secant modulus, E-sec)
+		These associations were all found to be statistically significant and independent of species membership.
 
-		From our results, we infer that there is little variation in the gross elastic behavior of the cuticle across the genus, in agreement with our current understanding of cuticle mechanobiology, making it unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum (singh 2016, jansen 2016).
+		Additionally, we observed a moderate stiffening size-effect with respect to rostral length.
+		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) snout would need to be more flexible to avoid fracture during oviposition.
+		Based on preliminary confocal microscopy data, we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
+		We hypothesize that a decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
+
+		From these results, we infer that the gross elastic behavior of the cuticle is consistent across the genus, in agreement with our current understanding of cuticle mechanobiology.
+		It is therefore unlikely that differences in sclerotization or chitin composition within the cuticle, rather than composite profile, are responsible for the mechanical behavior of the rostrum in different species.
 		Furthermore, the endocuticle apparently contributes more to the tensile strength of the rostrum than the exocuticle, likely because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
-		Thus, in addition to making the cuticle more flexible, the altered apical composite profile makes the rostrum less likely to break, suggesting a possible means by which this system evolved\info{(via negative selection of breakage, maybe say this in conclusions)}.
-
-		\info{DO I really want to discuss this? Occupies precious space...}Finally, we observed that the elastic modulus is higher in specimens with a longer snout.
-		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) would need to be more flexible to avoid fracture during oviposition.
-		Based on preliminary confocal microscopy data, we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
-		A decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
+		Thus, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.
 
 	\section{Load cycling of \textit{Curculio caryae}}
-	%375 words, add no more than 100, ONLY IF NEEDED
-		The shearing motion and accommodation of strain led us to question how the cuticle might accommodate repeated strain, as is seen in the living organism.
-		We therefore performed fatigue testing on a female curculio longinasus, which exhibits the most extreme degree of bending in the sis species examined.
+	%480 words
+		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature.
+		The head was removed and rehydrated for 24 hours in de-ionized water, then coated in grease to prevent loss of moisture and stiffening of the specimen during the duration of the test.
+		The specimen was then aligned so that uniaxial tension would induce elongation of the apical portion of the rostrum.
+		The strain per cycle was fixed at an amplitude sufficient to completely elongate the snout and generate a tensile load equal to $0.2\times$ the average tensile strength of conspecifics in the straightened configuration, at a frequency of $0.33\,\text{Hz}$.
+		The test was terminated after a period of two weeks (ca.~400K cycles) when the strain amplitude appeared to reach an asymptotic minimum.
 
-		~400K cycles, complete elongation, rehydrated, coated in grease to prevent loss of moisture and stiffening of the specimen.
-		We observed viscoelastic behavior in the specimen, as indicated by histeresis in the stress-strain relationship during each cycle.
-		Fmax decreased logarithmically with cycle number, etc., and the specimen appeared to have deformed plastically during the test.
-		We initially believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that the specimen returned to its original shape.
-		The specimen did not show any evidence of fractures or shear cusps anywhere in the surface of the exocuticle, and, furthermore, the tensile strength of the specimen was consistent with other members of its species.
-		Given this surprising result, it appears that the specimen was undamaged by the testing.
-		We therefore believe that under normal conditions in life, repeated bending of the snout does exceed the yield/plastic limit of the cuticle, and the bending strain is purely elastic/viscoelastic.
+		We observed viscoelastic behavior in the rostrum, as indicated by histeresis in the stress-strain relationship during each cycle.
+		Strain amplitude decreased logarithmically with cycle number, and the specimen initially appeared to have deformed plastically during the test.
+		We believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that rostrum returned to its original shape.
 
-		We cannot fully account for these results, but we speculate that the microstructure of the endocuticle is responsible for what we observed.
+		We cannot fully account for the stress relaxation of the rostrum after testing, but we speculate that it arose from the same mechanism as the viscoelastic behavior observed in tensile testing.
 		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains along their length.
 		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
-		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test, but without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform viscoelastically/viscoplastically.
-		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain without any damage to the specimen.
+		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test; however, without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform viscoelastically (REF\info{stress history dependence paper}).
+		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain.
+
+		The specimen did not show any evidence of fractures, micro-tears, or shear cusps anywhere in the surface of the exocuticle, and, furthermore, the tensile strength of the specimen was consistent with other members of its species ($F_max~=~5.02\,\text{N}$).
+		Given this surprising result, it appears that the specimen was undamaged by the testing.
+		We therefore expect that under normal conditions in life, repeated bending of the snout does not exceed the yield strength of the cuticle, and the bending strain is purely elastic.
 
 	\section{Fractography of test specimens}
-	%currently ~400 words, keep to this length or reduce to 2 short paragraphs
-		Examination of the fracture surfaces and adjacent cuticle of tensile testing specimens revealed that although fracture was comparatively brittle, the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure, in agreement with previous studies (name some).
+	%currently ~650 words
+		Due to the complex failure patterns present across the fracture surface, it was not always possible to identify void nucleation and crack initiation sites (both primary and secondary).
+		Nonetheless, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here.
 
-		In cross-section, the exocuticle consistently presented a smooth, nearly continuous fracture surface, indicative of relatively brittle fracture, likely due to the microarchitecture and resultant transverse isotropy of the Bouligand structure (explain this).
-		Endocuticle, on the other hand exhibited severe delamination, ply-splitting, and fiber pulling, consistent with viscoelastic/plastic behavior shown in previous studies and congruent with theoretical consideration of the microstructure of this material (name something here).
-		These patterns indicate that the endocuticle is probably less brittle than the exocuticle, most likely due to the alignment of the \textalpha-chitin fibers in each macrofiber.
+		Scanning electron and light microscopy of testing specimens revealed that the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure.	
+		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, indicative of comparatively brittle failure, likely due to the helicoidal arrangement of \textalpha-chitin laminae and resultant transverse isotropy of the Bouligand structure.
+		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident across the fracture surface and over unbroken endocuticle on the rostral surface adjacent to the plane of fracture.
+		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating that endocuticular macrofibers are less brittle than the exocuticle due to the unidirectional alignment of their \textalpha-chitin fibers.
+		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers, the presence of exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, which caused ply-splitting via mode II fracture between macrofibers at high strain.
+		This intra-laminar extension-shear coupling also yielded off-axis in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing inter-ply delamination.
+		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae.
 
-		In addition, the exocuticle typically appears to fracture before the endocuticle, with shear-cusp formation evident over unbroken endocuticle.
-		We note, however, that the exocuticle of weevils/beetles is anchored to the endocuticle by cross-linking fibers in a transition zone described by Kamp et al. (see refs.).
-		The presence of shear-cusps therefore indicates that the fibers of the endocuticle are shearing past each other within each ply along the radial/normal plane (type II shearing???, extension shear-coupling).
-		Furthermore, given the delamination observed between plies, the layers of endocuticle are likely shearing past each other (type III shearing) along the transverse plane.
-		We therefore infer that, under tension, the endocuticle tends to deform viscoelastically and plastically along the longitudinal axes of the macrofibers, while the overlying exocuticle exhibits brittle fracture due to shearing between the stretching endocuticle fibers to which it is anchored.
+		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum.
+		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and create large interfaces ideal for void nucleation.
+		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, continuing dorsolaterally and ventromedially.
+		Ventrally the cusps converged toward a prominent scarp where the crack fronts joined, although in specimens with large cross-sectional areas of endocuticle, this was often obscured by delaminated endocuticular macrofibers.
 
-		Additionally, the fracture surfaces show a characteristic failure mode, based on the pattern of fiber dislocation in the plies of the endocuticle.
-		(copy description from DATA PRESENTATION, for figure 5B).
+		The first layer of endocuticle usually fractured along the same plane as the exocuticle, and ventrally the endocuticle laminae typically converged toward a scarp-like region characterized by severe delamination and numerous de-bonded macrofibers.
+		Additionally, macrofibers oriented in the direction of crack propagation exhibited extensive ply-splitting with intermittent transverse shearing, while macrofibers oriented against the direction of crack propagation primarily displayed fracture by transverse shear along the plane of ply-splitting in adjacent laminae.
+		Thus, contralateral fibers in the same lamina displayed opposing fracture modes.
+		In addition, the ventrolateral surfaces often exhibited extensive inter-ply delamination and fiber de-bonding in scarp-like prominences, likely due to a combination of tensile failure and shearing along the dorsally-radiating crack front.
+		Dorsally, the coalescent crack fronts often caused significant de-bonding and ply-splitting followed by broom-like tensile failure; in some specimens, the contralateral crack fronts were out of plane and coalesced via transverse shearing through a large dorsal section of cuticle.
 
-		From this pattern we hypothesize that the exocuticle-rich gular sutures are the most likely site for the initiation of void nucleation and failure of the integrated rostral cuticle in cross section.
-		Structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
-		We speculate that this could be the reason why the cross-sectional profile of \textit{C. caryae} is flattened ventrally.
-		Flattening this region may reduce tensile-strain across the gular sutures when the snout is bent dorsally, thus reducing the risk of fracture in the elongate, strongly-curved rostrum in this species.
+		From these patterns we hypothesize that the exocuticle-rich occipital sulci are the most likely site for the initiation of void nucleation and catastrophic failure of the integrated rostral cuticle in cross section, as illustrated in Fig~REF.
+		According to this model, structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
+		Although other, more complex failure modes have been observed, we posit that in live specimens this is the most likely mechanism of tensile failure because typical bending behavior generates tension along the ventral surface of the rostrum.
 
 	\section{Conclusions}
-	%1-2 'short paragraphs', 300 words maximum
+	%1-2 'short paragraphs', 300 words maximum, already at 3200...
+		In this study, we...
+		raising the intriguing possibility that this system evolved via negative selection of breakage, rather than positive selection of flexibility.\info{(maybe say this in conclusions)}
+
+		evolutionary constraint (like occipital sulci)
+
+		The macrofibers are pulled longitudinal axis, explains fmax, toughness, and upcurved stress-strain curve.
+
+		We therefore infer that, under uniaxial tension, the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers, causing ply-splitting and delamination at high strain, while the overlying exocuticle exhibits brittle fracture, at a comparatively low strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
+		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
 
@@
 		\subsection*{Specimen imaging and microscopy}
+			Cold-fracture of rostra
 

### 2019-02-11 analysis/tensile.py

@@ -314,8 +314,8 @@ def xyplot(spec_set, save = True):
         new_l = spec.ufs() * 0.67
         new_u = spec.ufs() * 1.00
         spec.recalcE(lower = new_l, upper = new_u, graphical=False, normal=False)
-        x = ln(spec.ufs())
-        y = ln(spec.U())
+        x = ln(spec.endo() / spec.exo())
+        y = ln(spec.ufs())
         #spec.uts() / spec.ufs()
 
         if spec.species() == 'caryae':

### 2019-02-11 tex_v2/manuscript.tex
edits and begin conclusions

@@
 	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
-	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we estimate the effect of these differences using Classical Laminate Plate Theory (CLPT).
+	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we estimate the effect of these differences on the elasticity of the cuticle using Classical Laminate Plate Theory (CLPT).
 
-	Because recent studies have shown that the yield strength of the weevil rostrum exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
+	Because recent studies have shown that the yield strength of the beetle exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
 	We relate an observed increase in the volume fraction of endocuticle in the rostrum to higher tensile strength at the rostral apex in all tested species, and find that a strongly curved rostrum can be flexed repeatedly without harm to the structure.
@@
 	\info{780 words, reduce in revisions}
-		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin fibers; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
+		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
 		These layers are stacked at a more or less constant angle to each other, thus forming a quasi-isotropic laminate referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}.
@@
 
-		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens has revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
+		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens have revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
 		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle\info{thickness measurements here}.
-		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}) in two key characteristics:
-
-		\begin{enumerate}
-			\item The exocuticle is reduced to a thin shell\info{measurement here}, with the endocuticle thickened \info{measurement here} to offset this reduction to maintain a constant cuticle thickness in the head.
-			\item The endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
-		\end{enumerate}
+		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}).
+		Distally the exocuticle is reduced to a thin shell\info{measurement here}, with the endocuticle thickened \info{measurement here} to offset this reduction and maintain a constant cuticle thickness in the head.
+		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
 
@@
 
-		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications, namely:
-		\begin{enumerate}
-			\item A hybrid cuticle with the angle stacking sequence of regular cuticle (the head capsule) but possessing the layer thicknesses of the rostral apex.
-$E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases, respectively.
-			\item A hybrid cuticle with the layer thicknesses of regular cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle).
-$E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases, respectively.
-		\end{enumerate}
+		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications.
+		A hybrid configuration with the angle stacking sequence of typical cuticle (i.e., in the head capsule) but possessing the layer thicknesses of the rostral apex had effective moduli $E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases from unmodified cuticle, respectively.
+		Similarly, a hypothetical cuticle with the layer thicknesses of ordinary cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle) exhibited elastic moduli $E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases from unmodified cuticle, respectively.
 
@@
 	\section{Force-controlled loading to fracture}
-	%600 words... still must add numbers...
+	%590 words... still must add numbers...
 
 		To better characterize the mechanical behavior of the rostrum at failure, we performed tensile testing on the snouts of six \textit{Curculio} species, representing a mixture of closely and distantly related taxa.
-		Heads of female specimens of each species were rehydrated by immersion in de-ionized water for 24 hours
-		Each specimen was subjected to force-controlled, uniaxial loading to fracture at a constant stress rate (detailed in Methods).
+		Heads of female specimens of each species were rehydrated by immersion in de-ionized water for 24 hours to simulate the condition of the living tissue (but see Klocke and Schmitz 2009).
+		Each specimen was subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $0.05\,\text{N}\cdot\text{s}^{-1}$ (detailed in Methods).
 		We then examined the correspondence between composite structure and mechanical behavior of the snout, using phylogenetic linear mixed effects models to account for phylogenetic non-independence in the residual variances of the models.
@@
 		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in apparently brittle fracture after a period of strain hardening.
-		We observed that the maximum force sustained at the site of the break was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
-		Consequently, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
-		Inversely, the failure strength of the cuticle is positively associated with the relative proportion of the endocuticle.
+		We believe that strain hardening prior to fracture is a consequence of stiff longitudinal axis of the macrofibers becoming more closely aligned to the longitudinal axis of the rostrum, thereby resisting tension more directly at higher strains.
+		We observed that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
+		Consequently, there was a positive correlation between ultimate tensile strength of the specimen and the ratio of endocuticle to exocuticle cross-sectional area at the site of fracture.
 
-		Although CLPT predicts that the proportion of exocuticle is positively correlated with the stiffness of generalized cuticle, we found no associations between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
+		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
 		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.
@@
 
-		Additionally, we observed a moderate stiffening size-effect with respect to rostral length.
+		Additionally, we observed a moderate, stiffening size-effect with respect to rostral length.
 		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) snout would need to be more flexible to avoid fracture during oviposition.
@@
 
-		From these results, we infer that the gross elastic behavior of the cuticle is consistent across the genus, in agreement with our current understanding of cuticle mechanobiology.
-		It is therefore unlikely that differences in sclerotization or chitin composition within the cuticle, rather than composite profile, are responsible for the mechanical behavior of the rostrum in different species.
-		Furthermore, the endocuticle apparently contributes more to the tensile strength of the rostrum than the exocuticle, likely because the endocuticle is organized in large bundles of aligned, anisotropic fibers.
-		Thus, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.
+		From these tests, we infer that the gross elastic behavior of the cuticle is consistent across the genus, in agreement with our current understanding of cuticle mechanobiology.
+		It is therefore unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum in different species; instead, a single mechanism (i.e.
+the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
+		he endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers.
+		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.
 
@@
 	%currently ~650 words
-		Due to the complex failure patterns present across the fracture surface, it was not always possible to identify void nucleation and crack initiation sites (both primary and secondary).
+		Following tensile testing, the fractured specimens were subjected to fractographic analysis using light and scanning electron microscopy.
+		Due to the complex failure modes evident upon examination of the fractured specimens, it was not always possible to identify void nucleation and crack initiation sites (both primary and secondary).
 		Nonetheless, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here.
 
-		Scanning electron and light microscopy of testing specimens revealed that the fracture mechanics of the exocuticle and endocuticle differed according to their microstructure.	
-		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, indicative of comparatively brittle failure, likely due to the helicoidal arrangement of \textalpha-chitin laminae and resultant transverse isotropy of the Bouligand structure.
-		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident across the fracture surface and over unbroken endocuticle on the rostral surface adjacent to the plane of fracture.
-		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating that endocuticular macrofibers are less brittle than the exocuticle due to the unidirectional alignment of their \textalpha-chitin fibers.
+		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, indicative of comparatively brittle failure, due to the helicoidal arrangement of \textalpha-chitin laminae and resultant transverse isotropy of the Bouligand structure (REF).
+		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across unbroken endocuticle on the rostral surface adjacent to the plane of fracture.
+		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating that endocuticular macrofibers are less brittle than the exocuticle, likely as a consequence of the unidirectional alignment of their \textalpha-chitin nanofibrils (REF).
 		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers, the presence of exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, which caused ply-splitting via mode II fracture between macrofibers at high strain.
-		This intra-laminar extension-shear coupling also yielded off-axis in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing inter-ply delamination.
+		This intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing inter-ply delamination.
 		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae.
@@
 		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum.
-		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and create large interfaces ideal for void nucleation.
+		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation.
 		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, continuing dorsolaterally and ventromedially.
@@
 		Additionally, macrofibers oriented in the direction of crack propagation exhibited extensive ply-splitting with intermittent transverse shearing, while macrofibers oriented against the direction of crack propagation primarily displayed fracture by transverse shear along the plane of ply-splitting in adjacent laminae.
-		Thus, contralateral fibers in the same lamina displayed opposing fracture modes.
+		Because the laminae form a cylinder, contralateral fibers in the same lamina display opposing fracture modes.
 		In addition, the ventrolateral surfaces often exhibited extensive inter-ply delamination and fiber de-bonding in scarp-like prominences, likely due to a combination of tensile failure and shearing along the dorsally-radiating crack front.
-		Dorsally, the coalescent crack fronts often caused significant de-bonding and ply-splitting followed by broom-like tensile failure; in some specimens, the contralateral crack fronts were out of plane and coalesced via transverse shearing through a large dorsal section of cuticle.
+		Dorsally, the coalescent crack fronts often caused significant de-bonding and ply-splitting followed by broom-like tensile failure; in some specimens, the contralateral crack fronts were out of plane and coalesced via transverse shear through a large dorsal section of cuticle.
 
@@
 		According to this model, structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
-		Although other, more complex failure modes have been observed, we posit that in live specimens this is the most likely mechanism of tensile failure because typical bending behavior generates tension along the ventral surface of the rostrum.
+		Although other, more complex failure modes have been observed, we posit that in live specimens this is the most likely mechanism of tensile failure because typical bending behavior generates tension only along the ventral surface of the rostrum.
 
 	\section{Conclusions}
-	%1-2 'short paragraphs', 300 words maximum, already at 3200...
-		In this study, we...
-		raising the intriguing possibility that this system evolved via negative selection of breakage, rather than positive selection of flexibility.\info{(maybe say this in conclusions)}
+	%word count: (3157 so far...)
+	%Par 1 - Microstructure
+	%Par 2 - Tensile and fatigue
+	%Par 3 - Fracture
+	%Par 4 - Wrap-up
+		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile, wherein the cuticle is strongly differentiated along an anterior-posterior gradient.
+		In particular, the cuticle in the apical portion of the rostrum exhibits marked reduction in relative thickness of the and in stacking-angle between pseudo-orthogonal plies of the endocuticle, which are aligned at $\pm45\degree$ to the longitudinal axis of the snout.
+		Constitutive models demonstrated that an idealized apical cuticle could theoretically achieve 72\% and 60\% reductions in the effective longitudinal membrane and flexural moduli, respectively, compared to the cuticle of the head capsule, thus accounting for the observed flexibility of the rostral apex in live specimens.
+However, these reductions could only be realized with both modifications to the cuticle, which have a non-additive effect on cuticle elasticity.
+		Tensile testing data revealed that increased endocuticle thickness enhances the fracture toughness of the cuticle, raising the intriguing possibility that this system evolved via negative selection of breakage, rather than positive selection of flexibility.
 
-		evolutionary constraint (like occipital sulci)
+		We confirmed that altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, and .
 
-		The macrofibers are pulled longitudinal axis, explains fmax, toughness, and upcurved stress-strain curve.
+		Based on fractographic analysis of test-specimens, we infer that the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers under uniaxial tension, causing ply-splitting and delamination at high strain.
+		The overlying exocuticle then exhibits brittle fracture, at a comparatively low strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
+		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
 
-		We therefore infer that, under uniaxial tension, the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers, causing ply-splitting and delamination at high strain, while the overlying exocuticle exhibits brittle fracture, at a comparatively low strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
-		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
+		To the best of our knowledge... had not previously been reported
 

### 2019-02-11 tex_v2/manuscript.tex
conclusions done

@@
 	\section{Conclusions}
-	%word count: (3157 so far...)
-	%Par 1 - Microstructure
-	%Par 2 - Tensile and fatigue
-	%Par 3 - Fracture
-	%Par 4 - Wrap-up
+	%word count: (3176 so far...)
+
 		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile, wherein the cuticle is strongly differentiated along an anterior-posterior gradient.
@@
 		Constitutive models demonstrated that an idealized apical cuticle could theoretically achieve 72\% and 60\% reductions in the effective longitudinal membrane and flexural moduli, respectively, compared to the cuticle of the head capsule, thus accounting for the observed flexibility of the rostral apex in live specimens.
-However, these reductions could only be realized with both modifications to the cuticle, which have a non-additive effect on cuticle elasticity.
-		Tensile testing data revealed that increased endocuticle thickness enhances the fracture toughness of the cuticle, raising the intriguing possibility that this system evolved via negative selection of breakage, rather than positive selection of flexibility.
+		However, these reductions can only be realized with both modifications to the cuticle, which have a non-additive effect on cuticle elasticity, implying that these weevils require both modifications to function properly during oviposition.
 
-		We confirmed that altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, and .
+		Our tensile testing data revealed that increased endocuticle thickness correlates with greater fracture strength of the cuticle, while increased stiffness was associated with lower failure strains and reduced fracture toughness.
+		Additionally, fatigue testing of a highly-curved rostrum confirmed that the apical portion of the snout can be subjected to repeated antero-dorsal flexion without evidence of damage.
+		We therefore conclude that the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, and that this profile allows the rostrum to be flexed into a completely straight configuration without exceeding the elastic limits of the cuticle.
+		Because these associations were independent of species membership, we posit that the behavior of the cuticle is consistent across the genus, and that rostral flexibility is achieved exclusively through a modified cuticle lay-up in all \textit{Curculio} species.
 
-		Based on fractographic analysis of test-specimens, we infer that the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers under uniaxial tension, causing ply-splitting and delamination at high strain.
-		The overlying exocuticle then exhibits brittle fracture, at a comparatively low strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
+		Based on fractographic analysis of the test-specimens, we infer that the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers under uniaxial tension, causing ply-splitting and delamination at high strain.
+		However, the overlying exocuticle exhibits brittle fracture at a comparatively lower strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
 		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
+		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
 
-		To the best of our knowledge... had not previously been reported
+		From a biological perspective, these findings reveal an unexpected morphological source of evolutionary constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
+		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion; the doubly-thick exocuticle in the invagination thus creates an unavoidable, brittle weak-point in the otherwise endocuticle-dominated rostral apex.
+		This constraint, as well as the minimization of exocuticle thickness in the rostral apex, and the increased toughness derived from a thickened endocuticle, all point to avoidance of catastrophic structural failure as a driving selective pressure in the evolution of the rostrum.
 

### 2019-02-11 tex_v2/manuscript.tex
word count: ~3700
need to CUT

@@
 \section*{Introduction} \info{670 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
-%Intro - 670 words, reduce to 500 or so
+%word count: 673
 	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
@@
 	\section{Microstructure of the \textit{Curculio} rostrum}
-	\info{780 words, reduce in revisions}
+	%word count: 784
 		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
@@
 	\section{Force-controlled loading to fracture}
-	%590 words... still must add numbers...
-
+	%word count: 608
 		To better characterize the mechanical behavior of the rostrum at failure, we performed tensile testing on the snouts of six \textit{Curculio} species, representing a mixture of closely and distantly related taxa.
@@
 	\section{Load cycling of \textit{Curculio caryae}}
-	%480 words
+	%word count: 474
 		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature.
@@
 	\section{Fractography of test specimens}
-	%currently ~650 words
-		Following tensile testing, the fractured specimens were subjected to fractographic analysis using light and scanning electron microscopy.
+	%word count: 653
 		Due to the complex failure modes evident upon examination of the fractured specimens, it was not always possible to identify void nucleation and crack initiation sites (both primary and secondary).
@@
 	\section{Conclusions}
-	%word count: (3176 so far...)
+	%word count: 514 (3706 total)
 

### 2019-02-12 tex_v2/manuscript.tex
most of methods

@@
-\documentclass[twocolumn, linenumbers, superscriptaddress]{revtex4-1}
+\documentclass[twocolumn, linenumbers, superscriptaddress, nofootinbib]{revtex4-1}
 
@@
 	In general, impact-prone areas and exaggerated structures, such as horns and legs exhibit cuticle organization that resists deformation and fracture.
-	Acorn weevils in the genus \textit{Curculio} are typified by an elongate structure of the head, called the rostrum (snout), which instead exhibits unusual distal flexibility.
+	Acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
+	This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author.} are typified by an elongate structure of the head, called the rostrum (snout), which instead exhibits unusual distal flexibility.
 	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
@@
 
-		In previous work we identified these modifications to the composite structure of the cuticle within a single species, \textit{C. longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
+		In previous work we identified these modifications to the composite structure of the cuticle within a single species, \textit{C.~longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
 		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
@@
 		To investigate the effect of these cuticle modifications, we estimated uni-axial membrane and transverse flexural Young's moduli of the cuticle in both the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
-		We previously derived the effective elastic constants of the cuticle regions of \textit{C. longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
+		We previously derived the effective elastic constants of the cuticle regions of \textit{C.~longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
 		The cuticle was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively, in the head capsule; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
@@
 		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) snout would need to be more flexible to avoid fracture during oviposition.
-		Based on preliminary confocal microscopy data, we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
+		Based on preliminary confocal microscopy data (not presented herein), we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
 		We hypothesize that a decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
@@
 		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature.
-		The head was removed and rehydrated for 24 hours in de-ionized water, then coated in grease to prevent loss of moisture and stiffening of the specimen during the duration of the test.
 		The specimen was then aligned so that uniaxial tension would induce elongation of the apical portion of the rostrum.
@@
 		Reprints and permissions information is available online at www.nature.com/reprints.
-		Correspondence and requests for materials should be addressed to M.A.J.
+		Correspondence and requests for materials should be addressed to Andrew Jansen.
 
@@
 	\section*{Methods}
+		\subsection*{Specimen acquisition and taxon sampling}
+			Specimens for use in tensile and fatigue testing came from the Hasbrouck Insect Collection at Arizona State University [\textbf{\texttt{ASUC}}].
+This set of specimens was supplemented with material housed in the following collections, using the codens of Arnett	et al.
+(1993)(REF):
+
+			\begin{description}
+				\item [\texttt{CMNC}]  Canadian Museum of Nature Collection, Ottawa, Ontario, Canada
+				\item [\texttt{USNM}]  National Museum of Natural History, Washington, D.C., USA
+			\end{description}
+
+			Cold fracture, semi-thin sectioning, and tensile testing were conducted on randomly chosen female specimens belonging to six species obtainable through field work in the southwestern USA and northwestern Mexico.
+			Taxon sampling was targeted to represent a mixture of disparate radiations and sister taxa with a variety of rostral morphotypes, according to the phylogenetic hypotheses of authors(REF \& REF).
+			These six species of \textit{Curculio} used herein are \textit{C.~caryae} (Horn, 1873), \textit{C.~humeralis} (Casey, 1897), \textit{C.~proboscideus} Fabricius, 1775, \textit{C.~sulcatulus} (Casey, 1897), \textit{C.~uniformis} (LeConte, 1857), and \textit{C.~victoriensis} (Chittenden, 1904).
+
 		\subsection*{Histological sectioning}
+			To illustrate the relative proportions of the cuticle regions in cross-section, serial semi-thin sectioning was conducted on exemplary female specimens of \textit{C.~humeralis} and \textit{C.~longinasus} Chittenden 1927.
+			Live specimens of both species were collected into 95\% ethanol for
+			preservation.
+			A female specimen was selected and the rostrum was separated from the head capsule with a fine-edged razor blade.
+			The apical 1/4th of	the rostrum was also removed and then discarded.
+			The remaining portion	of the rostrum and the head capsule were then embedded in EMbed812, as follows.
+
+			The cuticle was first immersed in acetone for 24 hours, and then transferred to a 2:1 mixture of acetone to epoxy resin.
+			Samples remained at $21\degree\text{C}$ for 12 hours on a shaker table to prevent hardening.
+			These were then transferred into a 1:1 mixture of acetone to resin, following by a 1:2 mixture (each for 12 hours and at $21\degree\text{C}$ on a shaker table), before finally being placed into a silicone mold with pure resin.
+			The mold was placed	into an oven heated to $38\degree\text{C}$, and the resin was allowed to cure for 24 hours.
+			The resulting blocks were machined to prepare the apical surface of each
+			sample for microtomy.
+
+			A Leica Ultracut R Microtome and diamond knife were used to expose a cross-section (transverse plane) of the apical and basal portions of the rostrum and to remove excess material.
+			Semi-thin sections (0.5\,{\textmu}m thick) were kept and stained with
+			toluidine-blue-borax for light microscopy and imaging.
+
+		\subsection*{Cold-fracture of specimens}
+			Two pinned female specimens of each species were selected at random and retained for cold-fracturing of the rostrum.
+			The heads of the specimens were removed and cleaned using a 95\% ethanol solution and a thin paintbrush.
+			Any muscles protruding from the occipital foramen were removed with a fine-edged razor blade.
+			The antennae were removed directly using forceps to pull the scape (basal section of antenna) from the antennal insertion.
+			Cleaned specimens were stored at $-80\degree\text{C}$ for 24 hours, then fractured using forceps over a chilled aluminum block.
+			To fracture each specimen, the head capsule and rostrum were each gripped firmly in a pair of forceps; the forceps were then sharply rotated to fracture the base of the snout via dorsal flexion.
+			The rostrum was fractured a second time, after separation from the head capsule, using the same procedure.
+			The segmented specimens were then placed into individual glass vials to protect the fracture surfaces from contamination prior to microscopy.
 
 		\subsection*{Tensile and fatigue testing}
+			\paragraph*{Force-controlled loading to failure}
+				Five female specimens of each species were randomly allocated for use in tensile testing.
+				The head of each specimen was removed, cleaned, and prepared as described above in the cold-fracture protocol.
+				To avoid destroying the delicate, brittle specimens when gripping the ends of each head, a method was devised to create solid handles that could be clamped tightly into grips without risk of damage to the cuticle.
+
+				For each head, four $1\,\text{cm}^2$ strips of gaffer tape were cut; these were used as gripping and mounting points for the specimen.
+				A strip of tape would be laid flat, with a large drop of cyanoacrylate glue placed upon the upturned surface.
+				The curved portion of the snout was then placed into the drop such that the straight portion of the rostrum was aligned perpendicular to the edge of the strip.
+				Hardening of the cyanoacrylate effectively embedded the curved portion of the snout in a solid mass, isolating a straight section of the snout (from the base to a point distad of the apex of the scrobe) for testing.
+				A second strip of tape was fixed over this mass with an additional layer of cyanoacrylate to provide a dorsal gripping surface for the mass, and a small mark was made to indicate the extent of the head inside the mass.
+				This embedding procedure was likewise repeated for the head capsule, resulting in a finished specimen with anterior and posterior handles for testing.
+
+				Prior to testing, each specimen was placed in de-ionized water for 24 hours to allow full saturation of the cuticle, simulating the condition of live tissue (see also REF).
+				Once removed for testing, the specimen was gripped using the cyanoacrylate handles at the marked locations immediately beyond the anterior margin of the rostrum and the posterior margin of the head capsule
+				The exposed section of the snout was coated in petroleum jelly using a cotton swab to prevent loss of moisture and stiffening of the specimen during the test.
+				Specimens were loaded in a Tryton 250 Microforce Testing System equipped with a 5N load cell and mechanical clamp grip.
+				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ until failure.
+				Results were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
+
+			\paragraph*{Displacement-controlled cyclic loading}
+				To confirm that repeated, complete extension of a strongly curved rostrum would not result in fracture of the cuticle, a representative female specimen of \textit{C.~caryae} was allocated for fatigue testing.
+				The head capsule of the specimen was fixed to a push-pin using cyanoacrylate glue; this served as a pedestal and gripping location for the posterior portion of the specimen.
+				The apex of the rostrum was fixed to a strip of ripstop nylon fabric equal in length to the head using cyanoacrylate glue.
+				As with tensile testing, the specimen was placed in de-ionized water for 24 hours, then coated in petroleum jelly using a cotton swab immediately prior to load cycling.
+
+				The end of the fabric gripped and used to elongate the snout in tension; in this way, the rostrum would return to its original configuration in a springlike manner, as in living specimens, rather than being forced into position.
+				The snout was aligned such that complete elongation of the curved section would take place in tension with minimal off-axis deflection of the un-curved section.
+				The specimen was subjected to displacement-controlled loading sufficient to fully extend the snout and generate a load stress of 1\,N, or approximately 20\% of the tensile strength of the species average.
+				Load cycling took place at a rate of 0.33Hz, and was continued for 14 days (ca.~400K cycles), until the tensile stress in the sample approached an asymptotic minimum.
+
+				Once the test was concluded, the specimen was placed in a 50\% ethanol solution for 24 hours to clean the petroleum jelly from the rostrum.
+				The specimen was examined for surface fractures and micro-tears, then subjected to tensile testing via the same protocol as the other specimens to assess whether the cuticle had begun to fatigue.
 
 		\subsection*{Specimen imaging and microscopy}
-			Cold-fracture of rostra
+			The fracture surfaces of cold-fractured specimens were examined using scanning electron microscopy to characterize the composite profile and microstructure of the rostrum.
+			Fracture behavior was assessed using both light microscopy and SEM to image the fracture surfaces of the specimens in transverse view.
+			Electron microscopy was conducted using a JEOL JSM6300 scanning electron microscope, while light microscopy was conducted using a Leica M205 C stereomicroscope and attached	computer running the software Leica Application Suite (LAS), as well as a Visionary Digital Passport II system using a Canon EOS Mark 5D II camera outfitted with interchangeable microptic lenses.
+			Specimen length, layer thicknesses, macrofiber orientation angles, and cross-sectional areas were assessed in the Leica Application Suite and in Adobe Illustrator using pixel-wise measurements multiplied by a scaling factor for the image.
 
@@
 			\paragraph*{General Approach}
+
 			\paragraph*{Cuticle profile of model}
-				The cuticle of \textit{C. longinasus} is 50 microns thick; we use \textit{C. longinasus} because we have constitutive models for the cuticle regions of this species.
+				The cuticle of \textit{C.~longinasus} is 50 microns thick; we use \textit{C.~longinasus} because we have constitutive models for the cuticle regions of this species.
 				We assumed equal layer thicknesses in the endocuticle of the basal cuticle, and equal thicknesses of exocuticle and endocuticle.
@@
 				The angles of the endocuticle layers for both types of cuticle are as described above.
+
 			\paragraph*{Permutations}
 				We additionally calculated models for two hypothetical hybrid cuticles: one model has the layer thicknesses of the apex, but fiber orientations of the base, while the second has the fiber orientations of the apex, but the layer thicknesses of the base.
+
 			\paragraph*{Classical Laminate Plate Theory}
@@
 				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test and $R^{2}_{\sigma}$-difference test between the fully-specified model and a model lacking the phylogenetic effect.
+
 			\paragraph*{Hypothesis testing}

### 2019-02-14 tex_v2/manuscript.tex
done writing all sections

@@
 \begin{document}
-	%abstract - approx. 150 words
+	%word count: 156
 	\begin{abstract}
-		\blindtext
+		The acorn weevil snout exhibits remarkable flexibility and toughness that are derived from from the microarchitecture of its exoskeleton.
+		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
+		Using Classical Laminate Plate Theory (CLPT) we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
+		Additionally, changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton are related to the tensile and fatigue behavior of the snout in six species of diverse morphology.
+		We demonstrate that increased endocuticle thickness is strongly correlated with increased tensile strength in the snout.
+		Consequently, snout stiffness is shown to be inversely correlated with fracture toughness.
+		Finally, we identify exocuticle rich invaginations both as a likely site of crack initiation in tensile failure and as a source of morphological constraint on the evolution of the snout.
+
 	\end{abstract}
@@
 
-\section*{Introduction} \info{670 words, reduce to 600 or less in revisions, structure of argument/narrative is good}
+\section*{Introduction}
 %word count: 673
@@
 	Despite being composed of the same material as other rigid body parts, the snout can be repeatedly bent without evident damage.
-	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum\improvement{as illustrated in figure 1c}.
+	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum.
+\improvement{as illustrated in figure 1c}
 	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit; a single female may prepare hundreds of such sites during the adult stage.
@@
 		We previously derived the effective elastic constants of the cuticle regions of \textit{C.~longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
-		The cuticle was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively, in the head capsule; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
+		The cuticle of the head capsule was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
 
-		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hybrid cuticle lay-ups were also modeled, each with only one of the modifications.
-		A hybrid configuration with the angle stacking sequence of typical cuticle (i.e., in the head capsule) but possessing the layer thicknesses of the rostral apex had effective moduli $E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases from unmodified cuticle, respectively.
-		Similarly, a hypothetical cuticle with the layer thicknesses of ordinary cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle) exhibited elastic moduli $E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases from unmodified cuticle, respectively.
+		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hypothetical cuticle lay-ups were also modeled, each with only one of the modifications.
+		The effective moduli of a configuration with the angle stacking sequence of typical cuticle (i.e., in the head capsule) but possessing the layer thicknesses of the rostral apex were calculated as $E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases from unmodified cuticle, respectively.
+		Similarly, a hypothetical cuticle with the layer thicknesses of ordinary cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle) had effective elastic moduli of $E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases from unmodified cuticle, respectively.
 
@@
 		We then examined the correspondence between composite structure and mechanical behavior of the snout, using phylogenetic linear mixed effects models to account for phylogenetic non-independence in the residual variances of the models.
+		These associations reported below are statistically significant and, notably, independent of species membership.
 
@@
 		We observed that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
-		Consequently, there was a positive correlation between ultimate tensile strength of the specimen and the ratio of endocuticle to exocuticle cross-sectional area at the site of fracture.
+		Consequently, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
 
@@
 		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.
-		Instead, we found that the uniaxial elastic modulus of the snout under low strain (E-low) was inversely correlated with the ultimate strain and toughness of the snout (as was the secant modulus, E-sec)
-		These associations were all found to be statistically significant and independent of species membership.
+		Instead, we found that the uniaxial elastic modulus of the specimen under low strain (E-low) was inversely correlated with ultimate strain and fracture toughness (as was the secant modulus, E-sec).
+		We therefore can infer that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
 
@@
 the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
-		he endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers.
+		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers, leading to a trade-off between rigidity and toughness.
 		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant.
@@
 	\section*{Author contributions}
-		\begin{description}
-		\item[Andrew Jansen] Conducted sectioning and staining, microscopy and imaging, tensile and fatigue testing, statistical analysis, and participated in manuscript preparation.
-		\item[Jason Williams] Conducted tensile and fatigue testing, participated in manuscript preparation.
-		\item[Nikhilesh Chawla] Facilitated microscopy, tensile and fatigue testing, and participated in manuscript preparation.
-		\item[Nico Franz] Facilitated specimen acquisition and imaging, participated in manuscript preparation.
-		\end{description}
+		M.A.J. conducted sectioning and staining, microscopy and imaging, tensile and fatigue testing, statistical analysis, and participated in manuscript preparation.
+		J.W. conducted tensile and fatigue testing and participated in manuscript preparation.
+		N.C. facilitated microscopy, tensile and fatigue testing, and participated in manuscript preparation.
+		N.M.F. facilitated specimen acquisition and imaging and participated in manuscript preparation.
 
@@
 		Reprints and permissions information is available online at www.nature.com/reprints.
-		Correspondence and requests for materials should be addressed to Andrew Jansen.
+		Correspondence and requests for materials should be addressed to M.A.J.
 
@@
 				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ until failure.
-				Results were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
+				Engineering stresses and strains (using undeformed specimen dimensions) were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
 
@@
 			\paragraph*{General Approach}
+				The effective uni-axial membrane and transverse flexural elastic moduli of idealized cuticle organizations representing both the rostral apex and head capsule were estimated using Classical Laminate Plate Theory (CLPT).
+				The composite profiles of both types of cuticle were idealized using the layer thicknesses and stacking sequences observed in \textit{C.~longinasus}.
+				This particular species was chosen because we derived the effective elastic constants of the individual components of the cuticle in previous work.
+				In addition, \textit{C.~longinasus} exhibits a profile that is typical and representative for the genus \textit{Curculio}, based on examination of the six species used for tensile testing, as described previously.
 
-			\paragraph*{Cuticle profile of model}
-				The cuticle of \textit{C.~longinasus} is 50 microns thick; we use \textit{C.~longinasus} because we have constitutive models for the cuticle regions of this species.
-				We assumed equal layer thicknesses in the endocuticle of the basal cuticle, and equal thicknesses of exocuticle and endocuticle.
-				In the apical cuticle we assumed the exocuticle and upper 8 layers of the endocuticle were each 5 microns thick, with 4 thinner layers of endocuticle, each 1.25\,{\textmu}m thick.
-				The angles of the endocuticle layers for both types of cuticle are as described above.
+				For \textit{C.~longinasus} the total thickness of the cuticle in the head and rostrum is roughly 50\,{\textmu}m, as in most specimens of the other examined species.
+				In the head capsule, the exocuticle occupies between 30-50\% of the through-thickness of the laminate, with the remaining thickness nearly evenly divided between 12 layers of endocuticle.
+				We use the maximum (50\% of through-thickness, or 25\,{\textmu}m) for the model, since the cuticle appears to deviate from this value only in sulci (grooves), pores, and other scattered features of surface sculpture.
+				The macrofiber laminae of the endocuticle were assigned a stacking sequence sequence of $0\degree, 90\degree, 30\degree, -60\degree, 60\degree, -30\degree, 90\degree, 0\degree, -60\degree, 30\degree, -30\degree, 60\degree$, thereby representing pairs of orthogonal plies stacked at a constant rotation angle of $30\degree$, in approximation of the living tissue.
 
-			\paragraph*{Permutations}
-				We additionally calculated models for two hypothetical hybrid cuticles: one model has the layer thicknesses of the apex, but fiber orientations of the base, while the second has the fiber orientations of the apex, but the layer thicknesses of the base.
+				In the rostral apex, the exocuticle is reduced to a thin shell approximately 5\,{\textmu}m in thickness, or 10\% of the total cuticle thickness.
+				The endocuticle displays a more complex pattern of layer thicknesses in the rostral apex than in the head capsule.
+				Each of the eight outermost layers are of nearly equal thickness to the exocuticle (5\,{\textmu}m), while the four innermost layers have a combined thickness equal to that of the exocuticle or a single layer of outer endocuticle ($h_{outer}=5\,\text{{\textmu}m}, h_{inner}=1.25\,\text{{\textmu}m}$).
+				The stacking sequence with respect to the longitudinal axis of the rostrum forms an antisymmetric [$\pm45\degree$] angle-ply laminate.
+
+				To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hypothetical cuticle lay-ups were also modeled, each with only one of the modifications present in the cuticle of the rostral apex.
+				The first of these models has the layer thicknesses of the rostral apex, but fiber orientations of the head capsule, while the second has the fiber orientations of the rostral apex, but the layer thicknesses of the head capsule.
+
+				Because these laminates are not symmetric, there exists for each a bending-extension coupling matrix $[B]$ populated with non-zero terms, thus complicating the calculation of effective in-plane elastic moduli.
+				To circumvent this difficulty and enable meaningful comparison between each laminate, all of the layups are reflected about their inner surface.
+				This effectively doubles their thickness while producing a balanced, symmetric laminate with no coupling between bending and extension (i.e., $[B]=0_{3,3}$).
+				Estimation of in-plane elastic constants from the extension($[A]$) and bending ($[D]$) matrices is described in detail below; Matlab R2018b was used to numerically evaluate the final values of the effective elastic constants.
 
@@
 
-				Using the lay-ups specified for the cuticle permutations, we calculate the extensional stiffness matrix $[A]$, bending stiffness matrix $[D]$, and bending-extension coupling matrix $B$ for each laminate consisting of $n$ layers at a distance $z$ from the laminate mid-plane.
+				Using the lay-ups specified for each type of cuticle described above, we calculate the extensional stiffness matrix $[A]$, bending stiffness matrix $[D]$, and bending-extension coupling matrix $[B]$ for each laminate consisting of $n$ layers at a distance $z$ from the laminate mid-plane.
 				The elements of these matrices can be found according to:
@@
 
-				For symmetric laminates, $[B] = 0$, and therefore:
+				For symmetric laminates, $[B] = 0_{3,3}$, and therefore:
 
@@
 
-				and, by inverting this equation (let $A^* = A^{-1}$) and substituting based on the assumptions above, we infer:
+				and, by inverting this equation (let $A^* = A^{-1}$) and substituting $A^*_{12} = A^*_{16} = 0$ based on the assumptions above, we infer:
 
@@
 
-				Thus, given the assumption that $D_{12} = D_{16} = 0$, Young's modulus for the effective transverse flexural elasticity of the laminate can be found by making Eq.~\ref{eqn::bee} specific to transverse flexure of the x-axis and rearranging the terms:
+				Thus, given the assumption that $D^*_{12} = D^*_{16} = 0$, Young's modulus for the effective transverse flexural elasticity of the laminate can be found by making Eq.~\ref{eqn::bee} specific to transverse flexure of the x-axis and rearranging the terms:
 
@@
 		\subsection*{Statistical analysis}
-			\paragraph*{General approach}
-				To explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models to the data using maximum likelihood estimation.
+			\paragraph*{Model selection and fitting}
+				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation.
+				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''} using the \texttt{`nlme'} and \texttt{`ape'} packages.
+				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis (see Supplementary Data).
 				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
-				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion model of trait evolution along the phylogeny.
-				Response variables and covariates were natural-log transformed, as needed, to ensure model residuals were normally distributed and homoscedastic.
-				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test and $R^{2}_{\sigma}$-difference test between the fully-specified model and a model lacking the phylogenetic effect.
+				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion [\texttt{R:`ape':corBrownian}] model of trait evolution along a phylogeny.
+
+				The preferred phylogeny for North American \textit{Curculio} is that of (REF-Bonal et al 2016); this phylogeny was generated using Maximum-Likelihood methods and is untrametric, but has uniform internal branch lengths.
+				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species to represent the individual specimens examined for that species.
+				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.
+				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data; underestimation of branch length causes overestimation of phylogenetic signal, which was not found to be statistically significant for any of our models, and we therefore believe that this procedure does not affect the findings of this study.
+
+				Other models of trait evolution were considered during model exploration, including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models.
+				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance, viz.
+$R^2_{\sigma}$, and $R^2_{\beta*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
+				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects, and is generally used to compare between covariance structures (REF).
+				$R^2_{\beta*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare between fixed effects (REF).
+				We additionally estimated the phylogenetic signal in the residual variance of each model using two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$) and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$), as detailed below.
+
+				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect.
+				Models that incorporated covariance due to Brownian motion consistently produced a higher likelihood score and fit to the data; however, no model exhibited statistically significant phylogenetic signal in any of the variables.
+				Once models were fitted to the data, T and F-statistics were calculated to determine whether each cofactor was significantly different from zero.
 
 			\paragraph*{Hypothesis testing}
-				The following three hypotheses were tested using PGLMMs fitted using ML estimation:
-				\begin{enumerate}
-				\item The maximum sustained tensile force is proportional to the cross-sectional area of the endocuticle, and \emph{not} that of the exocuticle.
-				\item The ultimate tensile strength of the samples is inversely proportional to the ratio of exocuticle to endocuticle at the location of fracture.
-				\item Young's modulus of the samples is proportional to the length of the snout.
-				\end{enumerate}
+				Our objectives for hypothesis testing using the PGLMM's were threefold: \textbf{(1)} assess whether the altered ratio of exocuticle to endocuticle in the rostral apex has an effect on the tensile strength of the snout; \textbf{(2)} test whether a trade-off exists between specimen stiffness and resistance to fracture; and \textbf{(3)} examine whether snout length and flexibility are correlated.
 
-				We fitted a fully-specified model with the cross-sectional area of endocuticle and exocuticle at the site of fracture as fixed effects, including an interaction term, and with maximum tensile force sustained prior to fracture as a response variable.
-				This model was then compared to models with only cross-sectional area of either endocuticle or exocuticle as the sole fixed effect in the model.
-				We then tested the first hypothesis by using likelihood-ratio tests and $R^{2}_{\beta*}$-difference tests between each of the three models.
+				To test the relationship between relative layer thicknesses and tensile strength, we fitted a fully-specified model with the cross-sectional area of exocuticle and endocuticle at the site of fracture as fixed effects, including an interaction term, and with maximum tensile force sustained prior to fracture as a response variable.
+				This model was then compared to models with only cross-sectional area of either endocuticle or exocuticle, but not both, as the sole fixed effect in the model.
+				We then tested the whether one or both regions was significantly correlated to the maximum force sustained in tension by using likelihood-ratio tests and $R^{2}_{\beta*}$-difference tests between each of the three models.
+				We found that only the cross-sectional area of endocuticle was a significant fixed effect, and therefore we elected to remove cross-sectional area of exocuticle as a fixed effect to make the model more parsimonious.
+				The final model generated for our first aim featured the ratio of exocuticle to endocuticle in cross-section as a fixed effect, with ultimate tensile strength as the response variable.
 
-				The hypothesis that
+				For our second aim we examined the relationship between specimen stiffness and resistance to fracture.
+				Specimen stiffness was characterized using a low strain elastic modulus (averaged across the first 33\% of the stress-strain curve) and the secant modulus at failure.
+				Resistance to fracture was quantified in terms of failure strain and fracture toughness (measured as area under the stress-strain curve).
+				Four models were fitted, two of which had failure strain as the response variable and two with fracture toughness as response variable.
+				Each model then used one of the different measures of specimen stiffness as a fixed effect.
 
-			\paragraph*{Model selection and fitting}
-
-			\paragraph*{Estimating phylogenetic signal}
+				Finally, in the third aim we explored whether a size-effect might exist in the \textit{Curculio} rostrum, particularly if longer (and typically more curved) rostra were more flexible than shorter, straighter rostra.
+				We generated two models: the low strain and secant elastic moduli individually served as response variables, with specimen length as the fixed effect in both.
 
 		\subsection*{Code availability}
+			R, Python, and Matlab scripts used to manipulate and analyze the raw data (as well as their outputs), produce figures, and estimate effective elastic constants is available from the corresponding author upon request.
 
 		\subsection*{Data availability}
+			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 1).
+			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 2).
+PGLMM terms and output are provided in a PDF (Supplement 3).
+Raw and processed data will be provided by the corresponding author upon reasonable request.
 

### 2019-02-14 analysis/phylo_regress.R

@@ -67,7 +67,7 @@ Analysis1 = function(phylogeny, raw.data) {
   print(leveneTest(residuals(simple.model,
                              type = 'normalized') ~ raw.data$species,
                    data = raw.data))
-  print(ad.test(residuals(simple.model, type = 'normalized')))
+  print(shapiro.test(residuals(simple.model, type = 'normalized')))
   r2sigma = r2beta(simple.model, method = 'sgv')
   print(r2sigma)
   
@@ -76,7 +76,7 @@ Analysis1 = function(phylogeny, raw.data) {
   print(leveneTest(residuals(simple.brownian,
                              type = 'normalized') ~ raw.data$species,
                    data = raw.data))
-  print(ad.test(residuals(simple.brownian, type = 'normalized')))
+  print(shapiro.test(residuals(simple.brownian, type = 'normalized')))
   r2sigma.brownian = r2beta(simple.brownian, method = 'sgv')
   print(r2sigma.brownian)
   
@@ -92,7 +92,7 @@ Analysis1 = function(phylogeny, raw.data) {
   print(leveneTest(residuals(y.transformed,
                              type = 'normalized') ~ raw.data$species,
                    data = raw.data))
-  print(ad.test(residuals(y.transformed, type = 'normalized')))
+  print(shapiro.test(residuals(y.transformed, type = 'normalized')))
   r2sigma = r2beta(y.transformed, method = 'sgv')
   print(r2sigma)
   
@@ -101,7 +101,7 @@ Analysis1 = function(phylogeny, raw.data) {
   print(leveneTest(residuals(y.brownian,
                              type = 'normalized') ~ raw.data$species,
                    data = raw.data))
-  print(ad.test(residuals(y.brownian, type = 'normalized')))
+  print(shapiro.test(residuals(y.brownian, type = 'normalized')))
   r2sigma.brownian = r2beta(y.brownian, method = 'sgv')
   print(r2sigma.brownian)
   
@@ -116,7 +116,7 @@ Analysis1 = function(phylogeny, raw.data) {
   print(leveneTest(residuals(x.transformed,
                              type = 'normalized') ~ raw.data$species,
                    data = raw.data))
-  print(ad.test(residuals(x.transformed, type = 'normalized')))
[…]

### 2019-02-14 analysis/tensile.py

@@ -314,8 +314,8 @@ def xyplot(spec_set, save = True):
         new_l = spec.ufs() * 0.67
         new_u = spec.ufs() * 1.00
         spec.recalcE(lower = new_l, upper = new_u, graphical=False, normal=False)
-        x = ln(spec.endo() / spec.exo())
-        y = ln(spec.ufs())
+        x = ln(spec.length())
+        y = ln(spec.Etanl())
         #spec.uts() / spec.ufs()
 
         if spec.species() == 'caryae':

### 2019-02-14 tex_v2/manuscript.tex
proofreading main text done!!

@@
 		Using Classical Laminate Plate Theory (CLPT) we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
-		Additionally, changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton are related to the tensile and fatigue behavior of the snout in six species of diverse morphology.
+		Additionally, changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton are related to the tensile behavior of the snout in six species of diverse morphology, and we demonstrate that a highly curved rostrum can be completely straightened without structural damage.
 		We demonstrate that increased endocuticle thickness is strongly correlated with increased tensile strength in the snout.
@@
 
-{%manuscript outline:
-%3000 word limit, 150 word abstract, 500 word intro, 250 word fig caps
-
-%standard intro, talk about behavior, rostrum bending, and cuticle
-%weevil cuticle is known to be stronger in compression than in tension
-%in Curculio this means that the greatest risk of breakage would come from tension during bending of the snout
-%curculio appear to have mitigated the risk of breakage by making the rostral apex flexible
-%prior work suggests that this flexibility comes from a modified composite profile in the cuticle of the rostral apex
-%different parts of the head thus appear to have differently specialized cuticle to deal with extreme bending
-%only observed in 1 species, never any comparison across species/in genus, fractographic analysis never attempted to explain how snout is most likely to break/ how the cuticle fails (ie what failure modes are the weevils fighting against/trying to mitigate)
-
-%we therefore set out to characterize the modifications to the cuticle of the rostral apex that presumably allow bending, using six species representing a mix of phylogenetically related and disparate species.
-%because material rigidity and toughness are often inversely proportional, we also investigate how the composite structure of the cuticle contributes to the tensile strength of the snout, whether the flexible regions exhibit higher tensile strenth, and conduct fractographic analysis on tensile testing specimens to determine the failure modes of the rostrum/cuticle under tension.
-%We additionally present the results of a load cycling test for a representative specimen of a species with a highly elongate and curved snout to explore the possibility that bending of the rostrum results in elastic deformation of the cuticle.
-
-%This study demonstrates: (1) there are 2 microstructural modifications to the composite profile of the cuticle of the rostral apex that would enhance cuticular flexibility, present in all examined species, and likely across the genus (2) thickness of the encocuticle is directly proportional to the maximum tensile force sustained in tension, to the exclusion of the exocuticle, (3) the tensile strenth of the snout varies across the length of the head and is inversely proportional to exo/endocuticle ratio, such that the more flexible areas are also stronger under tension, (4) young's modulus is higher for longer snouts, reflecting increased reinforcement at the base of the snout, likely to prevent bucking fracture at the base, (5) because of the modifications to the cuticle of the rostral apex (which result in all of these changes to the mechanical properties of the snout), bending appears to be viscoelastic/elastic, with a relatively low risk of breakage during the lifetime of the beetle, (6) when the rostrum does fail, crack initiation appears to take place in the exocuticle-rich occipital sulci, which would be under tension during the drilling behavior in the live insect.
-
-%in conclusion, the cuticle of the rostral apex is modified to increase flexibility, allowing for nearly elastic bending, while simultaneously increasing tensile strength, thus resisting the predominant fracture modes observed in unaltered cuticle.
-%to our knowledge, this is the first time that modifications to the composite profile of the cuticle have been observed to optimized mechanical behavior in insects, or something like that.
-
-%this has important implications for understanding the evolution of this highly exaggerated structure, ie suggesting negative selection of breakage instead of positive selection of flexibility.
-possibly a single ancestral shift for each modification, opened a large region of morphospace for exploration, resulting in present day diversity of the genus.
-
-%-------------------------------------------------------------------------------
-%introduction - nature materials requires that introduction is given NO SECTION HEADER, 500 words
-}
-
-\section*{Introduction}
-%word count: 673
+	%word count:3486 - 71 = 3415
 	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
-	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure\unsure{give more numbers here}.
-	Beetle cuticle is divided into two structurally and mechanically distinct regions, the (outer) exocuticle and (inner) endocuticle.
-	The exocuticle is characterized by a transversely-isotropic helicoidal structure that is comparatively rigid (though brittle) and has been the subject of numerous prior studies\unsure{numbers and refs}.
-	By contrast, the endocuticle of beetles is comprised of large, aligned bundles (macrofibers) of chitin that are strongly anisotropic\unsure{numbers again}.
-	Arranged in unidirectional laminae, the macrofibers improve cuticle toughness by inhibiting crack formation and propagation between successive plies \cite{Kamp2010,Kamp2015,Hepburn1973}.
-
-	In general, impact-prone areas and exaggerated structures, such as horns and legs exhibit cuticle organization that resists deformation and fracture.
-	Acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
-	This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author.} are typified by an elongate structure of the head, called the rostrum (snout), which instead exhibits unusual distal flexibility.
+	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below.
+	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture; however, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
+		Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
+		This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author.}
+	instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout).
 	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
-	Despite being composed of the same material as other rigid body parts, the snout can be repeatedly bent without evident damage.
-	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition); the latter process causes significant, apparently elastic, deformation of the rostrum.
-\improvement{as illustrated in figure 1c}
-	By maintaining constant tension on the snout and rotating around the bore-hole, females are able to flex the rostrum into a near-perfectly straight configuration and thereby produce a linear channel into the fruit; a single female may prepare hundreds of such sites during the adult stage.
+	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~1c), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts.
+	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult.
 
@@
 	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we estimate the effect of these differences on the elasticity of the cuticle using Classical Laminate Plate Theory (CLPT).
-
 	Because recent studies have shown that the yield strength of the beetle exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
@@
 
-	We additionally describe the fracture mechanics of the snout, as pertains to both cuticle composite structure and tensile behavior, and consider how modification of the cuticle may reduce the risk of rostral fracture during oviposition.
-	Based on our findings, we posit that the composite profile of the rostral apex enables the snout to be flexed until straight while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
+	We additionally describe the fracture mechanics of the snout and consider how modification of the cuticle may prevent crack formation during oviposition.
+	Based on our findings, we posit that the composite profile of the rostral apex enables the snout to undergo repeated distal flexion while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.
 	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
@@
 	\section{Microstructure of the \textit{Curculio} rostrum}
-	%word count: 784
 		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
-		These layers are stacked at a more or less constant angle to each other, thus forming a quasi-isotropic laminate referred to as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}.
-		This layout effectively produces a transversely isotropic composite, mitigating the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton.
-
+		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
+		This layout effectively mitigates the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton.		
 		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.
-\info{~INSERT-FIGS-HERE} \cite{Kamp2010,Kamp2015}.
-		Typically, adjacent macrofiber plies are paired and pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other, see Figs.\info{~INSERT-REFS-HERE}), with a constant stacking angle \emph{between} pairs \info{what kind of laminate is this?}, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
+		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~REF \cite{Kamp2010,Kamp2015}.
+		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal (i.e., angled approx.
+$90\degree$ to each other, see Figs.~REF), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
 		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure.
@@
 
-		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens have revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
-		Additionally, in the head capsule, the thickness of the exocuticle in cross section is nearly equal to that of the endocuticle\info{thickness measurements here}.
-		However, we have also found that the cuticle composite lay-up of the rostral apex differs from that of the head capsule (see Fig.\info{~INSERT-REFS-HER}).
-		Distally the exocuticle is reduced to a thin shell\info{measurement here}, with the endocuticle thickened \info{measurement here} to offset this reduction and maintain a constant cuticle thickness in the head.
-		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, which are all oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
-
-		In previous work we identified these modifications to the composite structure of the cuticle within a single species, \textit{C.~longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
-		This composite profile has now been uncovered in the rostral apex of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
+		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
+		Additionally, in the head capsule, the thickness of the exocuticle and endocuticle in cross section are nearly equal (typically between 20-30\,{\textmu}m).
+		However, we found that the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule (see Fig.~REF).
+		Distally the exocuticle is reduced to a thin shell (ca.
+5\,{\textmu}m), with the endocuticle thickened to offset this reduction and maintain a constant cuticle thickness (ca.
+50\,{\textmu}m total) throughout its length.
+		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies, and are oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
+		We previously identified these modifications to the composite structure of the cuticle within a single species, \textit{C.~longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}; however this composite profile has now been uncovered in the rostrum of six additional, phylogenetically disparate, species (detailed in methods), indicating that this is likely a genus-wide trait.
 		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.
 
-		To investigate the effect of these cuticle modifications, we estimated uni-axial membrane and transverse flexural Young's moduli of the cuticle in both the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
-		We previously derived the effective elastic constants of the cuticle regions of \textit{C.~longinasus}, which we used here to construct constitutive equations for the entire cuticle of that species.
+		We estimated the effect of differential cuticle organization on uni-axial membrane and transverse-flexural Young's moduli of the cuticle in the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
+		The effective elastic constants of the cuticle regions of \textit{C.~longinasus} (estimated previously, see REF) were used to construct constitutive equations for the entire cuticle of that species.
 		The cuticle of the head capsule was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
-
-		To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hypothetical cuticle lay-ups were also modeled, each with only one of the modifications.
+		Two hypothetical cuticle lay-ups were also modeled to individually assess the contributions of either modified layer thickness or stacking angle sequence to cuticle flexibility.
 		The effective moduli of a configuration with the angle stacking sequence of typical cuticle (i.e., in the head capsule) but possessing the layer thicknesses of the rostral apex were calculated as $E_m~=~3.73\,\text{GPa}, E_f~=~4.31\,\text{GPa}$, representing 22\% and 29\% decreases from unmodified cuticle, respectively.
 		Similarly, a hypothetical cuticle with the layer thicknesses of ordinary cuticle but possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle) had effective elastic moduli of $E_m~=~3.77\,\text{GPa}, E_f~=~5.76\,\text{GPa}$, representing 21\% and 4.7\% decreases from unmodified cuticle, respectively.
-
 		Each of the cuticle modifications noted in the rostral apex individually decreased the elastic moduli of the cuticle; however, they appear to have a synergistic combined effect on cuticle elasticity, rather than a simple additive effect.
@@
 	\section{Force-controlled loading to fracture}
-	%word count: 608
-		To better characterize the mechanical behavior of the rostrum at failure, we performed tensile testing on the snouts of six \textit{Curculio} species, representing a mixture of closely and distantly related taxa.
-		Heads of female specimens of each species were rehydrated by immersion in de-ionized water for 24 hours to simulate the condition of the living tissue (but see Klocke and Schmitz 2009).
-		Each specimen was subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $0.05\,\text{N}\cdot\text{s}^{-1}$ (detailed in Methods).
-		We then examined the correspondence between composite structure and mechanical behavior of the snout, using phylogenetic linear mixed effects models to account for phylogenetic non-independence in the residual variances of the models.
-		These associations reported below are statistically significant and, notably, independent of species membership.
+		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods).
+		Each specimen was immersed in $\text{di-H}_2\text{O}$ for 24 hours, to simulate the living tissue (see REFS), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ (detailed in Methods).
+		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture.
+		We postulate that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain.
 
-		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in apparently brittle fracture after a period of strain hardening.
-		We believe that strain hardening prior to fracture is a consequence of stiff longitudinal axis of the macrofibers becoming more closely aligned to the longitudinal axis of the rostrum, thereby resisting tension more directly at higher strains.
-		We observed that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
-		Consequently, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
-
+		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLMM) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect.
+		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
+		Accordingly, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
 		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
 		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.
-		Instead, we found that the uniaxial elastic modulus of the specimen under low strain (E-low) was inversely correlated with ultimate strain and fracture toughness (as was the secant modulus, E-sec).
-		We therefore can infer that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
 
-		Additionally, we observed a moderate, stiffening size-effect with respect to rostral length.
-		This observation was initially quite puzzling, as we had expected that a longer (and typically more strongly curved [insert ecomorph paper ref here]) snout would need to be more flexible to avoid fracture during oviposition.
-		Based on preliminary confocal microscopy data (not presented herein), we speculate that this may be the result of a longer transition gradient from the basal profile to the apical profile in longer rostra.
-		We hypothesize that a decrease in exocuticle thickness along a longer portion of the base might reinforce the snout against buckling; however Young's modulus of the rostrum would be comparatively higher in these species because of the greater volume fraction of exocuticle.
-
-		From these tests, we infer that the gross elastic behavior of the cuticle is consistent across the genus, in agreement with our current understanding of cuticle mechanobiology.
-		It is therefore unlikely that differences in sclerotization or chitin composition within the cuticle are responsible for the mechanical behavior of the rostrum in different species; instead, a single mechanism (i.e.
+		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness, implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
+		We also observed a moderate, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (REFS) snout would require increased flexibility to avoid fracture during oviposition.
+		One possible explanation for this trend is that longer snouts may have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
+		Thus, Young's modulus of the rostrum would be comparatively higher in these species because of the higher volume fraction of exocuticle.
+		From these tests, we infer that the gross elastic behavior of the cuticle is consistent across the genus, and that a single mechanism (i.e.
 the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
+		in agreement with our current understanding of cuticle mechanobiology.
 		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers, leading to a trade-off between rigidity and toughness.
@@
 	\section{Load cycling of \textit{Curculio caryae}}
-	%word count: 474
 		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature.
-		The specimen was then aligned so that uniaxial tension would induce elongation of the apical portion of the rostrum.
-		The strain per cycle was fixed at an amplitude sufficient to completely elongate the snout and generate a tensile load equal to $0.2\times$ the average tensile strength of conspecifics in the straightened configuration, at a frequency of $0.33\,\text{Hz}$.
-		The test was terminated after a period of two weeks (ca.~400K cycles) when the strain amplitude appeared to reach an asymptotic minimum.
-
-		We observed viscoelastic behavior in the rostrum, as indicated by histeresis in the stress-strain relationship during each cycle.
+		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the un-curved section.
+		The strain per cycle was fixed at an amplitude sufficient to completely elongate the snout and generate a tensile load of 1.0\,N  in the straightened configuration (ca.
+20\% ultimate strength) at a frequency of $0.33\,\text{Hz}$.
+		The test was terminated after a period of two weeks (ca.~400K cycles) when the stress amplitude appeared to reach an asymptotic minimum.		
+		The rostrum behaved viscoelastically, indicated by histeresis in the stress-strain relationship during each cycle.
 		Strain amplitude decreased logarithmically with cycle number, and the specimen initially appeared to have deformed plastically during the test.
@@
 
-		We cannot fully account for the stress relaxation of the rostrum after testing, but we speculate that it arose from the same mechanism as the viscoelastic behavior observed in tensile testing.
-		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains along their length.
-		The viscoelasticity of the cuticle is thought, in part, to come from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
-		We believe that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test; however, without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and consequently deform viscoelastically (REF\info{stress history dependence paper}).
-		After 24 hours soaking in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain.
+		We cannot fully account for the stress relaxation of the rostrum after testing, but we speculate that it arose from the general mechanism associated with cuticle viscoelasticity.
+		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains along their length; macroscopic viscoelastic behavior results from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
+		We posit that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test; however, without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and deform viscoelastically (REF-stress history dependence paper).
+		After immersion in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain.
 
@@
 		Given this surprising result, it appears that the specimen was undamaged by the testing.
-		We therefore expect that under normal conditions in life, repeated bending of the snout does not exceed the yield strength of the cuticle, and the bending strain is purely elastic.
+		We therefore expect that under normal conditions in life, repeated bending of the snout does not exceed the yield strength of the cuticle.
 
@@
 	%word count: 653
-		Due to the complex failure modes evident upon examination of the fractured specimens, it was not always possible to identify void nucleation and crack initiation sites (both primary and secondary).
-		Nonetheless, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here.
+		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here.
 
-		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, indicative of comparatively brittle failure, due to the helicoidal arrangement of \textalpha-chitin laminae and resultant transverse isotropy of the Bouligand structure (REF).
-		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across unbroken endocuticle on the rostral surface adjacent to the plane of fracture.
-		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating that endocuticular macrofibers are less brittle than the exocuticle, likely as a consequence of the unidirectional alignment of their \textalpha-chitin nanofibrils (REF).
-		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers, the presence of exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, which caused ply-splitting via mode II fracture between macrofibers at high strain.
-		This intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing inter-ply delamination.
+		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure (REF).
+		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle (but with endocuticle intact) adjacent to the plane of fracture.
+		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers (REF).
+		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers, exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, implying that ply-splitting occurred via mode II fracture between macrofibers at high strain.
+		We Hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
 		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae.
@@
 		The first layer of endocuticle usually fractured along the same plane as the exocuticle, and ventrally the endocuticle laminae typically converged toward a scarp-like region characterized by severe delamination and numerous de-bonded macrofibers.
-		Additionally, macrofibers oriented in the direction of crack propagation exhibited extensive ply-splitting with intermittent transverse shearing, while macrofibers oriented against the direction of crack propagation primarily displayed fracture by transverse shear along the plane of ply-splitting in adjacent laminae.
+		Additionally, macrofibers aligned with the direction of crack propagation exhibited extensive ply-splitting with intermittent transverse shearing, while macrofibers oriented against the direction of crack propagation primarily displayed fracture by transverse shear along the plane of ply-splitting in adjacent laminae.
 		Because the laminae form a cylinder, contralateral fibers in the same lamina display opposing fracture modes.
@@
 
-		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile, wherein the cuticle is strongly differentiated along an anterior-posterior gradient.
-		In particular, the cuticle in the apical portion of the rostrum exhibits marked reduction in relative thickness of the and in stacking-angle between pseudo-orthogonal plies of the endocuticle, which are aligned at $\pm45\degree$ to the longitudinal axis of the snout.
-		Constitutive models demonstrated that an idealized apical cuticle could theoretically achieve 72\% and 60\% reductions in the effective longitudinal membrane and flexural moduli, respectively, compared to the cuticle of the head capsule, thus accounting for the observed flexibility of the rostral apex in live specimens.
+		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile, wherein the cuticle is strongly differentiated in terms of relative layer thicknesses and orientation angles along an anterior-posterior gradient.
+		These modifications are sufficient to achieve marked reduction in the effective membrane and flexural moduli of the cuticle (72\% and 60\%, respectively) in constitutive models based on CLPT, thereby accounting for the observed flexibility of the rostral apex in live specimens.
 		However, these reductions can only be realized with both modifications to the cuticle, which have a non-additive effect on cuticle elasticity, implying that these weevils require both modifications to function properly during oviposition.
+		Likewise, tensile and fatigue testing have revealed a trade-off between stiffness and fracture resistance (as measured by ultimate strain and toughness), mediated by the relative proportion of endocuticle in the laminate.
+		The altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, permitting the structure to be flexed without exceeding the elastic limits of the cuticle.
+		To our knowledge, this is the first time that the composite profile of the cuticle has been related to a gradient in elasticity and tensile performance across a cuticular structure in arthropods.
+		Because these associations were independent of species membership, we posit that the behavior of the cuticle is consistent across the genus, and that rostral flexibility is achieved exclusively through a modified cuticle lay-up in all \textit{Curculio} species.
+		This conclusion raises the intriguing possibility that a single ancestral shift in cuticle organization at the rostral apex -- yielding higher flexibility and tensile strength -- opened a large region of morphospace for exploration, resulting in present day diversity of the genus.
 
-		Our tensile testing data revealed that increased endocuticle thickness correlates with greater fracture strength of the cuticle, while increased stiffness was associated with lower failure strains and reduced fracture toughness.
-		Additionally, fatigue testing of a highly-curved rostrum confirmed that the apical portion of the snout can be subjected to repeated antero-dorsal flexion without evidence of damage.
-		We therefore conclude that the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, and that this profile allows the rostrum to be flexed into a completely straight configuration without exceeding the elastic limits of the cuticle.
-		Because these associations were independent of species membership, we posit that the behavior of the cuticle is consistent across the genus, and that rostral flexibility is achieved exclusively through a modified cuticle lay-up in all \textit{Curculio} species.
-
-		Based on fractographic analysis of the test-specimens, we infer that the endocuticle deforms viscoelastically along the longitudinal axes of the macrofibers under uniaxial tension, causing ply-splitting and delamination at high strain.
-		However, the overlying exocuticle exhibits brittle fracture at a comparatively lower strain, due to shearing between the stretching endocuticle fibers to which it is anchored.
+		Based on fractographic analysis of the test-specimens, we infer that the exocuticle exhibits brittle fracture at a comparatively low strain, due to shearing between the endocuticle macrofibers to which it is anchored; these macrofibers fail at higher strain as a result of mixed-mode shearing and tensile fracture within and between laminae.
 		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
+		These results imply that fracture initiation occurs in the comparatively brittle exocuticle, and that the reduction in exocuticle thickness in the snout apex might serve to mitigate crack formation in rostral bending.
 		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
-
 		From a biological perspective, these findings reveal an unexpected morphological source of evolutionary constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
 		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion; the doubly-thick exocuticle in the invagination thus creates an unavoidable, brittle weak-point in the otherwise endocuticle-dominated rostral apex.
-		This constraint, as well as the minimization of exocuticle thickness in the rostral apex, and the increased toughness derived from a thickened endocuticle, all point to avoidance of catastrophic structural failure as a driving selective pressure in the evolution of the rostrum.
+		This constraint, as well as the minimization of exocuticle thickness in the rostral apex, and the increased toughness derived from a thickened endocuticle, lead us to consider the avoidance of catastrophic structural failure as a driving selective pressure in the evolution of the rostrum.
 
@@
 				Specimens were loaded in a Tryton 250 Microforce Testing System equipped with a 5N load cell and mechanical clamp grip.
-				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ until failure.
-				Engineering stresses and strains (using undeformed specimen dimensions) were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
+				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ until failure, with a sampling interval of 0.1\,s.
+				Engineering stresses and strains were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
 
@@
 
-				The end of the fabric gripped and used to elongate the snout in tension; in this way, the rostrum would return to its original configuration in a springlike manner, as in living specimens, rather than being forced into position.
+				The end of the fabric gripped and used to elongate the snout in tension; in this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
 				The snout was aligned such that complete elongation of the curved section would take place in tension with minimal off-axis deflection of the un-curved section.
@@
 				Specimen stiffness was characterized using a low strain elastic modulus (averaged across the first 33\% of the stress-strain curve) and the secant modulus at failure.
-				Resistance to fracture was quantified in terms of failure strain and fracture toughness (measured as area under the stress-strain curve).
-				Four models were fitted, two of which had failure strain as the response variable and two with fracture toughness as response variable.
+				Resistance to fracture was quantified in terms of ultimate strain and fracture toughness (measured as area under the stress-strain curve).
+				Four models were fitted, two of which had ultimate strain as the response variable and two with fracture toughness as response variable.
 				Each model then used one of the different measures of specimen stiffness as a fixed effect.

### 2019-02-14 tex_v2/manuscript.tex
citations formatted

@@
 		Consequently, snout stiffness is shown to be inversely correlated with fracture toughness.
-		Finally, we identify exocuticle rich invaginations both as a likely site of crack initiation in tensile failure and as a source of morphological constraint on the evolution of the snout.
+		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure and as a source of morphological constraint on the evolution of the snout.
 
@@
 
-	%word count:3486 - 71 = 3415
-	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix.
-	Although \textalpha-chitin is brittle and strongly anisotropic\unsure{give some numbers here}, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below.
-	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture; however, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
+	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
+	Although \textalpha-chitin is brittle and strongly anisotropic, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below \cite{Kamp2010, Kamp2015, Neville1976}.
+	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture (e.g.
+\cite{Amini2015, Mccullough2014mech, Mccullough2014struc, Dirks2012, Dirks2013}); however, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
 		Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two part name) followed by the author and date of publication of the name.
-		This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author.}
-	instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout).
-	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts.
-	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~1c), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts.
-	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult.
+		This is not an in-line reference; it is a part of the name itself and refers to a particular species-concept as indicated in the description of the species by that author \cite{ICZN1999}.}
+	instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout) \cite{Toju2005, Jansen2016, Singh2016, Gibson1969}.
+	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts \cite{Morimoto2003, Ting1933, Ting1936, Dennell1942, Gibson1969}.
+	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~1c), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
+	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult \cite{Toju2005, Moffett1989, AguirreUribe1978}.
 
-	While this behavior has been observed in many species of \textit{Curculio}, it was unclear how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
+	While this behavior has been observed in many species of \textit{Curculio} \cite{Gibson1969, Moffett1989, AguirreUribe1978, Toju2005}, it was unclear how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
 	In this study we characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
 	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts, and we estimate the effect of these differences on the elasticity of the cuticle using Classical Laminate Plate Theory (CLPT).
-	Because recent studies have shown that the yield strength of the beetle exoskeleton is lower in tension than compression, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
+	Because recent studies have shown that the yield strength of the beetle exoskeleton is lower in tension than compression \cite{Longhai2017}, we perform a comparative analysis of the ultimate tensile strength of the rostrum across species and snout morphotypes; we also report the results of displacement-controlled load cycling of the snout in a species with strongly curved morphology.
 	We relate an observed increase in the volume fraction of endocuticle in the rostrum to higher tensile strength at the rostral apex in all tested species, and find that a strongly curved rostrum can be flexed repeatedly without harm to the structure.
@@
 	Thus, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
-	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton.
+	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (but see \cite{Matsumura2017}).
 }	
 	\section{Microstructure of the \textit{Curculio} rostrum}
-		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix.
-		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980,Bouligand1972,Neville1976}. 
-		This layout effectively mitigates the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton.		
-		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~REF \cite{Kamp2010,Kamp2015}.
+		In arthropods (including beetles), the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils; each layer is the thickness of a single fiber (2-4\,nm) embedded in a proteinaceous matrix \cite{Nikolov2010, Nikolov2011}.
+		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980, Bouligand1972, Neville1976}. 
+		This layout effectively mitigates the strong anisotropy of \textalpha-chitin to yield a versatile building material for the exoskeleton \cite{Vincent1982, Vincent2004, Nikolov2010, Nikolov2011}.		
+		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter in \textit{Curculio}) unidirectional bundles of chitin, called macrofibers.
+		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~REF \cite{Kamp2010, Kamp2015}.
 		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other, see Figs.~REF), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973,Kamp2010}.
-		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure.
-		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010,Kamp2015,Hepburn1973}.
+$90\degree$ to each other \cite{Cheng2009}, see Figs.~REF), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
+		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure \cite{Kamp2015, Nikolov2010}.
+		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010, Kamp2015, Hepburn1973}.
 
@@
 
-		We estimated the effect of differential cuticle organization on uni-axial membrane and transverse-flexural Young's moduli of the cuticle in the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods.
-		The effective elastic constants of the cuticle regions of \textit{C.~longinasus} (estimated previously, see REF) were used to construct constitutive equations for the entire cuticle of that species.
+		We estimated the effect of differential cuticle organization on uni-axial membrane and transverse-flexural Young's moduli of the cuticle in the rostral apex and head capsule using Classical Laminate Plate Theory (CLPT), as detailed in our methods \cite{Reddy2004, Jones2014}.
+		The effective elastic constants of the cuticle regions of \textit{C.~longinasus} (estimated previously, see \cite{Jansen2016}) were used to construct constitutive equations for the entire cuticle of that species.
 		The cuticle of the head capsule was estimated to have membrane and flexural moduli of $E_m~=~4.77\,\text{GPa}$ and $E_f~=~6.04\,\text{GPa}$, respectively; however, in the rostral apex we found that these values were reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36\,\text{GPa}, E_f~=~2.44\,\text{GPa}$).
@@
 	\section{Force-controlled loading to fracture}
-		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods).
-		Each specimen was immersed in $\text{di-H}_2\text{O}$ for 24 hours, to simulate the living tissue (see REFS), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ (detailed in Methods).
-		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture.
-		We postulate that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain.
+		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods) \cite{Hughes2004phylo, Hughes2004eco, Bonal2016, Bonal2011}.
+		Each specimen was immersed in $\text{di-H}_2\text{O}$ for 24 hours, to simulate the living tissue (see \cite{Klocke2011}), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ (detailed in Methods).
+		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture \cite{Mihai2017}.
+		We postulate that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain \cite{Munster2013}.
 
-		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLMM) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect.
+		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLMM) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
 		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
@@
 		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness, implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
-		We also observed a moderate, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (REFS) snout would require increased flexibility to avoid fracture during oviposition.
+		We also observed a moderate, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (see \cite{Hughes2004eco, Bonal2011}) snout would require increased flexibility to avoid fracture during oviposition.
 		One possible explanation for this trend is that longer snouts may have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
@@
 the modified composite profile) confers increased flexibility and tensile strength to the rostral apex in the genus \textit{Curculio}.
-		in agreement with our current understanding of cuticle mechanobiology.
 		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle, likely because of its organization into large bundles of aligned, anisotropic fibers, leading to a trade-off between rigidity and toughness.
@@
 	\section{Load cycling of \textit{Curculio caryae}}
-		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature.
+		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
 		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the un-curved section.
@@
 		We cannot fully account for the stress relaxation of the rostrum after testing, but we speculate that it arose from the general mechanism associated with cuticle viscoelasticity.
-		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains along their length; macroscopic viscoelastic behavior results from slippage between these chains as the hydrogen bonds break and reform in response to shearing between the chitin molecules.
-		We posit that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test; however, without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and deform viscoelastically (REF-stress history dependence paper).
-		After immersion in ethanol and water, the hydrogen bonds would relax sufficiently to allow the specimen to return to its original configuration, dissipating the accumulated strain.
+		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains and with the protein matrix along their length; macroscopic viscoelastic behavior results from slippage between these chains in response to shearing between the chitin molecules \cite{Vincent2004, Evans1967, Sun2012}.
+		We posit that repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test; however, without sufficient time for the material to \textit{completely} relax after deformation, the specimen would slowly accumulate strain and deform viscoelastically \cite{Munster2013}.
+		After immersion in ethanol and water, the cuticle would be sufficiently plasticized to allow the specimen to return to its original configuration, dissipating the accumulated strain.
 
@@
 	\section{Fractography of test specimens}
-	%word count: 653
-		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here.
+		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here (see \cite{Greenhalgh2009} for general approach).
 
-		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure (REF).
+		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure \cite{Nikolov2010, Nikolov2011}.
 		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle (but with endocuticle intact) adjacent to the plane of fracture.
-		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers (REF).
-		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers, exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, implying that ply-splitting occurred via mode II fracture between macrofibers at high strain.
-		We Hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
-		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae.
+		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers \cite{Kamp2010, Kamp2015}.
+		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers (see \cite{Kamp2015, Longhai2017}), exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae, implying that ply-splitting occurred via mode II fracture between macrofibers at high strain \cite{Jones2014, Reddy2004}.
+		We hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle; mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
+		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II (transverse tension/intra-laminar shear) fracture due to an increase in applied stress caused by ply-splitting in adjacent laminae \cite{Greenhalgh2009}.
 
-		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum.
+		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum \cite{Davis2014, Dennell1942}.
 		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation.
@@
 		From these patterns we hypothesize that the exocuticle-rich occipital sulci are the most likely site for the initiation of void nucleation and catastrophic failure of the integrated rostral cuticle in cross section, as illustrated in Fig~REF.
-		According to this model, structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate.
+		According to this model, structural failure would take place as cracks propagate through the endocuticle from these sutures, which penetrate the entire thickness of the laminate \cite{Davis2014}.
 		Although other, more complex failure modes have been observed, we posit that in live specimens this is the most likely mechanism of tensile failure because typical bending behavior generates tension only along the ventral surface of the rostrum.
@@
 	\section{Conclusions}
-	%word count: 514 (3706 total)
-
 		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile, wherein the cuticle is strongly differentiated in terms of relative layer thicknesses and orientation angles along an anterior-posterior gradient.
@@
 		Based on fractographic analysis of the test-specimens, we infer that the exocuticle exhibits brittle fracture at a comparatively low strain, due to shearing between the endocuticle macrofibers to which it is anchored; these macrofibers fail at higher strain as a result of mixed-mode shearing and tensile fracture within and between laminae.
-		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae.
+		This outcome is consistent with behavior shown in previous studies and is congruent with theoretical consideration of cuticle microstructure in CLPT, which predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae \cite{Jones2014, Reddy2004}.
 		These results imply that fracture initiation occurs in the comparatively brittle exocuticle, and that the reduction in exocuticle thickness in the snout apex might serve to mitigate crack formation in rostral bending.
@@
 	\section{Methods}
-	%actual methods after references
 		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.
@@
 	\begin{thebibliography}{50}
-	%50 references (as a guideline, in order of appearance in text)
 		\section*{References}
@@
 
-			\bibitem{Klocke2011}
-				Klocke, D. \& Schmitz, H.
-				Water as a major modulator of the mechanical properties of insect cuticle.
-				\textit{Acta Biomater.}
-				\textbf{7,}
-				2935--2942
-				(2011).
-
-			\bibitem{Vincent1982}
-				Vincent, J. F. V.
-				\textit{Structural biomaterials}
-				(Halsted Press,
-				New York,
-				1982).
-
 			\bibitem{Vincent2004}
@@
 				(2004).
-
-			\bibitem{Nikolov2011}
-				Nikolov, S. et al.
-				Robustness and optimal use of design principles of arthropod exoskeletons studied by ab initio-based multiscale simulations.
-				\textit{J. Mech. Behav. Biomed. Mater.}
-				\textbf{4:2,}
-				129--145,
-				(2011).
-
-			\bibitem{Nikolov2010}
-				Nikolov, S. et al.
-				Revealing the design principles of high-performance biological composites using Ab initio and multiscale simulations: The example of lobster cuticle.
-				\textit{Adv. Mater.}
-				\textbf{22:4,}
-				519--526,
-				(2010).
-
-			\bibitem{Blackwell1980}
-				Blackwell, J. \& Weih, M.
-				Structure of chitin-protein complexes: ovipositor of the ichneumon fly \textit{Megarhyssa}.
-				\textit{J. Mol. Biol.}
-				\textbf{137:1,}
-				49--60,
-				(1980).
-
-			\bibitem{Bouligand1972}
-				Bouligand, Y.
-				Twisted fibrous arrangements in biological materials and cholesteric mesophases.
-				\textit{Tissue Cell}
-				\textbf{4:2,}
-				189--217,
-				(1972).
-
-			\bibitem{Neville1976}
-				Neville, A. C., Parry, D. A. \& Woodhead-Galloway, J.
-				The chitin crystallite in arthropod cuticle.
-				\textit{J. Cell Sci.}
-				\textbf{21:1,}
-				73--82,
-				(1976).
-
-			\bibitem{Cheng2009}
-				Cheng, L., Wang, L., \& Karlsson, A. M.
-				Mechanics-based analysis of selected features of the exoskeletal microstructure of \textit{Popillia japonica}.
-				\textit{Mater. Res.}
-				\textbf{24:11,}
-				3253--3267,
-				(2009).
 
@@
 
+			\bibitem{Neville1976}
+				Neville, A. C., Parry, D. A. \& Woodhead-Galloway, J.
+				The chitin crystallite in arthropod cuticle.
+				\textit{J. Cell Sci.}
+				\textbf{21:1,}
+				73--82,
+				(1976).
+
+			\bibitem{Amini2015}
+				Amini, S., Tadayon, M., Idapalapati, S., and Miserez, A.
+				The role of quasi-plasticity in the extreme contact damage tolerance of the stomatopod dactyl club
+				\textit{Nat. Mater.}
+				\textbf{14:9,}
+				943--950,
+				(2015).
+
+			\bibitem{Mccullough2014struc}
+				McCullough, E. L., Tobalske, B. W., and Emlen, D. J.
+				Structural adaptations to diverse fighting styles in sexually selected weapons
+				\textit{Proc. Natl. Acad. Sci. USA}
+				\textbf{111:40,}
+				14484--14488,
+				(2014).
+
+			\bibitem{Mccullough2014mech}
+				McCullough, E. L.
+				Mechanical limits to maximum weapon size in a giant rhinoceros beetle
+				\textit{Proc. R. Soc. B}
+				\textbf{281,}
+				20140696,
+				(2014).
+
+			\bibitem{Dirks2013}
+				Dirks, J., Parle, E., and Taylor, D.
+				Fatigue of insect cuticle
+				\textit{J. Exp. Biol.}
+				\textbf{216:10,}
+				1924--1927,
+				(2013).
+
+			\bibitem{Dirks2012}
+				Dirks, J. and Taylor, D.
+				Fracture toughness of locust cuticle
+				\textit{J. Exp. Biol.}
+				\textbf{215:9,}
+				1502--1508,
+				(2012).
+
+			\bibitem{Singh2016}
+				Singh, S. S., Jansen, M. A., Franz, N. M., and Chawla, N.
+				Microstructure and nanoindentation of the rostrum of Curculio longinasus Chittenden, 1927 (Coleoptera: Curculionidae)
+				\textit{Mater. Charact.}
+				\textbf{118,}
+				206--211,
+				(2016).
+
+			\bibitem{Toju2005}
+				Toju, H. and Sota, T.
+				Imbalance of predator and prey armament: geographic clines in phenotypic interface and natural selection
+				\textit{Am. Nat.}
+				\textbf{167:1,}
+				105--117,
+				(2005).
+
+			\bibitem{Gibson1969}	
+				Gibson, L. P.
+				Monograph of the genus \textit{Curculio} in the New World (Coleoptera: Curculionidae). Part I. United States and Canada
+				\textit{Misc. Publ. Entomol. Soc. Am.}
+				\textbf{6,} 
+				239-–285,
+				(1969).
+
+			\bibitem{Dennell1942}
+				Dennell, R.
+				The structure and function of the mouth-parts, rostrum and fore-gut of the weevil \textit{ Calandra granaria} L.
+				\textit{Phil. Trans. R. Soc. Lond. B}
+				\textbf{231:581,}
+				247--291,
+				(1942).
+
+			\bibitem{Morimoto2003}
+				Morimoto, K. and Kojima, H.
+				Morphologic characters of the weevil head and phylogenetic implications (Coleoptera, Curculionoidea)
+				\textit{Esakia}
+				\textbf{43,}
+				133--169,
+				(2003).
+
+			\bibitem{Ting1933}
+				Ting, P. C.
+				Feeding mechanisms of weevils, their function, and relationship to classification
+				\textit{Mon. Bull. Dep. Agric. State Calif.}
+				\textbf{22,}
+				161–-165,
+				(1933).
+
+			\bibitem{Ting1936}
+				Ting, P. C.
+				The mouth parts of the coleopterous group Rhynchophora
+				\textit{Microentomology}
+				\textbf{1,}
+				93–-114,
+				(1936).
+
+			\bibitem{AguirreUribe1978}
+				Aguirre Uribe, L. A.
+				\textit{Biology of the immature stages of the pecan weevil \textnormal{Curculio caryae} (Horn) and oviposition habits of the adult weevil}
+				(Ph.D. thesis,
+				Texas A \& M University,
+				1978).
+
+			\bibitem{Moffett1989}
+				Moffett, M.
+				Life in a nutshell
+				\textit{Natl. Geogr. Mag.}
+				\textbf{176,}
+				783-–784,
+				(1989).
+
+			\bibitem{ICZN1999}
+				\textit{International Code of Zoological Nomenclature}
+				Fourth Edition. 
+				(The International Trust for Zoological Nomenclature,
+				London,
+				1999).
+
+			\bibitem{Longhai2017}
+				Longhai, L. et al.
+				Microstructure and mechanical properties of rostrum in Cyrtotrachelus longimanus (Coleoptera: Curculionidae)
+				\textit{Anim. Cells Syst.}
+				\textbf{21:3,}
+				199--206,
+				(2017).
+
+			\bibitem{Matsumura2017}
+				Matsumura, Y., Kovalev, A. E., and Gorb, S. N.
+				Penetration mechanics of a beetle intromittent organ with bending stiffness gradient and a soft tip,
+				\textit{Sci. Adv.}
+				\textbf{3:12,}
+				eaao5469,
+				(2017).
+
+			\bibitem{Nikolov2011}
+				Nikolov, S. et al.
+				Robustness and optimal use of design principles of arthropod exoskeletons studied by ab initio-based multiscale simulations.
+				\textit{J. Mech. Behav. Biomed. Mater.}
+				\textbf{4:2,}
+				129--145,
+				(2011).
+
+			\bibitem{Nikolov2010}
+				Nikolov, S. et al.
+				Revealing the design principles of high-performance biological composites using Ab initio and multiscale simulations: The example of lobster cuticle.
+				\textit{Adv. Mater.}
+				\textbf{22:4,}
+				519--526,
+				(2010).
+
+			\bibitem{Blackwell1980}
+				Blackwell, J. and Weih, M.
+				Structure of chitin-protein complexes: ovipositor of the ichneumon fly \textit{Megarhyssa}.
+				\textit{J. Mol. Biol.}
+				\textbf{137:1,}
+				49--60,
+				(1980).
+
+			\bibitem{Bouligand1972}
+				Bouligand, Y.
+				Twisted fibrous arrangements in biological materials and cholesteric mesophases.
+				\textit{Tissue Cell}
+				\textbf{4:2,}
+				189--217,
+				(1972).
+
+			\bibitem{Vincent1982}
+				Vincent, J. F. V.
+				\textit{Structural biomaterials}
+				(Halsted Press,
+				New York,
+				1982).
+
+			\bibitem{Cheng2009}
+				Cheng, L., Wang, L., \& Karlsson, A. M.
+				Mechanics-based analysis of selected features of the exoskeletal microstructure of \textit{Popillia japonica}.
+				\textit{Mater. Res.}
+				\textbf{24:11,}
+				3253--3267,
+				(2009).
+
+			\bibitem{Leopold1992}
+				Leopold, R. A., Newman, S. M., and Helgeson, G.
+				A comparison of cuticle deposition during the pre-and posteclosion stages of the adult weevil, Anthonomus grandis BOHEMAN (Coleoptera: Curculionidae)
+				\textit{Int. J. Insect Morphol. and Embryol.},
+				\textbf{21:1,}
+				37--62,
+				(1992).
+
+			\bibitem{Reddy2004}
+				Reddy, J. N.
+				\textit{Mechanics of laminated composite plates and shells: theory and analysis}
+				(CRC Press,
+				Philadelphia,
+				2004).
+
+			\bibitem{Jones2014}
+				Jones, R. M.
+				\textit{Mechanics of composite materials}
+				(CRC press,
+				Philadelphia,
+				2014).
+
+			\bibitem{Hughes2004phylo}
+				Hughes, J. and Vogler, A. P.
+				The phylogeny of acorn weevils (genus Curculio) from mitochondrial and nuclear DNA sequences: the problem of incomplete data
+				\textit{Mol. Phylogenet. Evol.}
+				\textbf{32:2,}
+				601--615,
+				(2004).
+
+			\bibitem{Bonal2016}
+				Bonal, R. et al.
+				Diversity in insect seed parasite guilds at large geographical scale: the roles of host specificity and spatial distance
+				\textit{J. Biogeogr.}
+				\textbf{43:8,}
+				1620--1630,
+				(2016).
+
+			\bibitem{Hughes2004eco}
+				Hughes, J. and Vogler, A. P.
+				Ecomorphological adaptation of acorn weevils to their oviposition site
+				\textit{Evolution}
+				\textbf{58:9,}
+				1971--1983,
+				(2004).
+
+			\bibitem{Bonal2011}
+				Bonal, R.,Espelta, J. M., and Vogler, A. P.
+				Complex selection on life-history traits and the maintenance of variation in exaggerated rostrum length in acorn weevils
+				\textit{Oecologia}
+				\textbf{167:4,}
+				1053--1061,
+				(2011).
+
+			\bibitem{Klocke2011}
+				Klocke, D. \& Schmitz, H.
+				Water as a major modulator of the mechanical properties of insect cuticle.
+				\textit{Acta Biomater.}
+				\textbf{7,}
+				2935--2942
+				(2011).
+
+			\bibitem{Mihai2017}
+				Mihai, L. A. and Goriely, A.
+				How to characterize a nonlinear elastic material? A review on nonlinear constitutive parameters in isotropic finite elasticity
+				\textit{Phil. Trans. R. Soc. Lond. A}
+				\textbf{473:2207,}
+				20170607,
+				(2017).
+
+			\bibitem{Munster2013}
+				M{\"u}nster, S. et al.
+				Strain history dependence of the nonlinear stress response of fibrin and collagen networks
+				\textit{Proc. Natl. Acad. Sci. USA}
+				\textbf{110:30,}
+				12197--12202,
+				(2013).
+
+			\bibitem{Felsenstein1985}
+				Felsenstein, J.
+				Phylogenies and the comparative method
+				\textit{Am. Nat.}
+				\textbf{125:1,}
+				1--15,
+				(1985).
+
+			\bibitem{Revell2010}
+				Revell, L. J.
+				Phylogenetic signal and linear regression on species data
+				\textit{Methods Ecol. Evol.}
+				\textbf{1:4,}
+				319--329,
+				(2010).
+
+			\bibitem{Evans1967}
+				Beament, J. W. L., Treherne, J. E., and Wigglesworth, V. B.
+				\textit{Advances in insect physiology}
+				Volume 4
+				(Academic Press Inc.,
+				London,
+				1967).
+
+			\bibitem{Sun2012}
+				Sun, J. Y. et al.
+				Differential constitutive equation of elytra cuticle by nanoindentation
+				\textit{Adv. Mater. Res.},
+				\textbf{343,}
+				1133--1139,
+				(2012).
+
+			\bibitem{Greenhalgh2009}
+				Greenhalgh, E.
+				\textit{Failure analysis and fractography of polymer composites}
+				(CRC press,
+				Boca Raton,
+				2009).
+
+			\bibitem{Davis2014}	
+				Davis, S. R.
+				\textit{Morphology, phylogeny, and evolutionary development in the	weevils (Insecta: Coleoptera: Curculionoidea)}
+				(Ph.D. thesis,
+				University of	Kansas,
+				2014).
 	\end{thebibliography}
 
+	\begin{acknowledgements}
+		\info{do this!!!}
 		%brief and 'without effusive statements'
@@
 		%Financial support/ funding sources/ grants.
-	\begin{acknowledgements}
-
 	\end{acknowledgements}
 
-%author contributions
-%MAJ: did tensile tests, SEM, confocal (if applicable), data analysis.
-%MAJ, NC: designed research, micro-CT (if applicable)
-%MAJ, NMF: field collection of test specimens, procurement of specimen loans, specimen prep.
-%MAJ, NC, NMF: All authors contributed to the writing, review, and revision of the manuscript.
 	\section*{Author contributions}
@@
 This set of specimens was supplemented with material housed in the following collections, using the codens of Arnett et al.
-(1993)(REF):
+\cite{Arnett1993}:
 
@@
 			Cold fracture, semi-thin sectioning, and tensile testing were conducted on randomly chosen female specimens belonging to six species obtainable through field work in the southwestern USA and northwestern Mexico.
-			Taxon sampling was targeted to represent a mixture of disparate radiations and sister taxa with a variety of rostral morphotypes, according to the phylogenetic hypotheses of authors(REF \& REF).
-			These six species of \textit{Curculio} used herein are \textit{C.~caryae} (Horn, 1873), \textit{C.~humeralis} (Casey, 1897), \textit{C.~proboscideus} Fabricius, 1775, \textit{C.~sulcatulus} (Casey, 1897), \textit{C.~uniformis} (LeConte, 1857), and \textit{C.~victoriensis} (Chittenden, 1904).
+			Taxon sampling was targeted to represent a mixture of disparate radiations and sister taxa with a variety of rostral morphotypes, according to the phylogenetic hypotheses of Hughes~et~al.~\cite{Hughes2004eco, Hughes2004phylo} and Bonal~et~al.~\cite{Bonal2016}.
+			The six species of \textit{Curculio} used herein are \textit{C.~caryae} (Horn, 1873), \textit{C.~humeralis} (Casey, 1897), \textit{C.~proboscideus} Fabricius, 1775, \textit{C.~sulcatulus} (Casey, 1897), \textit{C.~uniformis} (LeConte, 1857), and \textit{C.~victoriensis} (Chittenden, 1904).
 
@@
 			A female specimen was selected and the rostrum was separated from the head capsule with a fine-edged razor blade.
-			The apical 1/4th of	the rostrum was also removed and then discarded.
+			The apical $1/4^{\text{th}}$ of	the rostrum was also removed and then discarded.
 			The remaining portion	of the rostrum and the head capsule were then embedded in EMbed812, as follows.
@@
 
-				Prior to testing, each specimen was placed in de-ionized water for 24 hours to allow full saturation of the cuticle, simulating the condition of live tissue (see also REF).
+				Prior to testing, each specimen was placed in de-ionized water for 24 hours to allow full saturation of the cuticle, simulating the condition of live tissue (see \cite{Klocke2011}).
 				Once removed for testing, the specimen was gripped using the cyanoacrylate handles at the marked locations immediately beyond the anterior margin of the rostrum and the posterior margin of the head capsule
@@
 				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$ until failure, with a sampling interval of 0.1\,s.
-				Engineering stresses and strains were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
+				Engineering stresses ($\sigma_0=F/A_0$) and strains ($\epsilon_0=\Delta{l}/l_0$) were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.
 
@@
 
-				The end of the fabric gripped and used to elongate the snout in tension; in this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
+				The end of the fabric was gripped and used to elongate the snout in tension by pulling; in this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
 				The snout was aligned such that complete elongation of the curved section would take place in tension with minimal off-axis deflection of the un-curved section.
@@
 			The fracture surfaces of cold-fractured specimens were examined using scanning electron microscopy to characterize the composite profile and microstructure of the rostrum.
-			Fracture behavior was assessed using both light microscopy and SEM to image the fracture surfaces of the specimens in transverse view.
+			Fracture behavior of tensile testing specimens was assessed using both light microscopy and SEM to image the fracture surfaces of the specimens in transverse view.
 			Electron microscopy was conducted using a JEOL JSM6300 scanning electron microscope, while light microscopy was conducted using a Leica M205 C stereomicroscope and attached	computer running the software Leica Application Suite (LAS), as well as a Visionary Digital Passport II system using a Canon EOS Mark 5D II camera outfitted with interchangeable microptic lenses.
@@
 			\paragraph*{General Approach}
-				The effective uni-axial membrane and transverse flexural elastic moduli of idealized cuticle organizations representing both the rostral apex and head capsule were estimated using Classical Laminate Plate Theory (CLPT).
+				The effective uni-axial membrane and transverse flexural elastic moduli of idealized cuticle organizations representing both the rostral apex and head capsule were estimated using Classical Laminate Plate Theory (CLPT); for general approach see \cite{Jones2014, Reddy2004}).
 				The composite profiles of both types of cuticle were idealized using the layer thicknesses and stacking sequences observed in \textit{C.~longinasus}.
-				This particular species was chosen because we derived the effective elastic constants of the individual components of the cuticle in previous work.
+				This particular species was chosen because we derived the effective elastic constants of the individual components of the cuticle in previous work \cite{Jansen2016, Singh2016}.
 				In addition, \textit{C.~longinasus} exhibits a profile that is typical and representative for the genus \textit{Curculio}, based on examination of the six species used for tensile testing, as described previously.
@@
 				This effectively doubles their thickness while producing a balanced, symmetric laminate with no coupling between bending and extension (i.e., $[B]=0_{3,3}$).
-				Estimation of in-plane elastic constants from the extension($[A]$) and bending ($[D]$) matrices is described in detail below; Matlab R2018b was used to numerically evaluate the final values of the effective elastic constants.
+				Estimation of in-plane elastic constants from the extension($[A]$) and bending ($[D]$) matrices is described in detail below; Matlab R2018b was used to numerically evaluate the final values of the effective elastic constants \cite{Matlab}.
 
@@
 
-				If we make the simplifying assumptions (see ref.\info{reddy, 2007 and jones, 1998}) that \textbf{(1)} the laminate experiences pure axial loading and transverse bending (i.e., $N_{yy} = N_{xy} = 0$ and $M_{yy} = M_{xy} = 0$, respectively) and \textbf{(2)} the laminate is a beam of sufficiently high aspect ratio to minimize the Poisson effect and anisotropic shear coupling (i.e., below we effectively let $A^*_{12} = A^*_{16} = 0$ and $D^*_{12} = D^*_{16} = 0$), then we can calculate the in-plane effective flexural and axial Young's moduli of the laminate along the x-axis.
+				If we make the simplifying assumptions (see \cite{Jones2014, Reddy2004}) that \textbf{(1)} the laminate experiences pure axial loading and transverse bending (i.e., $N_{yy} = N_{xy} = 0$ and $M_{yy} = M_{xy} = 0$, respectively) and \textbf{(2)} the laminate is a beam of sufficiently high aspect ratio to minimize the Poisson effect and anisotropic shear coupling (i.e., below we effectively let $A^*_{12} = A^*_{16} = 0$ and $D^*_{12} = D^*_{16} = 0$), then we can calculate the in-plane effective flexural and axial Young's moduli of the laminate along the x-axis.
 
@@
 			\paragraph*{Model selection and fitting}
-				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation.
-				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''} using the \texttt{`nlme'} and \texttt{`ape'} packages.
+				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Houseworth2004, Stone2011glmm}.
+				Raw data was processed using a custom script in Python version 3.5.2 \cite{Python}.
+				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''} using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Ape, Nlme}.
 				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis (see Supplementary Data).
 				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
-				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion [\texttt{R:`ape':corBrownian}] model of trait evolution along a phylogeny.
+				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion [\texttt{R:`ape':corBrownian}] model of trait evolution along a phylogeny \cite{Munkemuller2012, Ape}.
 
-				The preferred phylogeny for North American \textit{Curculio} is that of (REF-Bonal et al 2016); this phylogeny was generated using Maximum-Likelihood methods and is untrametric, but has uniform internal branch lengths.
-				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species to represent the individual specimens examined for that species.
-				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.
-				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data; underestimation of branch length causes overestimation of phylogenetic signal, which was not found to be statistically significant for any of our models, and we therefore believe that this procedure does not affect the findings of this study.
+				The preferred phylogeny for North American \textit{Curculio} is that of Bonal~et~al.~\cite{Bonal2016}; this phylogeny was generated using Maximum-Likelihood methods and is untrametric, but has uniform internal branch lengths.
+				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species to represent the individual specimens examined for that species; because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.
+				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data; underestimation of branch length causes overestimation of phylogenetic signal, which was not found to be statistically significant for any of our models, and we therefore believe that this procedure does not affect the findings of this study \cite{Munkemuller2012, Molina2017, Stone2011star}.
 
-				Other models of trait evolution were considered during model exploration, including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models.
+				Other models of trait evolution were considered during model exploration, including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models \cite{Munkemuller2012, Ape}.
 				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance, viz.
 $R^2_{\sigma}$, and $R^2_{\beta*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
-				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects, and is generally used to compare between covariance structures (REF).
-				$R^2_{\beta*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare between fixed effects (REF).
-				We additionally estimated the phylogenetic signal in the residual variance of each model using two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$) and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$), as detailed below.
+				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects, and is generally used to compare between covariance structures \cite{Nakagawa2013, Jaeger2017, R2glmm}.
+				$R^2_{\beta*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare between fixed effects  \cite{Nakagawa2013, Jaeger2017, R2glmm}.
+				We additionally estimated the phylogenetic signal in the residual variance of each model using two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$) and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$) using the \texttt{R} package \texttt{`phytools'} \cite{Munkemuller2012, Phytools}.
 
-				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect.
+				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect \cite{Lmtest, R2glmm}.
 				Models that incorporated covariance due to Brownian motion consistently produced a higher likelihood score and fit to the data; however, no model exhibited statistically significant phylogenetic signal in any of the variables.
@@
 	\begin{thebibliography}{50}
+		\makeatletter
+		\addtocounter{NAT@ctr}{45}
+		\makeatother
 		\section*{References}	
-			\bibitem{lamport94}
-				Leslie Lamport,
-				\textit{\LaTeX: a document preparation system},
-				Addison Wesley, Massachusetts,
-				2nd edition,
-				1994.
+			\bibitem{Arnett1993}
+				Arnett Jr., R. H., Samuelson, G. A., and Nishida G. M.
+				\textit{The Insect and Spider Collections of the World, 2nd Edition.}
+				Fauna and Flora Handbook No. 11.
+				(Sandhill Crane Press,
+				Gainesville,
+				1993).
 
+			\bibitem{Matlab}
+				The MathWorks, Inc.,
+				\textit{MATLAB and Statistics Toolbox}
+				(Natick, Mass.,
+				Release R2018b,
+				2018).
+
+			\bibitem{Galecki2013}
+				Galecki, A. and Burzykowski, T.
+				\textit{Linear Mixed-Effects Models Using R.}
+				A Step-by-Step Approach
+				(Springer,
+				New York,
+				2013).
+
+			\bibitem{Hadfield2010}
+				Hadfield, J. D. and Nakagawa, S.
+				General quantitative genetic methods for comparative biology: phylogenies, taxonomies and multi-trait models for continuous and categorical characters
+				\textit{J. Evol. Biol.}
+				\textbf{23:3,}
+				494--508,
+				(2010).
+
+			\bibitem{Houseworth2004}
+				Housworth, E. A., Martins, E. P., and Lynch, M.
+				The Phylogenetic Mixed Model
+				\textit{Am. Nat.}
+				\textbf{163:1,}
+				84--96,
+				(2004).
+
+			\bibitem{Stone2011glmm}
+				Stone, G. N., Sean, N., and Felsenstein, J.
+				Controlling for non-independence in comparative analysis of patterns across populations within species
+				\textit{Philos. Trans. Royal Soc. B}
+				\textbf{366:1569,}
+				1410--1424,
+				(2011).
+
+			\bibitem{Python}
+				Python Core Team.
+				\textit{Python: A dynamic, open source programming language}
+				(Python Software Foundation, 
+				Version 3.5.2,
+				2018).
+
+			\bibitem{Rcite}
+				R Development Core Team
+				\textit{R: A Language and Environment for Statistical Computing}
+				(R Foundation for Statistical Computing,
+				Vienna,
+				Version 3.5.1,
+				2008).
+
+			\bibitem{Ape}
+				Paradis, E. and Schliep, K.
+				ape 5.0: an environment for modern phylogenetics and evolutionary analyses in R
+				\textit{Bioinformatics}
+				\textbf{5.0,}
+				(2018).
+
+			\bibitem{Nlme}
+				Pinheiro, J. et al.
+				nlme: Linear and Nonlinear Mixed Effects Models
+				\textbf{R package version 3.1-137,}
+				(2018).
+
+			\bibitem{Munkemuller2012}
+				M{\"u}nkem{\"u}ller, T. et al.
+				How to measure and test phylogenetic signal
+				\textit{Methods Ecol. Evol.}
+				\textbf{3:4,}
+				743--756,
+				(2012).
+
+			\bibitem{Stone2011star}
+				Stone, E. A.
+				Why the phylogenetic regression appears robust to tree misspecification
+				\textit{Syst. Biol.}
+				\textbf{60:3,}
+				245--260,
+				(2011).
+
+			\bibitem{Molina2017}
+				Molina-Venegas, R. and Rodr{\'\i}guez, M. {\'A}.
+				Revisiting phylogenetic signal; strong or negligible impacts of polytomies and branch length information?
+				\textit{BMC Evol. Biol.}
+				\textbf{17:1,}
+				1--53,
+				(2017).
+
+			\bibitem{Nakagawa2013}
+				Nakagawa, S. and Schielzeth, H.
+				A general and simple method for obtaining R2 from generalized linear mixed-effects models
+				\textit{Methods Ecol. Evol.}
+				\textbf{4:2,}
+				133--142,
+				(2013).
+
+			\bibitem{Jaeger2017}
+				Jaeger, B. C., Edwards, L. J., Das, K., and Sen, P. K.
+				An $\text{R}^2$ statistic for fixed effects in the generalized linear mixed model
+				\textit{J. Appl. Stat.}
+				\textbf{44:6,}
+				1086--1105,
+				(2017).
+
+			\bibitem{R2glmm}
+				Jaeger, B.
+				Computes R Squared for Mixed (Multilevel) Models
+				\textbf{Version 0.1.2}
+				(2017).
+
+			\bibitem{Phytools}
+				Liam, J. R.
+				phytools: An R package for phylogenetic comparative biology (and other things)
+				\textit{Methods Ecol. and Evol.}
+				\textbf{3,}
+				217--223,
+				(2012).
+
+			\bibitem{Lmtest}
+				Zeileis, A. and Hothorn, T.
+				Diagnostic Checking in Regression Relationships
+				\textit{R News}
+				\textbf{2:3,}
+				7--10,
+				(2002).
 	\end{thebibliography}

### 2019-02-15 tex_v2/manuscript.tex
got the kraken, updating everyhting

@@
 \usepackage{graphicx}
-%TODO remove package: blindtext
-\usepackage{blindtext}
 \usepackage{gensymb}
@@
 	}
-\usepackage{cleveref}
 \graphicspath{{figures/}}
-
-%TODO NOTES! REMOVE FROM COMPLETE DRAFT-----------------------------------------
-\usepackage{xargs}
-\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
-\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
-\newcommandx{\info}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{#2}}
-\newcommandx{\improvement}[2][1=]{\todo[linecolor=purple,backgroundcolor=purple!25,bordercolor=purple,#1]{#2}}
-\newcommandx{\noshow}[2][1=]{\todo[disable,#1]{#2}}
 %-------------------------------------------------------------------------------
@@
 \begin{document}
-	%word count: 156
 	\begin{abstract}
@@
 	%title and author block
-	{\title{Exoskeletal microstructure and tensile behavior of the acorn weevil rostrum}
-	%CONSIDER CHANGING THE TITLE AFTER WRITING MANUSCRIPT
+	{\title{Exoskeletal strength and cuticle composite profile of the acorn weevil rostrum}
 	\date{\today}
@@
 
-	{The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
+	The exoskeleton of Coleoptera (beetles) is a hierarchically-structured fibrous composite characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
 	Although \textalpha-chitin is brittle and strongly anisotropic, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (see kamp, vincent for review), which we characterize in detail below \cite{Kamp2010, Kamp2015, Neville1976}.
@@
 	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts \cite{Morimoto2003, Ting1933, Ting1936, Dennell1942, Gibson1969}.
-	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~1c), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
+	This structure is used by the female to feed and to excavate sites for egg-laying (oviposition, see Fig.~\ref{fig::curculio}), and can be repeatedly bent without evident damage despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
 	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit; a single female may prepare hundreds of such sites as an adult \cite{Toju2005, Moffett1989, AguirreUribe1978}.
@@
 
+	\begin{figure*}
+		\centering
+		\includegraphics[width=180mm]{fig1.pdf}
+		\caption{\textbf{Morphology and oviposition behavior of female \textit{Curculio}.} 
+			\textbf{a}, Lateral habitus image of female \textit{Curculio~sayi} ((Gyllenhal, 1836)) featuring the elongate, strongly curved rostrum.
+			\textbf{b}, Lateral view of the head of a female specimen of \textit{Curculio~longinasus}, with major anatomical features indicated.
+			\textbf{c}, Illustration of the oviposition behavior of female \textit{Curculio}, proceeding from left to right: a female makes an incision in the host fruit, flexes the head directly over the bore hole using the front legs, then maintains tension on the snout while rotating to excavate a linear channel into the fruit.
+During this process, the rostrum is bent until completely straight.
+		}
+		\label{fig::curculio}
+	\end{figure*}
+
 	We additionally describe the fracture mechanics of the snout and consider how modification of the cuticle may prevent crack formation during oviposition.
@@
 	To our knowledge, this is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (but see \cite{Matsumura2017}).
-}	
+
 	\section{Microstructure of the \textit{Curculio} rostrum}
@@
 		Beetle endocuticle, however, is unique among arthropods and is comprised of large (1-5\,{\textmu}m diameter in \textit{Curculio}) unidirectional bundles of chitin, called macrofibers.
-		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~REF \cite{Kamp2010, Kamp2015}.
+		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5\,\text{GPa}$, transverse: $ E_2~=~E_3~=~0.52\,\text{GPa}$ \cite{Jansen2016}) and arranged in unidirectional plies, seen in Figs.~\ref{fig::cuticle},~\ref{fig::profile} \cite{Kamp2010, Kamp2015}.
 		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal (i.e., angled approx.
-$90\degree$ to each other \cite{Cheng2009}, see Figs.~REF), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
+$90\degree$ to each other \cite{Cheng2009}, see Fig.~\ref{fig::profile}), with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
 		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure \cite{Kamp2015, Nikolov2010}.
@@
 
+		\begin{figure*}
+			\centering
+			\includegraphics[width=180mm]{fig2.pdf}
+			\caption{\textbf{Gross divisions of cuticle in the rostrum.}
+				\textbf{a}, Scanning electron micrograph of the head capsule, in frontal view, of female \textit{Curculio~sulcatulus} (Casey, 1897), with the rostrum removed.
+				\textbf{b}, Magnified view of the junction between the rostrum and head capsule showing the division of the cuticle into two general regions: the exocuticle and the endocuticle.
+			}
+			\label{fig::cuticle}
+		\end{figure*}
+
 		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens revealed that endocuticle in the head capsule fits this general profile, with an angle of approximately $30\degree$ between successive pairs of pseudo-orthogonal plies.
 		Additionally, in the head capsule, the thickness of the exocuticle and endocuticle in cross section are nearly equal (typically between 20-30\,{\textmu}m).
-		However, we found that the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule (see Fig.~REF).
+		However, we found that the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule (Fig.~\ref{fig::profile}).
 		Distally the exocuticle is reduced to a thin shell (ca.
@@
 		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.
+
+		\begin{figure*}
+			\centering
+			\includegraphics[width=180mm]{fig3.pdf}
+			\caption{\textbf{Composite profiles of rostral cuticle.}
+				\textbf{a}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~humeralis} (Casey, 1897), showing that the cuticle of the rostral apex is organized as a $\pm45\degree$ angle-ply laminate..
+				\textbf{b}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~caryae}  (Horn, 1873), showing that the cuticle of the rostral base and head capsule has a roughly $30\degree$ stacking angle between each pair of pseudo-orthogonal plies.
+				Semi-thin sections of cuticle from a specimen of \textit{C.~humeralis}, stained with toluidine-blue-borax, demonstrating:
+				(\textbf{c}) that the exocuticle of the rostral apex is reduced to a thin shell close to 5\,{\textmu}m thick, with the endocuticle thickened to maintain a constant laminate thickness; and,
+				(\textbf{d}) that the exocuticle of the head capsule and base of the snout occupies nearly half of the through-thickness of the cuticle.
+			}
+			\label{fig::profile}
+		\end{figure*}
 
@@
 	\section{Force-controlled loading to fracture}
+
+		\begin{figure*}
+			\centering
+			\includegraphics[width=180mm]{fig4.pdf}
+			\caption{\textbf{Tensile properties of the \textit{Curculio} rostrum.}
+				Each plot shows the relationship between two variables as predicted by a phylogenetic linear mixed-effect model, with species as a random effect and a variance-covariance matrix generated from Brownian motion over the preferred phylogeny of Bonal et al.
+\cite{Bonal2016}.
+				The gray regions represent the prediction interval and bootstrapped 95\% confidence interval of the model.
+				The estimated fixed effect $\hat{\beta}$ is given, along with the p-value of a t-test assessing whether $\hat{\beta}$ is significantly different from zero.
+				Also reported is a generalized marginal $\text{R}^2_{\beta^*}$ for assessing fixed effects (see online Methods for more information).
+				In general, we find that increased endocuticle thickness is associated with greater tensile strength, and that stiffness is inversely correlated with toughness.
+			}
+			\label{fig::tensile}
+		\end{figure*}
+
 		To better characterize the failure behavior of the rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa (see Methods) \cite{Hughes2004phylo, Hughes2004eco, Bonal2016, Bonal2011}.
@@
 
-		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLMM) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
-		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle, and not the exocuticle, at that site.
-		Accordingly, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture.
-		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
+		We examined the correspondence between composite structure and mechanical behavior of the snout using phylogenetic linear mixed effects models (PGLM, Fig.~\ref{fig::tensile}) to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
+		Our models show that the maximum force sustained at the site of failure was strongly correlated with the cross-sectional area of the endocuticle ($\hat{\beta}=1.28, p<0.0001$), and not the exocuticle ($\hat{\beta}=0.17, p=0.38$), at that site.
+		Accordingly, there was a negative correlation between ultimate tensile strength of the specimen and the ratio of exocuticle to endocuticle cross-sectional area at the site of fracture ($\hat{\beta}=-0.36, p=0.017$).
+		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no evidence of correspondence between the cross-sectional properties of the fracture site and the gross behavior of the entire rostrum.
 		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to the properties of the entire snout.
 
-		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness, implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
-		We also observed a moderate, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (see \cite{Hughes2004eco, Bonal2011}) snout would require increased flexibility to avoid fracture during oviposition.
-		One possible explanation for this trend is that longer snouts may have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
+		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness (see Fig.~\ref{fig::tensile}), implying that stiffer specimens, and by extension, stiffer cuticle profiles, are generally more brittle.
+		We also observed a moderate, but not statistically significant, stiffening size-effect with respect to rostral length, contrary to our expectation that a longer, more strongly curved (see \cite{Hughes2004eco, Bonal2011}) snout would require increased flexibility to avoid fracture during oviposition.
+		One possible explanation for this trend might be that longer snouts have a longer transition gradient from basal to apical profile to reinforce the junction between the rostrum and head capsule against buckling.
 		Thus, Young's modulus of the rostrum would be comparatively higher in these species because of the higher volume fraction of exocuticle.
@@
 	\section{Load cycling of \textit{Curculio caryae}}
-		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
+
+		\begin{figure*}
+			\centering
+			\includegraphics[width=180mm]{fig5.pdf}
+			\caption{\textbf{Fatigue testing of a female C. caryae.}
+				\textbf{a}, Shows the fatigue testing setup, wherein the head capsule was fixed to a pedestal and the rostrum was attached to a strip of rip-stop nylon fabric using cyanoacrylate adhesive.
+				The specimen was loaded in tension, and not compression, isolating the effect of tension on the fatigue behavior of the rostrum.
+				\textbf{b}, An overlay of the pre- and post-fatigue states of the head show a clear effect from repeated proscribed strain; however,
+				\textbf{c}, Demonstrates that this effect is apparently not permanent.
+			The first two panels to the left correspond to the pre- and post-fatigue states of the rostrum, respecively, but the third panel shows that the head has returned to its original shape after 24 hours soaking in a water/ethanol mixture.
+				\textbf{d}, The two plots presented here are the raw force and displacment data from the fatigue test.
+				The plot on the right shows clear viscoelastic behavior, indicated by histeresis in the stress-strain response of the specimen, and exhibits a logarithmic decrease stress amplitude over time.
+			}
+			\label{fig::fatigue}
+		\end{figure*}
+
+		To confirm that repeated, proscribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of \textit{Curculio caryae}, a species that exhibits extreme ($80-90\degree$, Fig.~\ref{fig::fatigue}) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
 		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the un-curved section.
@@
 		Strain amplitude decreased logarithmically with cycle number, and the specimen initially appeared to have deformed plastically during the test.
-		We believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that rostrum returned to its original shape.
+		We believed that this indicated damage to the specimen; however, after cleaning the specimen in a 24 hour wash with ethanol and water, we observed that rostrum returned to its original shape (Fig.~\ref{fig::fatigue}).
 
@@
 	\section{Fractography of test specimens}
-		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure, which we describe here (see \cite{Greenhalgh2009} for general approach).
+
+		\begin{figure*}
+			\centering
+			\includegraphics[width=180mm]{fig6.pdf}
+			\caption{\textbf{Fractography of the \textit{Curculio} rostrum.}
+				\textbf{a}, Light micrograph showing the fracture surface of a tensile tested rostrum, in the species \textit{C.~caryae}, exhibiting a typical failure mode, illustrated in (\textbf{b}).
+				The small, black arrows indicate the winding direction of the macrofiber lamina; green arrows and dotted lines indicate the direction of crack.
+				Scanning electron micrographs highlight: (\textbf{c}) shear cusp formation, (\textbf{d}) tensile failure and off-axis macrofiber fiber-pulling, (\textbf{e}) interlaminar delamination, (\textbf{f},\textbf{h}), crack formation near invaginated exocuticle (\textbf{g}) ventral crack-front coalescence and shear cusp formation.
+			}
+			\label{fig::fracture}
+		\end{figure*}
+
+		Although it was not always possible to identify void nucleation and crack initiation sites given the complex failure modes evident in the fractured specimens, we observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure (Fig.~\ref{fig::fracture}), which we describe here.
 
 		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface, characteristic of comparatively brittle failure, presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure \cite{Nikolov2010, Nikolov2011}.
-		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle (but with endocuticle intact) adjacent to the plane of fracture.
+		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle adjacent to the plane of fracture \cite{Greenhalgh2009}.
 		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface, indicating the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers \cite{Kamp2010, Kamp2015}.
@@
 		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticle invaginations that traverse the entire length of the rostrum \cite{Davis2014, Dennell1942}.
-		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation.
+		These sulci increase the volume fraction of exocuticle in the ventral part of the snout and contain large interfaces ideal for void nucleation (see Fig.~\ref{fig::fracture}).
 		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, continuing dorsolaterally and ventromedially.
@@
 	\begin{acknowledgements}
-		\info{do this!!!}
-		%brief and 'without effusive statements'
-		%thank everyone who helped with data collection, dave (SEM), charlie (specimens), paige (if I use confocal data), Sal (if I use histology/staining data or confocal data, he helped with specimen clearing).
-		%Financial support/ funding sources/ grants.
+		The authors are grateful to Robert Anderson (CMNC), Lourdes Chamorro (USNM), and Charlie O'Brien (CWOB) for their assistance and provision of specimens used in this study.
+		The authors thank Salvatore Anzaldo, Andrew Johnston, Brian Reilly, Sangmi Lee, and other ASUHIC members for their assistance in procuring, treating, and maintaining specimen loans upon entry into the ASUHIC.
+		The authors would also like to thank David Lowry (ASU CLAS Bioimaging Center) for his training and assistance with resin embedding, scanning electron microscopy, and microtomography, and Salvatore Anzaldo for his assistance with histological staining.
 	\end{acknowledgements}
@@
 
-%provide detailed methods here, subsection for each method, statistical analysis, statements of code and data avilability.
 	\section*{Methods}
@@
 
-				The end of the fabric was gripped and used to elongate the snout in tension by pulling; in this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
+				The end of the fabric was gripped and used to elongate the snout in tension by pulling on the fabric, thus isolating the effect of tension on the fatigue life of the rostral cuticle.
+				In this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
 				The snout was aligned such that complete elongation of the curved section would take place in tension with minimal off-axis deflection of the un-curved section.
@@
 				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance, viz.
-$R^2_{\sigma}$, and $R^2_{\beta*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
-				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects, and is generally used to compare between covariance structures \cite{Nakagawa2013, Jaeger2017, R2glmm}.
-				$R^2_{\beta*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare between fixed effects  \cite{Nakagawa2013, Jaeger2017, R2glmm}.
+$R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
+				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects, and is generally used to compare between covariance structures \cite{Nakagawa2013, Jaeger2016, R2glmm}.
+				$R^2_{\beta^*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare between fixed effects  \cite{Nakagawa2013, Jaeger2016, R2glmm}.
 				We additionally estimated the phylogenetic signal in the residual variance of each model using two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$) and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$) using the \texttt{R} package \texttt{`phytools'} \cite{Munkemuller2012, Phytools}.
@@
 				This model was then compared to models with only cross-sectional area of either endocuticle or exocuticle, but not both, as the sole fixed effect in the model.
-				We then tested the whether one or both regions was significantly correlated to the maximum force sustained in tension by using likelihood-ratio tests and $R^{2}_{\beta*}$-difference tests between each of the three models.
+				We then tested the whether one or both regions was significantly correlated to the maximum force sustained in tension by using likelihood-ratio tests and $R^{2}_{\beta^*}$-difference tests between each of the three models.
 				We found that only the cross-sectional area of endocuticle was a significant fixed effect, and therefore we elected to remove cross-sectional area of exocuticle as a fixed effect to make the model more parsimonious.
@@
 
-			\bibitem{Jaeger2017}
+			\bibitem{Jaeger2016}
 				Jaeger, B. C., Edwards, L. J., Das, K., and Sen, P. K.
@@
 				1086--1105,
-				(2017).
+				(2016).
 

### 2019-02-15 analysis/phylo_regress.R

@@ -643,8 +643,8 @@ main = function() {
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$fmax)
-  raw.data$cofactor = raw.data$exo
-  comparison = c('ln(Fmax [N])', 'Exocuticle CSA [mm2]')
+  raw.data$cofactor = log(raw.data$exo)
+  comparison = c('ln(Fmax [N])', 'ln(Exocuticle CSA [mm2])')
   Analysis3(phylo.vcv, tree.star, raw.data, comparison)
   Analysis4(phylo.vcv, raw.data, comparison)
 
@@ -652,7 +652,7 @@ main = function() {
   readline(prompt = 'uts vs. ratio: press enter to continue.')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
-  writeLines('Model 3a:log(uts) ~ log(ratio)')
+  writeLines('Model 1c:log(uts) ~ log(ratio)')
   writeLines(strrep('=',80))
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$uts)
@@ -671,7 +671,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$ufs)
   raw.data$cofactor = log(raw.data$Etanl)
-  comparison = c('ln(Ultimate Strain [mm2])', 'ln(Tangent Modulus [MPa])')
+  comparison = c('ln(Ultimate Strain [mm2])', 'ln(Elastic Modulus [MPa])')
   Analysis3(phylo.vcv, tree.star, raw.data, comparison)
   Analysis4(phylo.vcv, raw.data, comparison)
 
@@ -697,7 +697,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$U)
   raw.data$cofactor = log(raw.data$Etanl)
-  comparison = c('ln(Fracture Toughness [Jm-3])', 'ln(Tangent Modulus [MPa])')
+  comparison = c('ln(Fracture Toughness [Jm-3])', 'ln(Elastic Modulus [MPa])')
   Analysis3(phylo.vcv, tree.star, raw.data, comparison)
   Analysis4(phylo.vcv, raw.data, comparison)
 
@@ -723,7 +723,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$Etanl)
[…]

### 2019-02-15 analysis/tensile.py

@@ -59,7 +59,6 @@ class specimen:
         self.__E__ = linreg(xvals, yvals) #engineering stress and strain via Hooke's Law
 
         self.__Etanl__ = self.__E__[0]
-        self.__Etanh__ = 0
 
         t_upper = 0
         t_lower = 0
@@ -159,9 +158,6 @@ class specimen:
     def E(self, index = 0):
         return self.__E__[index]
 
-    def Etanh(self):
-        return self.__Etanh__
-
     def Etanl(self):
         return self.__Etanl__
 
@@ -250,10 +246,10 @@ def stress_strain(spec_set, save=True, i=0, j=-1):
         ax2.set_ylabel('Axial Displacement (mm)', color='r')
         ax2.tick_params('y', colors='r')
 
-        ax3.plot(spec.epsilonT()[i:j], spec.sigmaT()[i:j], color = 'gray')
+        # ax3.plot(spec.epsilonT()[i:j], spec.sigmaT()[i:j], color = 'gray')
         ax3.plot(spec.epsilonN()[i:j], spec.sigmaN()[i:j], color = 'black')
         ax3.set_ylabel('Stress (MPa)')
-        ax3.set_xlabel('Strain (ln(L/L0) or dL/L0)')
+        ax3.set_xlabel('Strain (dL/L0)')
 
         print(spec.asuid()+' '+spec.species())
         new_l = spec.ufs() * 0.00
@@ -263,18 +259,15 @@ def stress_strain(spec_set, save=True, i=0, j=-1):
                  spec.E()*spec.epsilonN()[spec.Eindex(0):spec.Eindex(1)] + spec.E(1),
                  color='r')
 
-        new_l = spec.ufs() * 0.67
-        new_u = spec.ufs() * 1.00
-        spec.recalcE(lower = new_l, upper = new_u, graphical=False, normal=False)
-        ax3.plot(spec.epsilonN()[spec.Eindex(0):spec.Eindex(1)],
-                 spec.E()*spec.epsilonN()[spec.Eindex(0):spec.Eindex(1)] + spec.E(1),
-                 color='r')
+        ax3.plot([min(spec.epsilonN()), max(spec.epsilonN()[0:-1])],
+                 [min(spec.sigmaN()), max(spec.sigmaN()[0:-1])],
+                 color='gray')
 
-        true = mpatches.Patch(color='gray', label='True')
-        eng = mpatches.Patch(color='black', label='Nominal')
[…]

### 2019-03-16 tex_v3/JansenEtAL2019-AcornWeevilRostrum-NMFedits-Feb2019.tex
nico edits

\documentclass[twocolumn, linenumbers, superscriptaddress, nofootinbib]{revtex4-1}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{textgreek}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
	\hypersetup{colorlinks,
	linkcolor={red!50!black},
	citecolor={blue!60!black},
	urlcolor={blue!40!black}
	}
\graphicspath{{figures/}}
%-------------------------------------------------------------------------------

\begin{document}
	\begin{abstract}
		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
		Rostrum stiffness is shown to be inversely correlated with fracture toughness; thus allowing a highly curved rostrum to completely straighten without structural damage.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.

	\end{abstract}

	%title and author block
	{\title{Exoskeletal strength and cuticle composite profile of the acorn weevil rostrum}
	\date{\today}

	\author{M. Andrew Jansen}
		\email[corresponding author, email:~]{majanse1@asu.edu}
		\affiliation{School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA}
	\author{Jason Williams}
		\affiliation{School for Engineering of Matter, Energy, and Transport, Arizona State University, Tempe, AZ 85287, USA}
	\author{Nikhilesh Chawla}
		\affiliation{School for Engineering of Matter, Energy, and Transport, Arizona State University, Tempe, AZ 85287, USA}
	\author{Nico M. Franz}
		\affiliation{School of Life Sciences, Arizona State University, Tempe, AZ 85287, USA}

	\maketitle
	}

	The exoskeleton of Coleoptera (beetles) is a hierarchically structured, fibrous composite -- characterized by variously arranged \textalpha-chitin (N-acetylglucosamine) nanofibrils that are embedded in a heterogeneous protein matrix \cite{Hepburn1973, Jansen2016, Vincent2004}.
	Although \textalpha-chitin is brittle and strongly anisotropic, beetle cuticle is simultaneously rigid and tough due to its unique laminate microstructure (reviewed in \cite{Kamp2010, Kamp2015, Neville1976}), characterized in detail below.
	Impact-prone areas and exaggerated structures in arthropods generally exhibit cuticle organization that resists deformation and fracture \cite{Amini2015, Mccullough2014mech, Mccullough2014struc, Dirks2012, Dirks2013}.
	However, acorn weevils in the genus \textit{Curculio} Linnaeus, 1758\footnote{
		Pursuant to the International Code of Zoological Nomenclature, the first mention of any specific epithet will include the full genus and species names as a binomen (two-part name) followed by the author and date of publication of the name.
		This is not an in-line reference; it is a part of the name itself and refers to a particular act establishing the validity and fixing the identity of the corresponding name by that author \cite{ICZN1999}.}
	(Curculionidae in the sense of \cite{Davis2014}) instead exhibit unusual distal flexibility in an elongate extension of the head called the rostrum (snout) \cite{Toju2005, Jansen2016, Singh2016, Gibson1969}.
	The rostrum is a hollow, strongly curved (over 90$\degree$ in some species), cylindrical, exoskeletal extension of the otherwise nearly-spherical head, which bears at its apex the terminal chewing mouthparts \cite{Morimoto2003, Ting1933, Ting1936, Dennell1942, Gibson1969}.
	This structure is used by the female to feed on fruit (plant) tissue and to excavate sites for egg-laying (oviposition, see Fig.~\ref{fig::curculio}).
	The rostrum can be repeatedly bent, without evident damage, and despite being composed of the same material as other rigid body parts \cite{Jansen2016, Singh2016, Gibson1969, Toju2005}.
	By maintaining constant pressure on the snout and rotating around the bore-hole, females are able to flex the rostrum into a straightened configuration and produce a linear channel into the fruit.
	A single adult female may prepare hundreds of such sites \cite{Toju2005, Moffett1989, AguirreUribe1978}.

	Despite its documented performance in many \textit{Curculio} species \cite{Gibson1969, Moffett1989, AguirreUribe1978, Toju2005}, it remains unknown how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
	In this study, we therefore characterize the composite profile of the rostral cuticle to account for the observed flexibility of the snout.
	We show that the relative layer thicknesses and fiber orientation angles of the exocuticle and endocuticle of the rostrum are strongly differentiated from the head capsule and other body parts.
	The effect of these differences on the elasticity of the cuticle is estimated using Classical Laminate Plate Theory (CLPT).
	Recent studies have shown that the yield strength of the beetle exoskeleton is lower in tension than compression \cite{Longhai2017}.
	To assess the validity of these findings, we compare the ultimate tensile strength of the rostrum across species and snout morphotypes.
	We also report displacement-controlled load cycling results for the snout of one \textit{Curculio} species with strongly curved morphology.
	Our results indicate that an increase in the volume fraction of endocuticle in the rostrum conveys higher tensile strength at the rostral apex across all tested species; and further, that a strongly curved rostrum can be flexed repeatedly without harm to the structure.

	\begin{figure*}
		\centering
		\includegraphics[width=180mm]{fig1.pdf}
		\caption{\textbf{Morphology and oviposition behavior of female \textit{Curculio} weevils.} 
			\textbf{a}, Lateral habitus image of female \textit{Curculio~sayi} (Gyllenhal, 1836) featuring the elongate, strongly curved rostrum.
			\textbf{b}, Lateral view of the head of a female specimen of \textit{Curculio~longinasus} Chittenden, 1927, with major anatomical features indicated.
			\textbf{c}, Illustration of the oviposition behavior, proceeding from left to right: a female makes an incision in the host fruit, flexes the head directly over the bore hole using the front legs, then maintains tension on the snout while rotating to excavate a linear channel into the fruit. During this process, the female rostrum is bent until completely straight.
		}
		\label{fig::curculio}
	\end{figure*}

	\textit{\textbf{[NMF note: Feel that intro could be shortened by not "giving away" so many results. Maybe 3-5 phrases less in 2 final paragraphs of intro]}}
	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition.
	[Based on our findings, we posit that the composite profile of the rostral apex enables the snout to undergo repeated distal flexion while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.]
	Accordingly, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	This is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (though see \cite{Matsumura2017}).

	\section{Microstructure of the \textit{Curculio} rostrum}
		In arthropods -- including beetles - the exocuticle is comprised of numerous unidirectional laminae of chitin nanofibrils.
		Each layer is the thickness of a single fiber (2-4 \,nm) embedded in a proteinaceous matrix \cite{Nikolov2010, Nikolov2011}.
		These layers are stacked at a more or less constant angle to each other, forming a quasi-isotropic laminate known as the Bouligand structure \cite{Blackwell1980, Bouligand1972, Neville1976}. 
		This layout effectively mitigates the strong anisotropy of \textalpha-chitin, thus yielding a versatile building material for the exoskeleton \cite{Vincent1982, Vincent2004, Nikolov2010, Nikolov2011}.		
		Beetle endocuticle, however, is unique among arthropods and is comprised of large -- 1-5 \,{\textmu}m diameter in \textit{Curculio} -- unidirectional bundles of chitin, called macrofibers.
		Chitin macrofibers are orthotropic (axial: $E_1~=~8.5~\,\text{GPa}$; transverse: $ E_2~=~E_3~=~0.52~\,\text{GPa}$ \cite{Jansen2016}), and arranged in unidirectional plies, as depicted in Figs.~\ref{fig::cuticle},~\ref{fig::profile} \cite{Kamp2010, Kamp2015}.
		Typically, adjacent macrofiber laminae are paired and pseudo-orthogonal -- i.e., angled nearly $90\degree$ to each other (see Fig.~\ref{fig::profile}; \cite{Cheng2009}) -- with a constant stacking angle \emph{between} pairs, although other configurations have been observed \cite{Hepburn1973, Kamp2010, Kamp2015, Leopold1992}.
		This geometric sequence of the macrofiber laminae yields an approximately transversely isotropic composite, similar to the Bouligand structure \cite{Kamp2015, Nikolov2010}.
		Notably, the resulting laminate is less rigid than the exocuticle, but exhibits greater toughness because the pseudo-orthogonal plies effectively inhibit crack formation and propagation between successive layers \cite{Kamp2010, Kamp2015, Hepburn1973}.

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig2.pdf}
			\caption{\textbf{Gross divisions of cuticle in the female \textit{Curculio} rostrum.}
				\textbf{a}, Scanning electron micrograph of the head capsule, in frontal view, of female \textit{Curculio~sulcatulus} (Casey, 1897), with the rostrum removed.
				\textbf{b}, Magnified view of the junction between the rostrum and head capsule, showing the division of the cuticle into two general regions: the exocuticle and the endocuticle.
			}
			\label{fig::cuticle}
		\end{figure*}

		Serial thin sectioning and scanning electron microscopy of fractured \textit{Curculio} specimens reveals that endocuticle in the head capsule fits this general profile, with an angle of nearly $30\degree$ between successive pairs of pseudo-orthogonal plies.
		Additionally, in the head capsule, the thickness of the exocuticle and endocuticle in cross-section is nearly equal -- typically between 20-30 \,{\textmu}m.
		However, the cuticle composite lay-up of the rostral apex is strongly differentiated from the head capsule, as shown in Fig.~\ref{fig::profile}.
		Distally the exocuticle is reduced to a thin shell (ca. 5 \,{\textmu}m), with the endocuticle thickened to offset this reduction and maintain a constant cuticle thickness (ca. 50 \,{\textmu}m total) throughout its length.
		Moreover, the endocuticular macrofibers exhibit no rotation between successive pseudo-orthogonal plies; and are oriented at approximately $\pm45\degree$ to the longitudinal axis of the snout (i.e., an antisymmetric $[\pm45\degree]$ angle-ply laminate).
		We previously identified these modifications to the composite structure of the cuticle within a single species, \textit{Curculio longinasus} Chittenden, 1927 \cite{Jansen2016, Singh2016}.
		However, this composite profile is herein reported in the rostrum of six additional, phylogenetically disjoint, species, suggesting that this is an evolutionarily conserved trait throughout the genus \textit{Curculio}.
		In all examined species, the portion of the snout between the head capsule and apex of the scrobe exhibits a gradual transition in composite profile along an anterior-posterior gradient.

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig3.pdf}
			\caption{\textbf{Composite profiles of rostral \textit{Curculio} cuticle.}
				\textbf{a}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~humeralis} (Casey, 1897), showing that the cuticle of the rostral apex is organized as a $\pm45\degree$ angle-ply laminate.
				\textbf{b}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~caryae}  (Horn, 1873), showing that the cuticle of the rostral base and head capsule has an approximately $30\degree$ stacking angle between each pair of pseudo-orthogonal plies.
				\textbf{c--d}, Semi-thin sections of cuticle from a specimen of \textit{Curculio humeralis}, stained with toluidine-blue-borax; demonstrating:
				(\textbf{c}) that the exocuticle of the rostral apex is reduced to a thin shell close to 5 \,{\textmu}m thick, with the endocuticle thickened to maintain a constant laminate thickness; and,
				(\textbf{d}) that the exocuticle of the head capsule and base of the snout occupies nearly half of the through-thickness of the cuticle.
			}
			\label{fig::profile}
		\end{figure*}

		Here we estimate the effect of differential cuticle organization on uni-axial membrane and on transverse-flexural Young's moduli of the cuticle in the rostral apex and head capsule using CLPT \cite{Reddy2004, Jones2014}.
		The effective elastic constants of the cuticle regions of \textit{Curculio longinasus} -- estimated previously \cite{Jansen2016} -- are used to construct constitutive equations for the entire cuticle of that species.
		The cuticle of the head capsule is estimated to have membrane and flexural moduli of $E_m~=~4.77~\,\text{GPa}$ and $E_f~=~6.04~\,\text{GPa}$, respectively.
		In the rostral apex these values are reduced by approximately 72\% and 60\%, respectively ($E_m~=~1.36~\,\text{GPa}; E_f~=~2.44~\,\text{GPa}$).
		Two hypothetical cuticle lay-ups are also modeled to individually assess the contributions of either modified layer thickness or stacking angle sequence to the cuticle's flexibility.
		The effective moduli of a configuration with the angle stacking sequence of an otherwise typical cuticle (i.e., in the head capsule) that shows layer thicknesses of the rostral apex are calculated as $E_m~=~3.73~\,\text{GPa}, E_f~=~4.31~\,\text{GPa}$; representing 22\% and 29\% decreases from unmodified cuticle, respectively.
		Similarly, a hypothetical cuticle with the layer thicknesses of ordinary cuticle yet possessing the angle stacking sequence of the rostral apex (i.e., $\pm45\degree$ angle-ply in the endocuticle) has effective elastic moduli of $E_m~=~3.77~\,\text{GPa}, E_f~=~5.76~\,\text{GPa}$; representing 21\% and 4.7\% decreases from unmodified cuticle, respectively.
		Each of the cuticle modifications noted in the rostral apex individually decrease the elastic moduli of the cuticle.
		However, they appear to have a synergistic combined effect on cuticle elasticity, rather than a simple additive effect.
		This result suggests that both modifications are necessary in order for the snout to function properly in the living individual -- where the combined effect allows the rostrum to bend until completely straight without fracture.

	\section{Force-controlled loading to fracture}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig4.pdf}
			\caption{\textbf{Tensile properties of the female \textit{Curculio} rostrum.}
				Each plot shows the relationship between two variables as predicted by a phylogenetic linear mixed-effect model; with species as a random effect and a variance-covariance matrix generated from Brownian motion over the preferred phylogeny of \cite{Bonal2016}.
				Gray regions represent the prediction interval and bootstrapped 95\% confidence interval of the model.
				The estimated fixed effect $\hat{\beta}$ is given, along with the p-value of a t-test assessing whether $\hat{\beta}$ is significantly different from zero.
				A generalized marginal $\text{R}^2_{\beta^*}$ for assessing fixed effects is also reported.
				In general, increased endocuticle thickness is associated with greater tensile strength, and stiffness is inversely correlated with toughness.
			}
			\label{fig::tensile}
		\end{figure*}

		To better characterize the failure behavior of the female rostrum, we performed tensile testing on the snouts of six \textit{Curculio} species that representing a mixture of closely and distantly related taxa \cite{Hughes2004phylo, Hughes2004eco, Bonal2016, Bonal2011}.
		Each specimen was first immersed in $\text{di-H}_2\text{O}$ for 24 hours to simulate the living tissue (see \cite{Klocke2011}), then subjected to force-controlled, uniaxial loading to fracture at a constant stress rate of $1.0\,\text{gf}\cdot\text{s}^{-1}$.
		In general, the specimens exhibited a non-linear viscoelastic response curve characterized by a sharp increase in stress at higher strains, terminating in brittle fracture \cite{Mihai2017}.
		We suggest that strain hardening occurs as the longitudinal axis of the macrofibers becomes more closely aligned to the cylindrical axis of the rostrum, thereby resisting tension more directly with increasing strain \cite{Munster2013}.

		We also examined the correspondence between composite structure and mechanical behavior of the snout in an evolutionary, comparative context.
		Phylogenetic linear mixed effects models (PGLM, Fig.~\ref{fig::tensile}) were used to account for phylogenetic non-independence in residual variance, with species membership included as a random effect \cite{Revell2010, Felsenstein1985}.
		The resulting models show that the maximum force sustained at the site of failure is strongly correlated with the cross-sectional area of the endocuticle ($\hat{\beta}=1.28,~p<0.0001$), and not the exocuticle ($\hat{\beta}=0.17,~p=0.38$), at that site.
		There is thus a negative correlation between ultimate tensile strength of the specimen and the cross-sectional exocuticle-to-endocuticle area ratio at the fracture site ($\hat{\beta}=-0.36,~p=0.017$).
		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no evidence of correspondence between the cross-sectional properties of the fracture site and the overall performance of the rostrum.
		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to properties of the entire rostrum.
		\textit{\textbf{[NMF note: How surprising/unexpected is this? If a lot -- state so..]}}

		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness (see Fig.~\ref{fig::tensile}).
		This implies that stiffer specimens -- and by extension, stiffer cuticle profiles -- are generally more brittle.
		We also observed a moderate, but not statistically significant, stiffening size-effect with respect to rostral length.
		This runs contrary to our expectation that a longer, more strongly curved rostrum would require increased flexibility to avoid fracture during oviposition \cite{Hughes2004eco, Bonal2011}.
		It is possible that longer rostra also have a longer transition gradient from basal to apical profile; thereby reinforcing the junction between the rostrum and head capsule against buckling.
		Young's modulus of the rostrum would be comparatively higher in such species, due to the higher volume fraction of exocuticle.
		We therefore posit that the gross elastic performance of the cuticle is consistent across the weevil genus \textit{Curculio}.
		A \emph{single} mechanism -- i.e., the modified composite profile -- likely confers increased flexibility and tensile strength to the rostral apex across all \textit{Curculio} species.
		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle; likely because of its organization into large bundles of aligned, anisotropic fibers, and amounting a trade-off between rigidity and toughness.
		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture-resistant.

	\section{Load cycling of the pecan weevil}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig5.pdf}
			\caption{\textbf{Fatigue testing of a female \textit{Curculio caryae} rostrum.}
				\textbf{a}, Set-up of fatigue testing, with the female's head capsule fixed to a pedestal and the rostrum attached to a strip of rip-stop nylon fabric using cyanoacrylate adhesive.
				The specimen was loaded in tension, as opposed to compression, therefore isolating the effect of tension on the fatigue behavior of the rostrum.
				\textbf{b}, Overlay of the pre- and post-fatigue states of the head, showing a clear (short-term) effect from repeated prescribed strain.
				\textbf{c}, Three-photograph sequence of pre- and post-fatigue states of the rostrum; with the left and central panels showing the conditions immediately prior to and after the testing, respectively, whereas the right panel shows the head having returned to its original shape after 24 hours off soaking in a water/ethanol mixture. Hence the immediate post-fatigue shape is not permanent.
				\textbf{d}, Raw force data (left) and displacement data (right) plots for the fatigue test.
				The displacement data plot shows clear viscoelastic behavior, indicated by histeresis in the stress-strain response of the specimen, and exhibits a logarithmic decrease stress amplitude over time.
			}
			\label{fig::fatigue}
		\end{figure*}

		To confirm that repeated, prescribed straightening of the rostrum does not result in damage to the cuticle, we performed displacement-controlled fatigue testing on a typical female specimen of the pecan weevil \textit{Curculio caryae} (Horn, 1873) -- a species that exhibits extreme ($80-90\degree$, Fig.~\ref{fig::fatigue}) rostral curvature \cite{AguirreUribe1978, Gibson1969}.
		The specimen was aligned so that uniaxial tension would induce elongation of the distal portion of the rostrum with minimal off-axis deflection of the uncurved section.
		The strain per cycle was fixed at an amplitude sufficient to completely elongate the rostrum and generate a tensile load of 1.0\,N  in the straightened configuration (ca. 20\% ultimate strength), at a frequency of $0.33\,\text{Hz}$.
		The test was terminated after a period of two weeks -- i.e., ca. 400,000 cycles -- when the stress amplitude appeared to reach an asymptotic minimum.		
		The rostrum behaved viscoelastically, indicated by histeresis in the stress-strain relationship during each cycle.
		Strain amplitude decreased logarithmically with cycle number, and the specimen appeared to have been deformed plastically and permanently during the test.
		However, after cleaning the specimen in a 24-hour wash with ethanol and water, the rostrum had returned to its original shape (Fig.~\ref{fig::fatigue}).

		While we cannot fully determine the cause for rostrum stress relaxation \emph{after} testing, we speculate that it arose from the general mechanism associated with cuticle viscoelasticity.
		The endocuticle is made of aligned \textalpha-chitin nanofibrils whose crystalline structure is enforced by hydrogen bonds between individual chitin chains and through the protein matrix along their length.
		Macroscopic viscoelastic behavior results from slippage between these chains in response to shearing between the chitin molecules \cite{Vincent2004, Evans1967, Sun2012}.
		Repeated strain may have caused such slippage in the endocuticle of the rostral apex during the fatigue test
		Without sufficient time for the material to \textit{completely} relax after deformation, the rostrum would slowly accumulate strain and deform viscoelastically \cite{Munster2013}.
		After immersion in ethanol and water, however, the cuticle would be sufficiently plasticized to allow the rostrum to return to its original configuration, thus dissipating the accumulated strain.

		The specimen did not show any evidence of fractures, micro-tears, or shear cusps anywhere in the surface of the exocuticle.
		Moreover, the tensile strength of the tested specimen's rostrum was consistent with that of other species members ($F_max~=~5.02~\,\text{N}$).
		Surprisingly, the specimen remained undamaged by the testing.
		We therefore conclude that under normal life conditions, repeated bending of the rostrum does not exceed the yield strength of the cuticle.

	\section{Fractography of \textit{Curculio} test specimens}

		\begin{figure*}
			\centering
			\includegraphics[width=180mm]{fig6.pdf}
			\caption{\textbf{Fractography of the female \textit{Curculio} rostrum.}
				\textbf{a}, Light micrograph showing the fracture surface of a tensile tested \textit{Curculio caryae} rostrum, displaying a typical failure mode; illustrated in (\textbf{b}):
				The small, black arrows indicate the winding direction of the macrofiber lamina; green arrows and dotted lines indicate the direction of crack.
				Scanning electron micrographs highlight: (\textbf{c}) shear cusp formation; (\textbf{d}) tensile failure and off-axis macrofiber fiber-pulling; (\textbf{e}) interlaminar delamination; (\textbf{f}-\textbf{h}) crack formation near invaginated exocuticle; and (\textbf{g}) ventral crack-front coalescence and shear cusp formation.
			}
			\label{fig::fracture}
		\end{figure*}

		In light of the complex failure modes evident in the fractured specimens, it was not always possible to identify void nucleation and crack initiation sites.
		We observed several patterns characteristic of both the micro-scale behavior of the cuticle and the meso-scale behavior of the rostrum during uniaxial tensile failure (Fig.~\ref{fig::fracture}).
		These patterns are described below.

		In transverse view, the exocuticle consistently presented a nearly continuous fracture surface.
		This is characteristic of comparatively brittle failure -- presumably due to the relatively homogeneous arrangement of \textalpha-chitin laminae in the Bouligand structure \cite{Nikolov2010, Nikolov2011}.
		The exocuticle typically appeared to fracture at lower strains than the endocuticle, with shear-cusp formation evident both at the fracture surface and across exocuticle adjacent to the plane of fracture \cite{Greenhalgh2009}.
		Conversely, the endocuticle exhibited severe delamination, off-axis ply-splitting, and fiber-pulling away from the fracture surface.
		This is indicative of the relatively high toughness of the unidirectional \textalpha-chitin organization within the macrofibers \cite{Kamp2010, Kamp2015}.
		Because the exocuticle of weevils is anchored to the endocuticle by cross-linking fibers (see \cite{Kamp2015, Longhai2017}), exocuticular shear-cusp formation in uniaxial tension suggests extension-shear coupling within individual endocuticle laminae; and further implies that ply-splitting occurred via mode II fracture between macrofibers at high strain \cite{Jones2014, Reddy2004}.
		We hypothesize that intra-laminar extension-shear coupling also yielded off-axis, in-plane resultant forces as a function of lamina orientation angle.
		Mode III shearing then occurred between laminae with opposing in-plane resultant forces, causing the observed inter-ply delamination.
		Tensile failure of the macrofiber laminae would ultimately occur via mixed-mode I/II -- i.e., transverse tension/intra-laminar shear -- fracture, due to an increase in applied stress caused by ply-splitting in adjacent laminae \cite{Greenhalgh2009}.

		At the meso-scale, most specimens fractured along a single plane across and between the occipital sulci, which are cuticular invaginations that traverse the entire length of the rostrum \cite{Davis2014, Dennell1942}.
		These sulci increase the volume fraction of exocuticle in the ventral part of the snout.
		They contain large interfaces ideal for void nucleation (Fig.~\ref{fig::fracture}).
		The exocuticle of the occipital sulci usually displayed shear cusps oriented outward from the center of the invagination, and continuing dorsolaterally and ventromedially.
		Ventrally, the cusps converged toward a prominent scarp where the crack fronts joined.
		This scarp was often obscured by delaminated endocuticular macrofibers in specimens with large cross-sectional areas of endocuticle,

		The first layer of endocuticle usually fractured along the same plane as the exocuticle.
		Ventrally, the endocuticle laminae typically converged toward a scarp-like region characterized by severe delamination and numerous de-bonded macrofibers.
		Moreover, macrofibers aligned with the direction of crack propagation exhibited extensive ply-splitting with intermittent transverse shearing.
		In contrast, macrofibers oriented against the direction of crack propagation primarily displayed fracture by transverse shear along the plane of ply-splitting in adjacent laminae.
		Because the laminae form a cylinder, contralateral fibers in the same lamina display opposing fracture modes.
		The ventrolateral surfaces often exhibited extensive inter-ply delamination and fiber de-bonding in scarp-like prominences.
		This is likely due to a combination of tensile failure and shearing along the dorsally-radiating crack front.
		Dorsally, the coalescent crack fronts often caused significant de-bonding and ply-splitting, followed by broom-like tensile failure.
		In some specimens, the contralateral crack fronts were out of plane and coalesced via transverse shear through a large dorsal section of cuticle.

		Based on these failure patterns, we hypothesize that the exocuticle-rich occipital sulci are the most likely site for the initiation of void nucleation and catastrophic failure of the integrated rostral cuticle in cross section, as illustrated in Fig.~\ref{fig::fracture}.
		Structural failure would take place as cracks propagate through the endocuticle from these sutures, and ultimately penetrating the entire thickness of the laminate \cite{Davis2014}.
		Although other, more complex failure modes have been observed, we posit that in live specimens this is the most likely mechanism of tensile failure because typical bending behavior generates tension \emph{only} along the ventral surface of the rostrum.

	\section{Conclusions}
		The rostrum of \textit{Curculio} is characterized by a discontinuous composite profile.
		The cuticle is strongly differentiated in terms of relative layer thicknesses and orientation angles along an anterior-posterior gradient.		
		These modifications are sufficient to achieve a marked reduction in the effective membrane and flexural moduli of the cuticle -- 72\% and 60\%, respectively -- in constitutive models based on CLPT, thereby accounting for the observed flexibility of the rostral apex in live specimens.
		However, the reductions can only be realized with \emph{both} modifications to the cuticle, which have a non-additive effect on cuticle elasticity.
		\textit{Curculio} females require both modifications to function properly during oviposition.

		Likewise, tensile and fatigue testing reveal a trade-off between stiffness and fracture resistance -- measured by ultimate strain and toughness -- mediated by the relative proportion of endocuticle in the laminate.
		The altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture resistant, permitting the structure to be flexed without exceeding the elastic limits of the cuticle.

		This is to our knowledge the first time in arthropods that the composite profile of the cuticle has been related to a gradient in elasticity and tensile performance across a cuticular structure.
		Because these associations are independent of species membership, we posit that the behavior of the cuticle is consistent across the genus.
		Rostral flexibility is achieved exclusively in all \textit{Curculio} species through a modified cuticle lay-up.
		This inference raises the intriguing possibility that a single ancestral shift in cuticle organization at the rostral apex -- yielding higher flexibility and tensile strength -- enabled the evolutionary "exploration" of a large morphospace region, promoting the the high species-level diversity of this lineage.

		Based on fractographic analysis of the test specimens, we infer that the exocuticle exhibits brittle fracture at a comparatively low strain, due to shearing between the endocuticle macrofibers to which it is anchored.
		These macrofibers fail at higher strain, mediated by mixed-mode shearing and tensile fracture within and between laminae.
		This outcome is consistent with behavior shown in previous studies, as well as theoretical consideration of cuticle microstructure in CLPT.
		The latter predicts extension shear-coupling ($A_{16}, A_{26}\neq{0}$) for individual off-axis macrofiber laminae \cite{Jones2014, Reddy2004}.

		Our results imply that fracture initiation occurs in the comparatively brittle exocuticle.
		The reduction in exocuticle thickness in the rostral apex might serve to mitigate crack formation in rostral bending.
		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
		From an evoluationary perspective, these findings reveal an unexpected morphological source of constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion.
		The doubly-thick exocuticle in the invagination thus creates an unavoidable, brittle weak-point in an otherwise endocuticle-dominated rostral apex.
		This constraint -- in conjunction with the minimization of exocuticle thickness in the rostral apex and the increased toughness derived from a thickened endocuticle -- lead us to consider that avoidance of catastrophic structural failure has been a driving selective pressure in the evolution of the female \textit{Curculio} rostrum.

	\section{Methods}
		Methods, including statements of data availability and any associated accession codes and references, are available in the online version of this paper.

	\begin{thebibliography}{50}
		\section*{References}
			\bibitem{Jansen2016}
				Jansen, M. A., Singh, S. S., Chawla, N., \& Franz, N. M.
				A multilayer micromechanical model of the cuticle of \textit{Curculio longinasus} Chittenden, 1927 (Coleoptera: Curculionidae).
				\textit{J. Struct. Biol.}
				\textbf{195: 2,}
				139--158,
				(2016).

			\bibitem{Vincent2004}
				Vincent, J. F. V. \& Wegst, U. G. K.
				Design and mechanical properties of insect cuticle.				
				\textit{Arthropod Struct. Dev.}
				\textbf{33:3,}
				187--199,
				(2004).

			\bibitem{Hepburn1973}
				Hepburn, H. R. \& Ball, A.
				On the structure and mechanical properties of beetle shells.
				\textit{J. Mater. Sci.}
				\textbf{8:5,}
				618--623,
				(1973).

			\bibitem{Kamp2010}
				van de Kamp, T. \& Greven, H.
				On the architecture of beetle elytra.
				\textit{Entomol. Heute}
				\textbf{22,}
				191--204,
				(2010).

			\bibitem{Kamp2015}
				van de Kamp, T., Riedel, A. \& Greven, H.
				Micromorphology of the elytral cuticle of beetles, with an emphasis on weevils (Coleoptera : Curculionoidea)
				\textit{Arthropod Struct. Dev.}
				\textbf{45:1,}
				14--22,
				(2016).

			\bibitem{Neville1976}
				Neville, A. C., Parry, D. A. \& Woodhead-Galloway, J.
				The chitin crystallite in arthropod cuticle.
				\textit{J. Cell Sci.}
				\textbf{21:1,}
				73--82,
				(1976).

			\bibitem{Amini2015}
				Amini, S., Tadayon, M., Idapalapati, S., and Miserez, A.
				The role of quasi-plasticity in the extreme contact damage tolerance of the stomatopod dactyl club
				\textit{Nat. Mater.}
				\textbf{14:9,}
				943--950,
				(2015).

			\bibitem{Mccullough2014struc}
				McCullough, E. L., Tobalske, B. W., and Emlen, D. J.
				Structural adaptations to diverse fighting styles in sexually selected weapons
				\textit{Proc. Natl. Acad. Sci. USA}
				\textbf{111:40,}
				14484--14488,
				(2014).

			\bibitem{Mccullough2014mech}
				McCullough, E. L.
				Mechanical limits to maximum weapon size in a giant rhinoceros beetle
				\textit{Proc. R. Soc. B}
				\textbf{281,}
				20140696,
				(2014).

			\bibitem{Dirks2013}
				Dirks, J., Parle, E., and Taylor, D.
				Fatigue of insect cuticle
				\textit{J. Exp. Biol.}
				\textbf{216:10,}
				1924--1927,
				(2013).

			\bibitem{Dirks2012}
				Dirks, J. and Taylor, D.
				Fracture toughness of locust cuticle
				\textit{J. Exp. Biol.}
				\textbf{215:9,}
				1502--1508,
				(2012).

			\bibitem{ICZN1999}
				\textit{International Code of Zoological Nomenclature}
				Fourth Edition. 
				(The International Trust for Zoological Nomenclature,
				London,
				1999).

			\bibitem{Davis2014}	
				Davis, S. R.
				\textit{Morphology, phylogeny, and evolutionary development in the weevils (Insecta: Coleoptera: Curculionoidea)}
				(Ph.D. thesis,
				University of Kansas,
				2014).

			\bibitem{Singh2016}
				Singh, S. S., Jansen, M. A., Franz, N. M., and Chawla, N.
				Microstructure and nanoindentation of the rostrum of Curculio longinasus Chittenden, 1927 (Coleoptera: Curculionidae)
				\textit{Mater. Charact.}
				\textbf{118,}
				206--211,
				(2016).

			\bibitem{Toju2005}
				Toju, H. and Sota, T.
				Imbalance of predator and prey armament: geographic clines in phenotypic interface and natural selection
				\textit{Am. Nat.}
				\textbf{167:1,}
				105--117,
				(2006).

			\bibitem{Gibson1969}	
				Gibson, L. P.
				Monograph of the genus \textit{Curculio} in the New World (Coleoptera: Curculionidae). Part I. United States and Canada
				\textit{Misc. Publ. Entomol. Soc. Am.}
				\textbf{6,} 
				239-–285,
				(1969).

			\bibitem{Dennell1942}
				Dennell, R.
				The structure and function of the mouth-parts, rostrum and fore-gut of the weevil \textit{ Calandra granaria} L.
				\textit{Phil. Trans. R. Soc. Lond. B}
				\textbf{231:581,}
				247--291,
				(1942).

			\bibitem{Morimoto2003}
				Morimoto, K. and Kojima, H.
				Morphologic characters of the weevil head and phylogenetic implications (Coleoptera, Curculionoidea)
				\textit{Esakia}
				\textbf{43,}
				133--169,
				(2003).

			\bibitem{Ting1933}
				Ting, P. C.
				Feeding mechanisms of weevils, their function, and relationship to classification
				\textit{Mon. Bull. Dep. Agric. State Calif.}
				\textbf{22,}
				161–-165,
				(1933).

			\bibitem{Ting1936}
				Ting, P. C.
				The mouth parts of the coleopterous group Rhynchophora
				\textit{Microentomology}
				\textbf{1,}
				93–-114,
				(1936).

			\bibitem{AguirreUribe1978}
				Aguirre Uribe, L. A.
				\textit{Biology of the immature stages of the pecan weevil \textnormal{Curculio caryae} (Horn) and oviposition habits of the adult weevil}
				(Ph.D. thesis,
				Texas A \& M University,
				1978).

			\bibitem{Moffett1989}
				Moffett, M.
				Life in a nutshell
				\textit{Natl. Geogr. Mag.}
				\textbf{176,}
				783-–784,
				(1989).

			\bibitem{Longhai2017}
				Longhai, L. et al.
				Microstructure and mechanical properties of rostrum in Cyrtotrachelus longimanus (Coleoptera: Curculionidae)
				\textit{Anim. Cells Syst.}
				\textbf{21:3,}
				199--206,
				(2017).

			\bibitem{Matsumura2017}
				Matsumura, Y., Kovalev, A. E., and Gorb, S. N.
				Penetration mechanics of a beetle intromittent organ with bending stiffness gradient and a soft tip,
				\textit{Sci. Adv.}
				\textbf{3:12,}
				eaao5469,
				(2017).

			\bibitem{Nikolov2011}
				Nikolov, S. et al.
				Robustness and optimal use of design principles of arthropod exoskeletons studied by ab initio-based multiscale simulations.
				\textit{J. Mech. Behav. Biomed. Mater.}
				\textbf{4:2,}
				129--145,
				(2011).

			\bibitem{Nikolov2010}
				Nikolov, S. et al.
				Revealing the design principles of high-performance biological composites using Ab initio and multiscale simulations: The example of lobster cuticle.
				\textit{Adv. Mater.}
				\textbf{22:4,}
				519--526,
				(2010).

			\bibitem{Blackwell1980}
				Blackwell, J. and Weih, M.
				Structure of chitin-protein complexes: ovipositor of the ichneumon fly \textit{Megarhyssa}.
				\textit{J. Mol. Biol.}
				\textbf{137:1,}
				49--60,
				(1980).

			\bibitem{Bouligand1972}
				Bouligand, Y.
				Twisted fibrous arrangements in biological materials and cholesteric mesophases.
				\textit{Tissue Cell}
				\textbf{4:2,}
				189--217,
				(1972).

			\bibitem{Vincent1982}
				Vincent, J. F. V.
				\textit{Structural biomaterials}
				(Halsted Press,
				New York,
				1982).

			\bibitem{Cheng2009}
				Cheng, L., Wang, L., \& Karlsson, A. M.
				Mechanics-based analysis of selected features of the exoskeletal microstructure of \textit{Popillia japonica}.
				\textit{Mater. Res.}
				\textbf{24:11,}
				3253--3267,
				(2009).

			\bibitem{Leopold1992}
				Leopold, R. A., Newman, S. M., and Helgeson, G.
				A comparison of cuticle deposition during the pre-and posteclosion stages of the adult weevil, Anthonomus grandis BOHEMAN (Coleoptera: Curculionidae)
				\textit{Int. J. Insect Morphol. and Embryol.},
				\textbf{21:1,}
				37--62,
				(1992).

			\bibitem{Reddy2004}
				Reddy, J. N.
				\textit{Mechanics of laminated composite plates and shells: theory and analysis}
				(CRC Press,
				Philadelphia,
				2004).

			\bibitem{Jones2014}
				Jones, R. M.
				\textit{Mechanics of composite materials}
				(CRC press,
				Philadelphia,
				2014).

			\bibitem{Hughes2004phylo}
				Hughes, J. and Vogler, A. P.
				The phylogeny of acorn weevils (genus Curculio) from mitochondrial and nuclear DNA sequences: the problem of incomplete data
				\textit{Mol. Phylogenet. Evol.}
				\textbf{32:2,}
				601--615,
				(2004).

			\bibitem{Bonal2016}
				Bonal, R. et al.
				Diversity in insect seed parasite guilds at large geographical scale: the roles of host specificity and spatial distance
				\textit{J. Biogeogr.}
				\textbf{43:8,}
				1620--1630,
				(2016).

			\bibitem{Hughes2004eco}
				Hughes, J. and Vogler, A. P.
				Ecomorphological adaptation of acorn weevils to their oviposition site
				\textit{Evolution}
				\textbf{58:9,}
				1971--1983,
				(2004).

			\bibitem{Bonal2011}
				Bonal, R.,Espelta, J. M., and Vogler, A. P.
				Complex selection on life-history traits and the maintenance of variation in exaggerated rostrum length in acorn weevils
				\textit{Oecologia}
				\textbf{167:4,}
				1053--1061,
				(2011).

			\bibitem{Klocke2011}
				Klocke, D. \& Schmitz, H.
				Water as a major modulator of the mechanical properties of insect cuticle.
				\textit{Acta Biomater.}
				\textbf{7,}
				2935--2942
				(2011).

			\bibitem{Mihai2017}
				Mihai, L. A. and Goriely, A.
				How to characterize a nonlinear elastic material? A review on nonlinear constitutive parameters in isotropic finite elasticity
				\textit{Phil. Trans. R. Soc. Lond. A}
				\textbf{473:2207,}
				20170607,
				(2017).

			\bibitem{Munster2013}
				M{\"u}nster, S. et al.
				Strain history dependence of the nonlinear stress response of fibrin and collagen networks
				\textit{Proc. Natl. Acad. Sci. USA}
				\textbf{110:30,}
				12197--12202,
				(2013).

			\bibitem{Felsenstein1985}
				Felsenstein, J.
				Phylogenies and the comparative method
				\textit{Am. Nat.}
				\textbf{125:1,}
				1--15,
				(1985).

			\bibitem{Revell2010}
				Revell, L. J.
				Phylogenetic signal and linear regression on species data
				\textit{Methods Ecol. Evol.}
				\textbf{1:4,}
				319--329,
				(2010).

			\bibitem{Evans1967}
				Beament, J. W. L., Treherne, J. E., and Wigglesworth, V. B.
				\textit{Advances in insect physiology}
				Volume 4
				(Academic Press Inc.,
				London,
				1967).

			\bibitem{Sun2012}
				Sun, J. Y. et al.
				Differential constitutive equation of elytra cuticle by nanoindentation
				\textit{Adv. Mater. Res.},
				\textbf{343,}
				1133--1139,
				(2012).

			\bibitem{Greenhalgh2009}
				Greenhalgh, E.
				\textit{Failure analysis and fractography of polymer composites}
				(CRC Press,
				Boca Raton,
				2009).
	\end{thebibliography}

	\begin{acknowledgements}
		The authors are grateful to Robert Anderson (CMNC), Lourdes Chamorro (USNM), and Charlie O'Brien (CWOB) for their assistance and provision of \textit{Curculio} specimens used in this study.
		The authors thank Salvatore Anzaldo, Andrew Johnston, Brian Reilly, Sangmi Lee, and other Arizona State University Hasbrouck Insect Collection (ASUHIC) members for their assistance in procuring, treating, and maintaining specimen loans upon entry into the ASUHIC.
		The authors would also like to thank David Lowry (ASU CLAS Bioimaging Center) for his training and assistance with resin embedding, scanning electron microscopy, and microtomography; and Salvatore Anzaldo for his assistance with histological staining.
	\end{acknowledgements}

	\section*{Author contributions}
		M.A.J. conducted sectioning and staining, microscopy and imaging, tensile and fatigue testing, statistical analysis, and participated in manuscript preparation.
		J.W. conducted tensile and fatigue testing and participated in manuscript preparation.
		N.C. facilitated microscopy, tensile and fatigue testing, and participated in manuscript preparation.
		N.M.F. facilitated specimen acquisition and imaging and participated in manuscript preparation.

	\section*{Additional information}
		Supplementary information is available in the online version of the paper.
		Reprints and permissions information is available online at www.nature.com/reprints.
		Correspondence and requests for materials should be addressed to M.A.J.

	\section*{Competing financial interests}
		The authors declare no competing financial interests.

	\newpage

	\section*{Methods}
		\subsection*{Specimen acquisition and taxon sampling}
			Specimens for use in tensile and fatigue testing came from the Hasbrouck Insect Collection at Arizona State University [\textbf{\texttt{ASUHIC}}].  This set of specimens was supplemented with material housed in the following collections, using the codens of Arnett et al. \cite{Arnett1993}:

			\begin{description}
				\item [\texttt{CMNC}]  Canadian Museum of Nature Collection, Ottawa, Ontario, Canada
				\item [\texttt{USNM}]  National Museum of Natural History, Washington, D.C., USA
			\end{description}

			Cold fracture, semi-thin sectioning, and tensile testing were conducted on randomly chosen female specimens belonging to six \textit{Curculio} species obtainable through field work in the southwestern United States and northwestern Mexico.
			Taxon sampling was targeted to represent a mixture of disparate radiations and sister taxa with a variety of rostral morphotypesin accordance with the phylogenetic hypotheses of Hughes~et~al.~\cite{Hughes2004eco, Hughes2004phylo} and Bonal~et~al.~\cite{Bonal2016}.
			The six species of \textit{Curculio} used herein are (in alphabetical sequence): \textit{Curculio~caryae} (Horn, 1873), \textit{Curculio~humeralis} (Casey, 1897), \textit{Curculio~proboscideus} Fabricius, 1775, \textit{Curculio~sulcatulus} (Casey, 1897), \textit{Curculio~uniformis} (LeConte, 1857), and \textit{Curculio~victoriensis} (Chittenden, 1904).
			Specimens were identified to taxonomic (species) concepts using \cite{Gibson1969} and other resources.

		\subsection*{Histological sectioning}
			To illustrate the relative proportions of the cuticle regions in cross-section, serial semi-thin sectioning was conducted on exemplary female specimens of \textit{Curculio~humeralis} and \textit{Curculio~longinasus} Chittenden, 1927.
			Live specimens of both species were collected into 95\% ethanol for preservation.
			A female specimen was selected; and the rostrum was separated from the head capsule with a fine-edged razor blade.
			The apical $1/4^{\text{th}}$ of the rostrum was also removed and then discarded.
			The remaining portion of the rostrum and the head capsule were then embedded in EMbed812, as follows.

			The cuticle was first immersed in acetone for 24 hours, and then transferred to a 2:1 mixture of acetone to epoxy resin.
			Samples remained at $21\degree\text{C}$ for 12 hours on a shaker table to prevent hardening.
			They were then transferred into a 1:1 mixture of acetone to resin, followed by a 1:2 mixture (each for 12 hours and at $21\degree\text{C}$ on a shaker table), before finally being placed into a silicone mold with pure resin.
			The mold was placed	into an oven heated to $38\degree\text{C}$, and the resin was allowed to cure for 24 hours.
			The resulting blocks were machined to prepare the apical surface of each
			sample for microtomy.

			A Leica Ultracut R Microtome and diamond knife were used to expose a cross-section (transverse plane) of the apical and basal portions of the rostrum and to remove excess material.
			Semi-thin sections (0.5 \,{\textmu}m thick) were kept and stained with
			toluidine-blue-borax for light microscopy and imaging.

		\subsection*{Cold-fracture of specimens}
			Two pinned female specimens of each \textit{Curculio} species were selected at random and retained for cold-fracturing of the rostrum.
			The heads of the specimens were removed and cleaned using a 95\% ethanol solution and a thin paintbrush.
			Any muscles protruding from the occipital foramen were removed with a fine-edged razor blade.
			The antennae were removed directly using forceps to pull the scape (basal section of antenna) from the antennal insertion.
			Cleaned specimens were stored at $-80\degree\text{C}$ for 24 hours, then fractured using forceps over a chilled aluminum block.
			To fracture each specimen, the head capsule and rostrum were each gripped firmly in a pair of forceps.
			The forceps were then sharply rotated to fracture the base of the snout via dorsal flexion.
			The rostrum was fractured a second time, after separation from the head capsule, using the same procedure.
			The segmented specimens were then placed into individual glass vials to protect the fracture surfaces from contamination prior to microscopy.

		\subsection*{Tensile and fatigue testing}
			\paragraph*{Force-controlled loading to failure.}
				Five female specimens of each \textit{Curculio} species were randomly allocated for use in tensile testing.
				The head of each specimen was removed, cleaned, and prepared as described above in the cold-fracture protocol.
				To avoid destroying the delicate, brittle specimens when gripping the ends of each head, a method was devised to create solid handles that could be clamped tightly into grips without risk of damage to the cuticle.

				For each head, four $1\,\text{cm}^2$ strips of gaffer tape were cut; these were used as gripping and mounting points for the specimen.
				A strip of tape would be laid flat, with a large drop of cyanoacrylate glue placed upon the upturned surface.
				The curved portion of the snout was then placed into the drop, such that the straight portion of the rostrum was aligned perpendicular to the edge of the strip.
				Hardening of the cyanoacrylate effectively embedded the curved portion of the snout in a solid mass, isolating a straight section of the snout -- from the base to a point distad of the apex of the scrobe -- for testing.
				A second strip of tape was fixed over this mass with an additional layer of cyanoacrylate to provide a dorsal gripping surface for the mass.
				A small mark was made to indicate the extent of the head inside the mass.
				This embedding procedure was repeated for the head capsule, resulting in a finished specimen with anterior and posterior handles for testing.

				Prior to testing, each specimen was placed in de-ionized water for 24 hours to allow full saturation of the cuticle, simulating the condition of live tissue \cite{Klocke2011}.
				Once removed for testing, the specimen was gripped using the cyanoacrylate handles at the marked locations immediately beyond the anterior margin of the rostrum and the posterior margin of the head capsule
				The exposed section of the snout was coated in petroleum jelly using a cotton swab to prevent loss of moisture and stiffening of the specimen during the test.
				Specimens were loaded in a Tryton 250 Microforce Testing System equipped with a 5N load cell and mechanical clamp grip.
				All specimens were subjected to force-controlled uniaxial tension at a rate of $1.0 \,\text{gf}\cdot\text{s}^{-1}$ until failure, with a sampling interval of 0.1\,s.
				Engineering stresses ($\sigma_0=F/A_0$) and strains ($\epsilon_0=\Delta{l}/l_0$) were reported only for specimens that did not fracture due to strain accumulation at the interface between the rostrum and the cyanoacrylate handles.

			\paragraph*{Displacement-controlled cyclic loading.}
				To confirm that repeated, complete extension of a strongly curved rostrum would not result in fracture of the cuticle, a representative female specimen of \textit{Curculio~caryae} was allocated for fatigue testing.
				The head capsule of the specimen was fixed to a push-pin using cyanoacrylate glue.
				This served as a pedestal and gripping location for the posterior portion of the specimen.
				The apex of the rostrum was fixed to a strip of ripstop nylon fabric equal in length to the head, using cyanoacrylate glue.
				As with tensile testing, the specimen was placed in de-ionized water for 24 hours, then coated in petroleum jelly using a cotton swab immediately prior to load cycling.

				The end of the fabric was gripped and used to elongate the rostrum in tension by pulling on the fabric, thus isolating the effect of tension on the fatigue life of the rostral cuticle.
				In this way, the rostrum would return to its original configuration in a spring-like manner, as in living specimens, rather than being forced to return to the initial position.
				The rostrum was aligned such that complete elongation of the curved section would take place in tension, with minimal off-axis deflection of the un-curved section.
				The specimen was subjected to displacement-controlled loading sufficient to fully extend the rostrum and generate a load stress of 1\,N, or approximately 20\% of the tensile strength of the species average.
				Load cycling took place at a rate of 0.33Hz, and was continued for 14 days -- i.e., ~400,000 cycles -- until the tensile stress in the sample approached an asymptotic minimum.

				Once the test was concluded, the specimen was placed in a 50\% ethanol solution for 24 hours to clean the petroleum jelly from the rostrum.
				The specimen was examined for surface fractures and micro-tears, then subjected to tensile testing via the same protocol as the other specimens to assess whether the cuticle had begun to fatigue.

		\subsection*{Specimen imaging and microscopy}
			The fracture surfaces of cold-fractured specimens were examined using scanning electron microscopy to characterize the composite profile and microstructure of the rostrum.
			Fracture behavior of tensile testing specimens was assessed using both light microscopy and SEM to image the fracture surfaces of the specimens in transverse view.
			Electron microscopy was conducted using a JEOL JSM6300 scanning electron microscope.
			Light microscopy was conducted using a Leica M205 C stereomicroscope and attached computer running the software Leica Application Suite (LAS); as well as a Visionary Digital Passport II system using a Canon EOS Mark 5D II camera outfitted with interchangeable macro lenses.
			Specimen length, layer thicknesses, macrofiber orientation angles, and cross-sectional areas were measured in the LAS and in Adobe Illustrator, using pixel-wise measurements multiplied by a scaling factor for the image.
			%%%%%
		\subsection*{Constitutive modeling of the cuticle}
			\paragraph*{General approach.}
				The effective uni-axial membrane and transverse flexural elastic moduli of idealized cuticle organizations, representing both the rostral apex and head capsule, were estimated using Classical Laminate Plate Theory (CLPT).
				For general information on this approach see \cite{Jones2014, Reddy2004}.
				The composite profiles of both types of cuticle were idealized using the layer thicknesses and stacking sequences observed in \textit{Curculio~longinasus}.
				This particular species was chosen because we derived the effective elastic constants of the individual components of the cuticle in previous work \cite{Jansen2016, Singh2016}.
				In addition, \textit{Curculio~longinasus} exhibits a profile that is typical and representative for the genus \textit{Curculio}, based on examination of the six species used for tensile testing.

				For \textit{Curculio~longinasus} the total thickness of the cuticle in the head and rostrum is roughly 50 \,{\textmu}m, as in most specimens of the other examined species.
				In the head capsule, the exocuticle occupies between 30-50\% of the through-thickness of the laminate, with the remaining thickness nearly evenly divided between 12 layers of endocuticle.
				We use the maximum (50\% of through-thickness, or 25 \,{\textmu}m) for the model, since the cuticle appears to deviate from this value only in regions with sulci (grooves), pores, and other scattered features of the surface sculpture.
				The macrofiber laminae of the endocuticle were assigned a stacking sequence of $0\degree, 90\degree, 30\degree, -60\degree, 60\degree, -30\degree, 90\degree, 0\degree, -60\degree, 30\degree, -30\degree, 60\degree$; thereby representing pairs of orthogonal plies stacked at a constant rotation angle of $30\degree$, in approximation of the living tissue.

				In the rostral apex, the exocuticle is reduced to a thin shell approximately 5 \,{\textmu}m in thickness, or 10\% of the total cuticle thickness.
				The endocuticle displays a more complex pattern of layer thicknesses in the rostral apex than in the head capsule.
				Each of the eight outermost layers are of nearly equal thickness to the exocuticle (5 \,{\textmu}m), whereas the four innermost layers have a combined thickness equal to that of the exocuticle or to a single layer of outer endocuticle ($h_{outer}=5~\,\text{{\textmu}m}, h_{inner}=1.25~\,\text{{\textmu}m}$).
				The stacking sequence with respect to the longitudinal axis of the rostrum forms an antisymmetric angle-ply laminate of $\pm45\degree$.

				To assess the individual contributions of layer thickness and stacking angle sequence to cuticle flexibility in the rostral apex, two hypothetical cuticle lay-ups were modeled -- each with only one of the modifications present in the cuticle of the rostral apex.
				The first of these models has the layer thicknesses of the rostral apex, but fiber orientations of the head capsule; whereas the second has the fiber orientations of the rostral apex, but the layer thicknesses of the head capsule.

				Because these laminates are not symmetric, each has a bending-extension coupling matrix $[B]$ populated with non-zero terms; thus complicating the calculation of effective in-plane elastic moduli.
				To circumvent this difficulty and enable meaningful comparisons between each laminate, all of the lay-ups are reflected about their inner surface.
				This effectively doubles their thickness while producing a balanced, symmetric laminate with no coupling between bending and extension (i.e., $[B]=0_{3,3}$).
				Estimation of in-plane elastic constants from the extension($[A]$) and bending ($[D]$) matrices is described in detail below.
				The program Matlab R2018b was used to numerically evaluate the final values of the effective elastic constants \cite{Matlab}.

			\paragraph*{Classical Laminate Plate Theory.}
				We begin by calculating the 2D reduced stiffness matrix for each part of the cuticle.
				For orthotropic materials with the principal axes parallel to the ply edges, the reduced stiffness matrix is defined as follows:

				\begin{equation}
				[Q] =
					\begin{bmatrix}
						Q_{11} & Q_{12} & 0 \\
						Q_{21} & Q_{22} & 0 \\
						0 & 0 & Q_{66}
					\end{bmatrix}\,,
				\end{equation}

				where:

				\begin{equation}
				\begin{aligned}
					Q_{11} & = \frac{E_{1}}{1 - \nu_{12}\nu_{21}}\,, \\
					Q_{12} & = \frac{E_{1}\nu_{21}}{1 - \nu_{12}\nu_{21}} = Q_{21}\,, \\
					Q_{21} & = \frac{E_{2}\nu_{12}}{1 - \nu_{12}\nu_{21}} = Q_{12}\,, \\
					Q_{22} & = \frac{E_{2}}{1 - \nu_{12}\nu_{21}}\,, \\
					Q_{66} & = G_{12}\,.
				\end{aligned}
				\end{equation}

				For each layer $k$, the reduced stiffness matrix is transformed to account for the layer orientation angle $\theta$ within the laminate coordinate system, yielding a reduced transformed stiffness matrix according to:

				\begin{equation}
					[\bar{Q}] = [T]^{-1}[Q][T]^{-T}\,,
				\end{equation}

				where the transformation matrix $[T]$ is defined as:

				\begin{equation}
				[T] =
				 \begin{bmatrix}
					 \cos^2\theta & \sin^2\theta & 2\cos\theta\sin\theta \\
					 \sin^2\theta & \cos^2\theta & -2\cos\theta\sin\theta \\
					 -\cos\theta\sin\theta & \cos\theta\sin\theta & \cos^2\theta - \sin^2\theta
				 \end{bmatrix}\,.
				\end{equation}

				Using the lay-ups specified for each type of cuticle described above, we calculate the extensional stiffness matrix $[A]$, bending stiffness matrix $[D]$, and bending-extension coupling matrix $[B]$ for each laminate consisting of $n$ layers at a distance $z$ from the laminate mid-plane.
				The elements of these matrices can be identified according to:

				\begin{equation}
					\begin{aligned}
						A_{ij} & = \sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z_{k} - z_{k-1})\,, \\
						B_{ij} & = \frac{1}{2}\sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z^2_{k} - z^2_{k-1})\,, \\
						D_{ij} & = \frac{1}{3}\sum_{k = 1}^{n}(\bar{Q}_{ij})_k(z^3_{k} - z^3_{k-1})\,.
					\end{aligned}
				\end{equation}

				These stiffness matrices respectively relate vectors of resultant forces $\{N\}$ and bending moments $\{M\}$ to mid-surface strains and curvatures $\{\epsilon^{\degree}\}$ and $\{\kappa\}$ in the laminate, according to the following relationship:

				\begin{equation}
					\begin{Bmatrix}
						\{N\} \\
						\{M\}
					\end{Bmatrix}
					=
					\begin{bmatrix}
						[A] & [B] \\
						[B] & [D]
					\end{bmatrix}
					\begin{Bmatrix}
					\{\epsilon^{\degree}\} \\
					\{\kappa\}
					\end{Bmatrix}\,.
				\end{equation}

				For symmetric laminates, $[B] = 0_{3,3}$, and therefore:

				\begin{equation}
				\label{eqn::abbd}
					\begin{aligned}
						\{N\} & = [A]\{\epsilon^{\degree}\}\,, \\
						\{M\} & = [D]\{\kappa\}\,,						
					\end{aligned}\,,
				\end{equation}

				or, in expanded form:

				\begin{equation}
					\begin{aligned}
						\begin{Bmatrix}
							N_{xx} \\
							N_{yy} \\
							N_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							A_{11} & A_{12} & A_{16} \\
							A_{21} & A_{22} & A_{26} \\
							A_{61} & A_{62} & A_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\epsilon^{\degree}_{xx} \\
							\epsilon^{\degree}_{yy} \\
							\gamma^{\degree}_{xy}
						\end{Bmatrix}\,,
						\\
						\begin{Bmatrix}
							M_{xx} \\
							M_{yy} \\
							M_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							D_{11} & D_{12} & D_{16} \\
							D_{21} & D_{22} & D_{26} \\
							D_{61} & D_{62} & D_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\kappa_{xx} \\
							\kappa_{yy} \\
							\kappa_{xy}
						\end{Bmatrix}\,.
					\end{aligned}
				\end{equation}

				If we make the simplifying assumptions \cite{Jones2014, Reddy2004} that \textbf{(1)} the laminate experiences pure axial loading and transverse bending (i.e., $N_{yy} = N_{xy} = 0$ and $M_{yy} = M_{xy} = 0$, respectively), and \textbf{(2)} the laminate is a beam of sufficiently high aspect ratio to minimize the Poisson effect and anisotropic shear coupling (i.e., below we effectively let $A^*_{12} = A^*_{16} = 0$ and $D^*_{12} = D^*_{16} = 0$), then we can calculate the in-plane effective flexural and axial Young's moduli of the laminate along the x-axis.

				For axial Young's modulus of the laminate, we first define the average membrane stresses in the laminate as:

				\begin{equation}
					\{\bar{\sigma}^m\} = \frac{\{N\}}{z_1 - z_n}\,.
				\end{equation}

				By substitution in Eq.~\ref{eqn::abbd}, we obtain:

				\begin{equation}
					\begin{aligned}
						\begin{Bmatrix}
							\bar{\sigma}^m_{xx} \\
							\bar{\sigma}^m_{yy} \\
							\bar{\tau}^m_{xy}
						\end{Bmatrix}
						& = \frac{1}{(z_1 - z_n)}
						\begin{bmatrix}
							A_{11} & A_{12} & A_{16} \\
							A_{21} & A_{22} & A_{26} \\
							A_{61} & A_{62} & A_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\epsilon^m_{xx} \\
							\epsilon^m_{yy} \\
							\gamma^m_{xy}
						\end{Bmatrix}\,,
					\end{aligned}
				\end{equation}

				and, by inverting this equation (let $A^* = A^{-1}$) and substituting $A^*_{12} = A^*_{16} = 0$ based on the assumptions above, we infer:

				\begin{equation}
					\epsilon^m_{xx} = (z_1 - z_n)A^{*}_{11}\bar{\sigma}^m_{xx}\,.
				\end{equation}

				We therefore define Young's modulus for effective axial elasticity as:

				\begin{equation}
					E^m_{xx} = \frac{\bar{\sigma}^m_{xx}}{\epsilon^m_{xx}} = \frac{1}{(z_1 - z_k)A^*_{11}}\,.
				\end{equation}

				To find the transverse flexural Young's modulus of the laminate, we first specify the moment-curvature relation of an Euler-Bernoulli beam:

				\begin{equation}
				\label{eqn::bee}
				M = EI\kappa\,.
				\end{equation}

				Along the x-axis, the second moment of area for a rectangular cross-section is:

				\begin{equation}
					I_{yy} = \frac{b(z_1 - z_n)^3}{12}\,.
				\end{equation}

				Given the assumption that $M_{yy} = M_{xy} = 0$, the moment along the x-axis is related to the moment of the beam by:

				\begin{equation}
					M = M_{xx}b\,.
				\end{equation}

				Thus, given the assumption that $D^*_{12} = D^*_{16} = 0$, Young's modulus for the effective transverse flexural elasticity of the laminate can be found by making Eq.~\ref{eqn::bee} specific to transverse flexure of the x-axis and rearranging the terms:

				\begin{equation}
					E^f_{xx} = \frac{12M_{xx}}{(z_1 - z_n)^3\kappa_{xx}}\,.
				\end{equation}

				From inversion of Eq.~\ref{eqn::abbd} (let $D^* = D^{-1}$), this reduces to:

				\begin{equation}
					E^f_{xx} = \frac{12}{(z_1 - z_n)^3D^*_{11}}\,.
				\end{equation}

		\subsection*{Statistical analysis}
			\paragraph*{Model selection and fitting.}
				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Houseworth2004, Stone2011glmm}.
				Raw data was processed using a custom script in Python version 3.5.2 \cite{Python}.
				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''}, using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Ape, Nlme}.
				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis (see Supplementary Data).
				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion model of trait evolution [\texttt{R:`ape':corBrownian}] along a phylogeny \cite{Munkemuller2012, Ape}.

				The preferred phylogeny for select North American \textit{Curculio} species is that of Bonal~et~al.~\cite{Bonal2016}.
				This phylogeny was generated using Maximum-Likelihood methods; it is untrametric yet has uniform internal branch lengths.
				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species' root node to represent the individual specimens examined for that species.
				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.

				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data.
				Underestimation of branch lengths causes overestimation of phylogenetic signal.
				This phenomenon was not found to be statistically significant for any of our models.
				Therefore we consider our study's principal findings to be unaffected by branch length underestimation \cite{Munkemuller2012, Molina2017, Stone2011star}.

				Other models of trait evolution were considered during model exploration; including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models \cite{Munkemuller2012, Ape}.
				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance; viz. $R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects.
				This measure is generally used for comparison of covariance structures \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				$R^2_{\beta^*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare fixed effects \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				Using the \texttt{R} package \texttt{`phytools'} \cite{Munkemuller2012, Phytools}, we also estimated the phylogenetic signal in the residual variance of each model; taking into account two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$), and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$).

				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect \cite{Lmtest, R2glmm}.
				Models that incorporated covariance due to Brownian motion consistently produced higher likelihood scores and fit to the data.
				However, no model exhibited statistically significant phylogenetic signal in any of the variables.
				Once models were fitted to the data, T- and F-statistics were calculated to determine whether each cofactor was significantly different from zero.

			\paragraph*{Hypothesis testing.}
				Our objectives for hypothesis testing using the PGLMMs were threefold: \textbf{(1)} to assess whether the altered ratio of exocuticle to endocuticle in the rostral apex has an effect on the tensile strength of the rostrum; \textbf{(2)} to test whether a trade-off exists between specimen stiffness and resistance to fracture; and \textbf{(3)} to examine whether rostrum length and flexibility are correlated.

				To test the relationship between relative layer thicknesses and tensile strength, we fitted a fully-specified model with the cross-sectional area of exocuticle and endocuticle at the site of fracture as fixed effects, including an interaction term, and with the maximum tensile force sustained prior to fracture as a response variable.
				This model was then compared to models with only cross-sectional area of either endocuticle or exocuticle, but not both, as the sole fixed effect in the model.
				We then tested whether one or both regions were significantly correlated to the maximum force sustained in tension, using likelihood-ratio tests and $R^{2}_{\beta^*}$-difference tests between each of the three models.
				Only the cross-sectional area of endocuticle produced a significant fixed effect.
				We therefore elected to make the model more parsimonious by removing cross-sectional area of exocuticle as a fixed effect.
				The final model generated for our first aim featured the ratio of exocuticle to endocuticle in cross-section as a fixed effect, with ultimate tensile strength as the response variable.

				For our second aim we examined the relationship between specimen stiffness and resistance to fracture.
				Specimen stiffness was characterized using a low strain elastic modulus -- averaged across the first 33\% of the stress-strain curve -- and the secant modulus at failure.
				Resistance to fracture was quantified in terms of ultimate strain and fracture toughness, measured as area under the stress-strain curve.
				Four models were fitted, two of which had ultimate strain as the response variable, and two with fracture toughness as response variable.
				Each model then used one of the different measures of specimen stiffness as a fixed effect.

				Finally, for the third aim we explored whether a size-effect might exist in the \textit{Curculio} rostrum; and specifically if longer and typically more curved rostra were more flexible than shorter, straighter rostra.
				We generated two models; the low strain and secant elastic moduli individually served as response variables, with specimen length as the fixed effect in both.

		\subsection*{Code availability}
			R, Python, and Matlab scripts used to manipulate and analyze the raw data (as well as their outputs), produce figures, and estimate effective elastic constants are available from the corresponding author upon request.

		\subsection*{Data availability}
			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 1).
			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 2). PGLMM terms and output are provided in a PDF (Supplement 3). Raw and processed data will be provided by the corresponding author upon reasonable request.

%references for methods only
	\begin{thebibliography}{50}
		\makeatletter
		\addtocounter{NAT@ctr}{45}
		\makeatother
		\section*{References}	
			\bibitem{Arnett1993}
				Arnett Jr., R. H., Samuelson, G. A., and Nishida G. M.
				\textit{The Insect and Spider Collections of the World, 2nd Edition.}
				Fauna and Flora Handbook No. 11.
				(Sandhill Crane Press,
				Gainesville,
				1993).

			\bibitem{Matlab}
				The MathWorks, Inc.,
				\textit{MATLAB and Statistics Toolbox}
				(Natick, Mass.,
				Release R2018b,
				2018).

			\bibitem{Galecki2013}
				Galecki, A. and Burzykowski, T.
				\textit{Linear Mixed-Effects Models Using R.}
				A Step-by-Step Approach
				(Springer,
				New York,
				2013).

			\bibitem{Hadfield2010}
				Hadfield, J. D. and Nakagawa, S.
				General quantitative genetic methods for comparative biology: phylogenies, taxonomies and multi-trait models for continuous and categorical characters
				\textit{J. Evol. Biol.}
				\textbf{23:3,}
				494--508,
				(2010).

			\bibitem{Houseworth2004}
				Housworth, E. A., Martins, E. P., and Lynch, M.
				The Phylogenetic Mixed Model
				\textit{Am. Nat.}
				\textbf{163:1,}
				84--96,
				(2004).

			\bibitem{Stone2011glmm}
				Stone, G. N., Sean, N., and Felsenstein, J.
				Controlling for non-independence in comparative analysis of patterns across populations within species
				\textit{Philos. Trans. Royal Soc. B}
				\textbf{366:1569,}
				1410--1424,
				(2011).

			\bibitem{Python}
				Python Core Team.
				\textit{Python: A dynamic, open source programming language}
				(Python Software Foundation, 
				Version 3.5.2,
				2018).

			\bibitem{Rcite}
				R Development Core Team
				\textit{R: A Language and Environment for Statistical Computing}
				(R Foundation for Statistical Computing,
				Vienna,
				Version 3.5.1,
				2008).

			\bibitem{Ape}
				Paradis, E. and Schliep, K.
				ape 5.0: an environment for modern phylogenetics and evolutionary analyses in R
				\textit{Bioinformatics}
				\textbf{5.0,}
				(2018).

			\bibitem{Nlme}
				Pinheiro, J. et al.
				nlme: Linear and Nonlinear Mixed Effects Models
				\textbf{R package version 3.1-137,}
				(2018).

			\bibitem{Munkemuller2012}
				M{\"u}nkem{\"u}ller, T. et al.
				How to measure and test phylogenetic signal
				\textit{Methods Ecol. Evol.}
				\textbf{3:4,}
				743--756,
				(2012).

			\bibitem{Stone2011star}
				Stone, E. A.
				Why the phylogenetic regression appears robust to tree misspecification
				\textit{Syst. Biol.}
				\textbf{60:3,}
				245--260,
				(2011).

			\bibitem{Molina2017}
				Molina-Venegas, R. and Rodr{\'\i}guez, M. {\'A}.
				Revisiting phylogenetic signal; strong or negligible impacts of polytomies and branch length information?
				\textit{BMC Evol. Biol.}
				\textbf{17:1,}
				1--53,
				(2017).

			\bibitem{Nakagawa2013}
				Nakagawa, S. and Schielzeth, H.
				A general and simple method for obtaining R2 from generalized linear mixed-effects models
				\textit{Methods Ecol. Evol.}
				\textbf{4:2,}
				133--142,
				(2013).

			\bibitem{Jaeger2016}
				Jaeger, B. C., Edwards, L. J., Das, K., and Sen, P. K.
				An $\text{R}^2$ statistic for fixed effects in the generalized linear mixed model
				\textit{J. Appl. Stat.}
				\textbf{44:6,}
				1086--1105,
				(2016).

			\bibitem{R2glmm}
				Jaeger, B.
				Computes R Squared for Mixed (Multilevel) Models
				\textbf{Version 0.1.2}
				(2017).

			\bibitem{Phytools}
				Liam, J. R.
				phytools: An R package for phylogenetic comparative biology (and other things)
				\textit{Methods Ecol. and Evol.}
				\textbf{3,}
				217--223,
				(2012).

			\bibitem{Lmtest}
				Zeileis, A. and Hothorn, T.
				Diagnostic Checking in Regression Relationships
				\textit{R News}
				\textbf{2:3,}
				7--10,
				(2002).
	\end{thebibliography}
\end{document}

### 2019-03-19 tex_v4/manuscript.tex
v4

--- 2019-03-16 tex_v3/JansenEtAL2019-AcornWeevilRostrum-NMFedits-Feb2019.tex
@@
 
+	\begin{figure*}
+		\centering
+		\includegraphics[width=180mm]{fig1.pdf}
+		\caption{\textbf{Morphology and oviposition behavior of female \textit{Curculio} weevils.} 
+			\textbf{a}, Lateral habitus image of female \textit{Curculio~sayi} (Gyllenhal, 1836) featuring the elongate, strongly curved rostrum.
+			\textbf{b}, Lateral view of head of female specimen of \textit{Curculio~longinasus} Chittenden, 1927, with major anatomical features indicated.
+			\textbf{c}, Illustration of oviposition behavior, proceeding from left to right: female makes incision in host fruit, flexes head directly over bore-hole using front legs, then maintains tension on snout while rotating to excavate linear channel into fruit.
+During this process, female rostrum is bent until completely straight.
+		}
+		\label{fig::curculio}
+	\end{figure*}
+
 	Despite its documented performance in many \textit{Curculio} species \cite{Gibson1969, Moffett1989, AguirreUribe1978, Toju2005}, it remains unknown how the rostrum of female acorn weevils can withstand the repeated, often extreme bending incurred during the process of egg-chamber excavation.
@@
 
-	\begin{figure*}
-		\centering
-		\includegraphics[width=180mm]{fig1.pdf}
-		\caption{\textbf{Morphology and oviposition behavior of female \textit{Curculio} weevils.} 
-			\textbf{a}, Lateral habitus image of female \textit{Curculio~sayi} (Gyllenhal, 1836) featuring the elongate, strongly curved rostrum.
-			\textbf{b}, Lateral view of the head of a female specimen of \textit{Curculio~longinasus} Chittenden, 1927, with major anatomical features indicated.
-			\textbf{c}, Illustration of the oviposition behavior, proceeding from left to right: a female makes an incision in the host fruit, flexes the head directly over the bore hole using the front legs, then maintains tension on the snout while rotating to excavate a linear channel into the fruit.
-During this process, the female rostrum is bent until completely straight.
-		}
-		\label{fig::curculio}
-	\end{figure*}
-
-	\textit{\textbf{[NMF note: Feel that intro could be shortened by not "giving away" so many results. Maybe 3-5 phrases less in 2 final paragraphs of intro]}}
 	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition.
-	[Based on our findings, we posit that the composite profile of the rostral apex enables the snout to undergo repeated distal flexion while remaining within the elastic limits of the material, mitigating the risk of structural damage, and without evident alteration of the mechanical properties of the individual components of the cuticle across the structure and between species.]
 	Accordingly, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
@@
 			\caption{\textbf{Gross divisions of cuticle in the female \textit{Curculio} rostrum.}
-				\textbf{a}, Scanning electron micrograph of the head capsule, in frontal view, of female \textit{Curculio~sulcatulus} (Casey, 1897), with the rostrum removed.
-				\textbf{b}, Magnified view of the junction between the rostrum and head capsule, showing the division of the cuticle into two general regions: the exocuticle and the endocuticle.
+				\textbf{a}, Scanning electron micrograph of head capsule, in frontal view, of female \textit{Curculio~sulcatulus} (Casey, 1897), with rostrum removed.
+				\textbf{b}, Magnified view of junction between rostrum and head capsule, showing division of cuticle into two general regions: exocuticle and endocuticle.
 			}
@@
 			\caption{\textbf{Composite profiles of rostral \textit{Curculio} cuticle.}
-				\textbf{a}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~humeralis} (Casey, 1897), showing that the cuticle of the rostral apex is organized as a $\pm45\degree$ angle-ply laminate.
-				\textbf{b}, Scanning electron micrograph of a fractured specimen of \textit{Curculio~caryae}  (Horn, 1873), showing that the cuticle of the rostral base and head capsule has an approximately $30\degree$ stacking angle between each pair of pseudo-orthogonal plies.
-				\textbf{c--d}, Semi-thin sections of cuticle from a specimen of \textit{Curculio humeralis}, stained with toluidine-blue-borax; demonstrating:
-				(\textbf{c}) that the exocuticle of the rostral apex is reduced to a thin shell close to 5 \,{\textmu}m thick, with the endocuticle thickened to maintain a constant laminate thickness; and,
-				(\textbf{d}) that the exocuticle of the head capsule and base of the snout occupies nearly half of the through-thickness of the cuticle.
+				\textbf{a}, Scanning electron micrograph of fractured specimen of \textit{Curculio~humeralis} (Casey, 1897), showing that cuticle of rostral apex is organized as $\pm45\degree$ angle-ply laminate.
+				\textbf{b}, Scanning electron micrograph of fractured specimen of \textit{Curculio~caryae} (Horn, 1873), showing that cuticle of rostral base and head capsule has approximately $30\degree$ stacking angle between each pair of pseudo-orthogonal plies.
+				\textbf{c--d}, Semi-thin sections of cuticle from specimen of \textit{Curculio humeralis}, stained with toluidine-blue-borax; demonstrating:
+				(\textbf{c}) that exocuticle of rostral apex is reduced to thin shell close to 5\,{\textmu}m thick, with endocuticle thickened to maintain a constant laminate thickness; and, 
+				(\textbf{d}) that exocuticle of head capsule and base of snout occupies nearly half of through-thickness of cuticle.
 			}
@@
 			\caption{\textbf{Tensile properties of the female \textit{Curculio} rostrum.}
-				Each plot shows the relationship between two variables as predicted by a phylogenetic linear mixed-effect model; with species as a random effect and a variance-covariance matrix generated from Brownian motion over the preferred phylogeny of \cite{Bonal2016}.
-				Gray regions represent the prediction interval and bootstrapped 95\% confidence interval of the model.
-				The estimated fixed effect $\hat{\beta}$ is given, along with the p-value of a t-test assessing whether $\hat{\beta}$ is significantly different from zero.
-				A generalized marginal $\text{R}^2_{\beta^*}$ for assessing fixed effects is also reported.
+				Each plot shows relationship between two variables as predicted by phylogenetic linear mixed-effect model; with species as random effect and variance-covariance matrix generated from Brownian motion over preferred phylogeny of \cite{Bonal2016}.
+				Gray regions represent the prediction interval and bootstrapped 95\% confidence interval of model.
+				The estimated fixed effect $\hat{\beta}$ is given, along with p-value of t-test assessing whether $\hat{\beta}$ is significantly different from zero.
+				Generalized marginal $\text{R}^2_{\beta^*}$ for assessing fixed effects is also reported.
 				In general, increased endocuticle thickness is associated with greater tensile strength, and stiffness is inversely correlated with toughness.
@@
 		Although CLPT predicts a positive association between the proportion of exocuticle and stiffness of a generalized cuticle, we found no evidence of correspondence between the cross-sectional properties of the fracture site and the overall performance of the rostrum.
-		Because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to properties of the entire rostrum.
-		\textit{\textbf{[NMF note: How surprising/unexpected is this? If a lot -- state so..]}}
+		This result is not too surprising however; because the cross-sectional areas of the cuticle regions vary across the length of the head along an anterior-posterior gradient, it is not possible to correlate measurements from the fracture surface to properties of the entire rostrum.
 
@@
 			\caption{\textbf{Fatigue testing of a female \textit{Curculio caryae} rostrum.}
-				\textbf{a}, Set-up of fatigue testing, with the female's head capsule fixed to a pedestal and the rostrum attached to a strip of rip-stop nylon fabric using cyanoacrylate adhesive.
-				The specimen was loaded in tension, as opposed to compression, therefore isolating the effect of tension on the fatigue behavior of the rostrum.
-				\textbf{b}, Overlay of the pre- and post-fatigue states of the head, showing a clear (short-term) effect from repeated prescribed strain.
-				\textbf{c}, Three-photograph sequence of pre- and post-fatigue states of the rostrum; with the left and central panels showing the conditions immediately prior to and after the testing, respectively, whereas the right panel shows the head having returned to its original shape after 24 hours off soaking in a water/ethanol mixture.
-Hence the immediate post-fatigue shape is not permanent.
-				\textbf{d}, Raw force data (left) and displacement data (right) plots for the fatigue test.
-				The displacement data plot shows clear viscoelastic behavior, indicated by histeresis in the stress-strain response of the specimen, and exhibits a logarithmic decrease stress amplitude over time.
+				\textbf{a}, Set-up of fatigue testing, with female's head capsule fixed to pedestal and rostrum attached to strip of rip-stop nylon fabric using cyanoacrylate adhesive.
+				Specimen was loaded in tension, as opposed to compression, therefore isolating effect of tension on fatigue behavior of rostrum.
+				\textbf{b}, Overlay of pre- and post-fatigue states of head, showing clear (short-term) effect from repeated prescribed strain.
+				\textbf{c}, Three-photograph sequence of pre- and post-fatigue states of the rostrum; with left and central panels showing conditions immediately prior to and after testing, respectively, whereas right panel shows head having returned to its original shape after 24 hours off soaking in a water/ethanol mixture.
+Hence immediate post-fatigue shape is not permanent.
+				\textbf{d}, Raw force data (left) and displacement data (right) plots for fatigue test.
+				Displacement data plot shows clear viscoelastic behavior, indicated by histeresis in stress-strain response of specimen, and exhibits logarithmic decrease stress amplitude over time.
 			}
@@
 			\includegraphics[width=180mm]{fig6.pdf}
-			\caption{\textbf{Fractography of the female \textit{Curculio} rostrum.}
-				\textbf{a}, Light micrograph showing the fracture surface of a tensile tested \textit{Curculio caryae} rostrum, displaying a typical failure mode; illustrated in (\textbf{b}):
-				The small, black arrows indicate the winding direction of the macrofiber lamina; green arrows and dotted lines indicate the direction of crack.
+			\caption{\textbf{Fractography of female \textit{Curculio} rostrum.}
+				\textbf{a}, Light micrograph showing the fracture surface of tensile tested \textit{Curculio caryae} rostrum, displaying typical failure mode; illustrated in (\textbf{b}):
+				Small, black arrows indicate the winding direction of macrofiber lamina; green arrows and dotted lines indicate the direction of crack.
 				Scanning electron micrographs highlight: (\textbf{c}) shear cusp formation; (\textbf{d}) tensile failure and off-axis macrofiber fiber-pulling; (\textbf{e}) interlaminar delamination; (\textbf{f}-\textbf{h}) crack formation near invaginated exocuticle; and (\textbf{g}) ventral crack-front coalescence and shear cusp formation.

### 2019-05-03 tex_v4/manuscript.tex
post-defense, corrections and revisions

@@
 		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
-		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
+		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and toughness of the distal portion of the snout.
 		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
@@
 		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
-		Rostrum stiffness is shown to be inversely correlated with fracture toughness; thus allowing a highly curved rostrum to completely straighten without structural damage.
+		Rostrum stiffness is shown to be inversely correlated with work of fracture; thus allowing a highly curved rostrum to completely straighten without structural damage.
 		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
-
 	\end{abstract}
@@
 		The specimen did not show any evidence of fractures, micro-tears, or shear cusps anywhere in the surface of the exocuticle.
-		Moreover, the tensile strength of the tested specimen's rostrum was consistent with that of other species members ($F_max~=~5.02~\,\text{N}$).
+		Moreover, the tensile strength of the tested specimen's rostrum was consistent with that of other species members ($F_{max}~=~5.02~\,\text{N}$).
 		Surprisingly, the specimen remained undamaged by the testing.

### 2019-05-03 tex_v5/manuscript.tex
post-defense, corrections and revisions

--- 2019-05-03 tex_v4/manuscript.tex
@@
 \usepackage{xcolor}
+\usepackage{microtype}
 \usepackage[colorlinks]{hyperref}
@@
 
-		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness (see Fig.~\ref{fig::tensile}).
+		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and work of fracture (see Fig.~\ref{fig::tensile}).
 		This implies that stiffer specimens -- and by extension, stiffer cuticle profiles -- are generally more brittle.
@@
 		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
-		From an evoluationary perspective, these findings reveal an unexpected morphological source of constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
+		From an evolutionary perspective, these findings reveal an unexpected morphological source of constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
 		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion.
@@
 					\begin{Bmatrix}
-					\{\epsilon^{\degree}\} \\
+					\{\epsilon^{\circ}\} \\
 					\{\kappa\}
@@
 					\begin{aligned}
-						\{N\} & = [A]\{\epsilon^{\degree}\}\,, \\
+						\{N\} & = [A]\{\epsilon^{\circ}\}\,, \\
 						\{M\} & = [D]\{\kappa\}\,,						
@@
 						\begin{Bmatrix}
-							\epsilon^{\degree}_{xx} \\
-							\epsilon^{\degree}_{yy} \\
-							\gamma^{\degree}_{xy}
+							\epsilon^{\circ}_{xx} \\
+							\epsilon^{\circ}_{yy} \\
+							\gamma^{\circ}_{xy}
 						\end{Bmatrix}\,,
@@
 			\paragraph*{Model selection and fitting.}
-				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Houseworth2004, Stone2011glmm}.
+				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Housworth2004, Stone2011glmm}.
 				Raw data was processed using a custom script in Python version 3.5.2 \cite{Python}.
-				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''}, using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Ape, Nlme}.
-				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis (see Supplementary Data).
+				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''}, using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Nlme, Ape}.
+				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis.
 				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
@@
 				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.
-
-				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data.
-				Underestimation of branch lengths causes overestimation of phylogenetic signal.
-				This phenomenon was not found to be statistically significant for any of our models.
-				Therefore we consider our study's principal findings to be unaffected by branch length underestimation \cite{Munkemuller2012, Molina2017, Stone2011star}.
-
-				Other models of trait evolution were considered during model exploration; including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models \cite{Munkemuller2012, Ape}.
-				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance; viz.
-$R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
+				Underestimation of branch lengths causes overestimation of phylogenetic signal \cite{Munkemuller2012, Molina2017, Stone2011star}.
+				We were unable to detect statistically significant phylogenetic signal for any of our models.
+				For this reason we consider our study's principal findings to be unaffected by branch length underestimation.
+				However, because of the relatively poor taxonomic resolution afforded by a sample of five species, we are unable to conclude whether significant phylogenetic signal exists in the traits examined herein.
+
+				Other models of trait evolution were considered during model exploration; including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) models [\texttt{R:`ape':corBlomberg}] \cite{Munkemuller2012, Ape}.
+				These models were chosen to compare among competing hypotheses of trait evolution, namely stabilizing selection and variable-rate evolution, respectively.
+				We additionally tested a Brownian motion model of trait evolution, corresponding to variation constrained by clade membership (i.e., sister taxa are more similar to each other than to an out-group).
+				However, neither of the former produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance; viz.
+$R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}].
 				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects.
@@
 				However, no model exhibited statistically significant phylogenetic signal in any of the variables.
+				Our findings imply that \emph{residual variance} in the examined traits does not exhibit significant rate-variation or stabilizing selection in the amount or direction of residual variance between groups; rather, variation within and between clades is distributed randomly by clade, with limited phylogenetic dependence.
 				Once models were fitted to the data, T- and F-statistics were calculated to determine whether each cofactor was significantly different from zero.
@@
 				Specimen stiffness was characterized using a low strain elastic modulus -- averaged across the first 33\% of the stress-strain curve -- and the secant modulus at failure.
-				Resistance to fracture was quantified in terms of ultimate strain and fracture toughness, measured as area under the stress-strain curve.
-				Four models were fitted, two of which had ultimate strain as the response variable, and two with fracture toughness as response variable.
+				Resistance to fracture was quantified in terms of ultimate strain and work of fracture, measured as area under the stress-strain curve.
+				Four models were fitted, two of which had ultimate strain as the response variable, and two with work of fracture as response variable.
 				Each model then used one of the different measures of specimen stiffness as a fixed effect.
@@
 
-			\bibitem{Houseworth2004}
+			\bibitem{Housworth2004}
 				Housworth, E. A., Martins, E. P., and Lynch, M.

### 2019-05-19 tex_v5/manuscript.tex
MS ready for submission

@@
 		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
+		We conclude that avoidance of catastrophic structural failure, as initiated in these sutures under tension, is the driving selective pressure in the evolution of the female \textit{Curculio} rostrum.
 	\end{abstract}
@@
 	%title and author block
-	{\title{Exoskeletal strength and cuticle composite profile of the acorn weevil rostrum}
+	{\title{Avoidance of catastrophic structural failure as an evolutionary constraint: Biomechanics of the acorn weevil rostrum}
 	\date{\today}
@@
 
-	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition.
 	Accordingly, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
 	This is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (though see \cite{Matsumura2017}).
+	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition, particularly in the fracture-prone and exocuticle-rich occipital sulci.
+	Based on these findings we conclude that avoidance of catastrophic structural failure is the primary selective pressure in the evolution of the female \textit{Curculio} rostrum.
 
@@
 		\subsection*{Code availability}
-			R, Python, and Matlab scripts used to manipulate and analyze the raw data (as well as their outputs), produce figures, and estimate effective elastic constants are available from the corresponding author upon request.
+			R, Python, and Matlab scripts used to manipulate and analyze the raw data and processed data, produce figures, and estimate effective elastic constants are provided in Supplement 1.
 
 		\subsection*{Data availability}
-			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 1).
-			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 2).
-PGLMM terms and output are provided in a PDF (Supplement 3).
+			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 2).
+			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 3).
+PGLMM terms and output are provided in a PDF (Supplement 4).
+All measurements used for PGLMM generation are provided as a CSV file (Supplement 5).
 Raw and processed data will be provided by the corresponding author upon reasonable request.

### 2019-05-19 analysis/phylo_regress.R

@@ -697,7 +697,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$U)
   raw.data$cofactor = log(raw.data$Etanl)
-  comparison = c('ln(Fracture Toughness [Jm-3])', 'ln(Elastic Modulus [MPa])')
+  comparison = c('ln(Work of Fracture [Jm-3])', 'ln(Elastic Modulus [MPa])')
   Analysis3(phylo.vcv, tree.star, raw.data, comparison)
   Analysis4(phylo.vcv, raw.data, comparison)
 
@@ -710,7 +710,7 @@ main = function() {
   writeLines(strrep('=',80))
   raw.data$phenotype = log(raw.data$U)
   raw.data$cofactor = log(raw.data$Esec)
-  comparison = c('ln(Fracture Toughness [Jm-3])', 'ln(Secant Modulus [MPa])')
+  comparison = c('ln(Work of Fracture [Jm-3])', 'ln(Secant Modulus [MPa])')
   Analysis3(phylo.vcv, tree.star, raw.data, comparison)
   Analysis4(phylo.vcv, raw.data, comparison)
 

### 2019-05-19 tex_v5/supplementary data/S1/effective_moduli.m

@@ -1,0 +1,190 @@
+%Author: Andrew Jansen
+%Date: January 2019
+%Title: Effective moduli of Curculio cuticle
+
+%--------------------------------------------------------------------------
+%Definition of elastic constants by cuticle region
+%--------------------------------------------------------------------------
+%Elastic constants for exocuticle, denoted by 'x'
+%E denotes Young's modulus along material axis denoted by number
+Ex1 = 6.4;
+Ex2 = 6.4;
+
+%nu denotes Poisson ratio along material axes denoted by numbers
+nux12 = 0.36;
+nux21 = 0.36;
+
+%G denotes shear modulus along material axes denoted by numbers
+Gx12 = 2.5;
+
+%Matrix of constants for exocuticle
+exo = [Ex1 Ex2 nux12 nux21 Gx12];
+
+%--------------------------------------------------------------------------
+%For macrofiber (a.k.a. 'balken') layer in endocuticle, denoted by 'b'
+Eb1 = 8.5;
+Eb2 = 0.52;
+nub12 = 0.49;
+nub21 = 0.03;
+Gb12 = 0.17;
+
+endo = [Eb1 Eb2 nub12 nub21 Gb12];
+
+cuticle = [exo; endo];
+%--------------------------------------------------------------------------
+%Cuticle lay-up: layer thicknesses, orientation angles & stacking sequence  
+%--------------------------------------------------------------------------
+%sequence for basal portion of head
+base_z = (1 / 480) * [-24 -12:-1 1:12 24]; %48 layers, 50 microns => mm
+base_theta = 30 * [ 3:8 -(-5:0); 0:5 -(-8:-3)]; %30 deg angle sequence
+%subtract 5 deg  after matrix to maximize flexural stiffness
+base_seq = [1, repelem(2, 24) 1]; %1 for exocuticle, 2 for endocuticle
+base_layup = [base_z; 0 base_theta(1:24) 0; base_seq];
+
+%--------------------------------------------------------------------------
+%sequence for apical portion of head
+apex_z = (1 / 800) * [-40 -36:4:-4 -3:-1 1:3 4:4:36 40];
+apex_theta = [0 repelem(-45, 6) repelem(45, 6);
+    repelem(45, 6) repelem(-45, 6) 0];
+apex_seq = [1, repelem(2, 24) 1]; %same sequence as base
+apex_layup = [apex_z; apex_theta(1:26); apex_seq];
+
[…]

### 2019-05-19 tex_v5/supplementary data/S1/phylo_regress.R

@@ -1,0 +1,748 @@
+# Andrew Jansen
+# created x.10.2018
+
+# Miscellaneous Functions ------------------------------------------------------
+# Import the data
+GetData = function(file.name) {
+  test.data = data.frame(read.csv(file.name, sep = ','))
+  return(test.data)
+}
+
+# Analysis 1 -------------------------------------------------------------------
+# Fits models with and without log transformation and phylogenetic covariance
+Analysis1 = function(phylogeny, raw.data) {
+  
+  # Fit models with variable transformation and phylogenetic covariance
+  simple.model = lme(phenotype ~ cofactor,
+                     random = ~ 1 | species,
+                     data = raw.data,
+                     method = 'ML')
+  
+  simple.brownian = lme(phenotype ~ cofactor,
+                        random = ~ 1 | species,
+                        data = raw.data,
+                        correlation = phylogeny,
+                        method = 'ML')
+  
+  y.transformed = lme(log(phenotype) ~ cofactor,
+                           random = ~ 1 | species,
+                           data = raw.data,
+                           method = 'ML')
+  
+  y.brownian = lme(log(phenotype) ~ cofactor,
+                             random = ~ 1 | species,
+                             data = raw.data,
+                             correlation = phylogeny,
+                             method = 'ML')
+  
+  x.transformed = lme(phenotype ~ log(cofactor),
+                      random = ~ 1 | species,
+                      data = raw.data,
+                      method = 'ML')
+  
+  x.brownian = lme(phenotype ~ log(cofactor),
+                   random = ~ 1 | species,
+                   data = raw.data,
+                   correlation = phylogeny,
+                   method = 'ML')
+  
+  xy.transformed = lme(log(phenotype) ~ log(cofactor),
+                       random = ~ 1 | species,
+                       data = raw.data,
+                       method = 'ML')
+  
+  xy.brownian = lme(log(phenotype) ~ log(cofactor),
+                    random = ~ 1 | species,
+                    data = raw.data,
+                    correlation = phylogeny,
+                    method = 'ML')
+  
[…]

### 2019-05-19 tex_v5/supplementary data/S1/tensile.py

@@ -1,0 +1,469 @@
+# -*- coding: utf-8 -*-
+'''
+Created on Tue Aug  1 13:59:50 2017
+
+@author: Flashcock
+'''
+
+import os
+from scipy.stats import linregress as linreg
+from itertools import takewhile
+from numpy import array, poly1d, sqrt, absolute, arcsin, pi
+from scipy.optimize import curve_fit, minimize_scalar
+from scipy.integrate import quad
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from mpl_toolkits.mplot3d import Axes3D
+from numpy import log as ln
+
+class specimen:
+    '''
+    Args should be formatted as follows:
+        asuid (str or int), species (str), length (str), exo (str), endo (str),
+        total (str), N (list [float]), mm (list [float]), Sec (list [float]).
+    These are converted to their respective types below and used to calculate
+        true and engineering stress and strain, young's modulus, and UTS.
+    '''
+    def __init__(self, asuid, species, length, exo, endo, total, N, mm, Sec):
+        self.__asuid__ = asuid
+        self.__species__ = species.lower()
+
+        self.__length__ = float(length)
+        self.__exo__ = float(exo)
+        self.__endo__ = float(endo)
+        self.__total__ = float(total)
+
+        self.__N__ = array(N)
+        self.__mm__ = array(mm) - min(mm)
+        self.__Sec__ = array(Sec) - Sec[0]
+
+        self.__sigmaN__ = self.__N__ / self.__total__ #P/A0
+        self.__epsilonN__ = (self.__mm__ / self.__length__) #dL/L0
+        self.__sigmaT__ = self.__sigmaN__ * (1 + self.__epsilonN__) #s(1+e)
+        self.__epsilonT__ = ln(1 + self.__epsilonN__) #ln(1+e)
+
+        self.__fmax__ = max(N)
+        self.__uts__ = max(self.__sigmaN__) #evaluated from engineering stress
+
+        self.__dmax__ = max(self.__mm__)
+        self.__ufs__ = max(self.__epsilonN__) #evaluated from engineering strain
+
+
+        self.__Eindex__ = []
+        lower = self.__ufs__ * 0
+        upper = self.__ufs__ * 0.33
+        self.__Eindex__.append(len([x for x in takewhile(lambda x: x[1] <= lower, enumerate(self.__epsilonN__))]))
+        self.__Eindex__.append(len([y for y in takewhile(lambda y: y[1] <= upper, enumerate(self.__epsilonN__))]))
[…]

### 2019-05-20 cover letter/cover_letter.tex
full submission

\documentclass[10pt,letterpaper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[letterpaper,left=1in, right=1in, top=1in, bottom=1in]{geometry}

\makeatletter
\let\@texttop\relax
\makeatother

\address{Andrew Jansen\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501} 
\signature{Andrew Jansen, PhD\\
Corresponding Author – majanse1@asu.edu\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA} 
\begin{document} 
\begin{letter}{Xin Li\\
Associate Editor\\
xin.li@nature.com\\
}
\opening{Dear Dr. Li,}

I am writing to inquire about the suitability of our manuscript entitled, ``Avoidance of catastrophic structural failure as an evolutionary constraint: Biomechanics of the acorn weevil rostrum'', for consideration for publication in Nature Materials.
The primary focus of the manuscript is the characterization of tensile and failure mechanics of the exoskeleton (a laminate composite) in a type of beetle (acorn weevils).
I believe that the methods and results we present would be an outstanding fit for your readers, while also having the potential to reach an expanded audience comprised of evolutionary biologists and biomechanists.
Our work is novel in showing that the composite microstructure of the exoskeleton of acorn weevils is locally modified to enhance both flexibility and toughness of a part of the head called the rostrum, something never before observed in insects.
In addition, we find that the modifications to the material are likely an evolutionary response to circumvent unavoidable, brittle features inherent to the anatomy the structure.

We believe that our insights are very impactful: these results are the first of their kind to conclusively demonstrate that the \emph{composite microstructure} of the exoskeleton, rather than its thickness or elastomer composition, can be optimized across species for avoidance of structural failure during extreme bending.
Our results also highlight long-ignored and under-appreciated design patterns in the composite structure of the cuticle as an essential feature of the functionality of insect exoskeletons.
This study shows how the laminate organization of the cuticle, and local variations in particular, are intimately linked to the gross morphology of macroscale structures and are of critical importance to understanding the mechanical behavior and evolution of the exoskeleton both as a material and source of biomechanical constraint upon these fascinating organisms.

I hope you will consider evaluating the manuscript to see if it falls within your journal’s scope.
Our title and abstract are copied below; I have also attached the full manuscript, in case you or another editor would like to read further.
All of the authors have agreed to submit the manuscript to Nature Materials; we feel that this journal will permit us to reach the ideal readership to disseminate our findings.
If you have any questions about the manuscript, please let me know.

Thank you very much, and we look forward to your reply.

\closing{Sincerely,}

\newpage

\cc{Jason Williams\\
Co-author\\
Research Associate\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nikhilesh Chawla\\
Co-author\\
Professor\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nico Franz\\
Co-author\\
Professor\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA}

\bigskip

\textbf{Title:}
Avoidance of catastrophic structural failure as an evolutionary constraint: Biomechanics of the acorn weevil rostrum

\textbf{Abstract:}
The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and toughness of the distal portion of the snout.
Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
Rostrum stiffness is shown to be inversely correlated with work of fracture; thus allowing a highly curved rostrum to completely straighten without structural damage.
Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
We conclude that avoidance of catastrophic structural failure, as initiated in these sutures under tension, is the driving selective pressure in the evolution of the female \textit{Curculio} rostrum.
%\ps{adding a postscript} 
%\encl{list of enclosed material} 
\end{letter} 
\end{document}
